### 2016-01-08 paper.tex
Draft of introduction

\section{Introduction}

Reactive systems are ubiquitous in real-world problems such as circuit design,
industrial automation, or device drivers. Automatic synthesis can provide a
\textit{correct by construction} controller for a reactive system from a
specification.  However, the reactive synthesis problem is 2EXPTIME-complete so
naive algorithms are infeasible on even simple systems.

Reactive synthesis is formalised as a game between the \texit{system} and its
\texit{environment}. In this work we focus on safety games, in which the system
must prevent the environment from forcing the game into an error state. Much of
the complexity of reactive synthesis stems from tracking the set of states in
which a player is winning.

There are several techniques that aim to mitigate this complexity by
representing states symbolically.  Historically the most successful technique
has been to use \texit{Binary Decision Diagrams} (BDDs).  BDDs efficiently
represent a relation on a set of game variables but in the worst case the
representation may be exponential. This means that BDDs are not a
one-size-fits-all solution for all reactive synthesis specifications.

Advances in SAT solving technology has prompted research into its applicability
to synthesis as an alternative to BDDs. One approach is to find sets of states
in CNF \cit{demiurge}. Another approach is to eschew states and focus on
\texit{runs} of the game. Previous work has applied this idea to solving
bounded games \cite{nina} by forming abstract trees of individual runs. In this
paper, we extend this idea to unbounded games by constructing approximate sets
of winning states from abstract trees.

### 2016-01-10 paper.tex
Formatting

@@
+\documentclass{llncs}
+
+\begin{document}
+
+\title{Untitled CAV Paper}
+
+\author{Alexander Legg\inst{1} 
+    \and Leonid Ryzhyk\inst{2}
+    \and Nina Narodytska\inst{2}}
+
+\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
+    Communications and the Digital Economy and the Australian Research Council through the ICT
+    Centre of Excellence program.} and UNSW \\
+    \email{alexander.legg@nicta.com.au}
+    \and Samsung Research}
+
+\maketitle
+
+\begin{abstract}
+    This is the abstract
+\end{abstract}
+
 \section{Introduction}
@@
 industrial automation, or device drivers. Automatic synthesis can provide a
-\textit{correct by construction} controller for a reactive system from a
+\emph{correct by construction} controller for a reactive system from a
 specification.  However, the reactive synthesis problem is 2EXPTIME-complete so
@@
 
-Reactive synthesis is formalised as a game between the \texit{system} and its
-\texit{environment}. In this work we focus on safety games, in which the system
+Reactive synthesis is formalised as a game between the \emph{system} and its
+\emph{environment}. In this work we focus on safety games, in which the system
 must prevent the environment from forcing the game into an error state. Much of
@@
 representing states symbolically.  Historically the most successful technique
-has been to use \texit{Binary Decision Diagrams} (BDDs).  BDDs efficiently
+has been to use \emph{Binary Decision Diagrams} (BDDs).  BDDs efficiently
 represent a relation on a set of game variables but in the worst case the
@@
 to synthesis as an alternative to BDDs. One approach is to find sets of states
-in CNF \cit{demiurge}. Another approach is to eschew states and focus on
-\texit{runs} of the game. Previous work has applied this idea to solving
+in CNF \cite{demiurge}. Another approach is to eschew states and focus on
+\emph{runs} of the game. Previous work has applied this idea to solving
 bounded games \cite{nina} by forming abstract trees of individual runs. In this
@@
 of winning states from abstract trees.
+
+\section{Reactive Synthesis}
+
+\end{document}

### 2016-01-10 paper.tex
Definitions of games/synthesis

@@
 
-Reactive synthesis is formalised as a game between the \emph{system} and its
-\emph{environment}. In this work we focus on safety games, in which the system
-must prevent the environment from forcing the game into an error state. Much of
-the complexity of reactive synthesis stems from tracking the set of states in
-which a player is winning.
+Reactive synthesis is formalised as a game between the \emph{controller} and
+its \emph{environment}. In this work we focus on safety games, in which the
+controller must prevent the environment from forcing the game into an error
+state.  Much of the complexity of reactive synthesis stems from tracking the
+set of states in which a player is winning.
 
@@
 in CNF \cite{demiurge}. Another approach is to eschew states and focus on
-\emph{runs} of the game. Previous work has applied this idea to solving
-bounded games \cite{nina} by forming abstract trees of individual runs. In this
-paper, we extend this idea to unbounded games by constructing approximate sets
-of winning states from abstract trees.
+\emph{runs} of the game. Previous work has applied this idea to realizability
+of bounded games \cite{nina} by forming abstract representations of the game.
+In this paper, we extend this idea to unbounded games by constructing
+approximate sets of winning states from abstract trees.
 
@@
 
+A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
+consists of a set of state variables, sets of controllable and uncontrollable
+label variables, a transition relationship $\delta : (X, L_c, L_u) \to X$, an
+initial state, and an error state. The \emph{controller} and \emph{environment}
+players choose controllable and uncontrollable labels respectively and the game
+proceeds according to $\delta$.
+
+An \emph{run} of a game $(x_0, c_0, u_0), (x_1, c_1, u_1) \dots (x_n, c_n,
+u_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
+\leftarrow \delta(x_k, c_k, u_k)$. A run is winning for the controller if $x_0
+= I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
+all runs are restricted to length $n$, whereas unbounded games consider runs of
+infinite length.
+
+A \emph{controller strategy} $\pi^c : X \to L_c$ is a mapping of states to
+controllable labels. A controller strategy is winning in a bounded game of rank
+$n$ if all runs $(x_0, \pi^c(x_0), u_0), (x_1, \pi^c(x_1), u_1) \dots (x_n,
+\pi^c(x_n), u_n)$ are winning. Bounded \emph{realizability} is the problem of
+determining the existence of such a strategy for a bounded game.
+
+An \emph{environment strategy} $\pi^e : (X, L_c) \to L_u$ is a mapping of
+states and controllable labels to uncontrollable labels. A bounded run is
+winning for the environment if $x_0 = I \land \exists i \in \{1..n\} (x_i = E)$
+and an environment strategy is winning for a bounded game if there exists a run
+$(x_0, c_1, \pi^e(x_1, c_1)), (x_1, c_1, \pi^e(x_1, c_1)) \dots (x_n, c_n,
+\pi^e(x_n, c_n))$ that wins for the environment. Safety games are zero sum,
+therefore the existence of a controller strategy implies the nonexistence of an
+environment strategy and vice versa.
+
+A set of runs for a bounded game of rank $n$ can be represented by a \emph{game
+tree} of depth $n$ with states in the nodes and labels as edges. A
+\emph{strategy tree} similarly represents a strategy by restricting the edges
+in a game tree to obey $\pi^c$ for a controller strategy or $\pi^e$ for an
+environment strategy.
+
+\subsection{Abstract Game Trees}
+
+The bounded synthesis algorithm from \cite{nina} constructs abstractions of the
+game in the form of \emph{abstract game trees}.
+
 \end{document}

### 2016-01-17 paper.tex
Added a section on abstract game trees

@@
 
-A set of runs for a bounded game of rank $n$ can be represented by a \emph{game
-tree} of depth $n$ with states in the nodes and labels as edges. A
-\emph{strategy tree} similarly represents a strategy by restricting the edges
-in a game tree to obey $\pi^c$ for a controller strategy or $\pi^e$ for an
-environment strategy.
-
 \subsection{Abstract Game Trees}
 
-The bounded synthesis algorithm from \cite{nina} constructs abstractions of the
-game in the form of \emph{abstract game trees}.
+A set of runs can be symbolically represented by a tree of labels. Each edge of
+the tree is either a \emph{fixed} valuation of controllable or uncontrollable
+variables, or it is \emph{unfixed} denoting any values are possible. Also, the
+tree must alternate between controllable and uncontrollable edges to reflect
+the alternating turns of the game. The entire set of runs of a bounded game of
+rank $n$ is symbolically represented by a tree of depth $2n$ and width 1
+populated by unfixed edges. Reducing the set of runs in a game forms an
+abstract game, which can be represented symbolically by an \emph{abstract game
+tree}.
+
+A strategy is equivalent to the set of all runs with the player's labels
+obeying the strategy mapping $\pi$. Therefore, a strategy can also be
+represented by a tree. Relaxing the restriction of the strategy mapping allows
+for a \emph{partial strategy} in which multiple labels are now possible for a
+single state.
+
+A strategy or partial strategy can also be thought of as an abstract game. A
+partial strategy for the controller is a restriction only on the controllable
+labels in the game. So if the environment can not win in the abstract game
+equivalent to the controller's partial strategy, then all strategies allowable
+by that partial strategy must be winning.
 

### 2016-01-20 paper.tex
Added treeFormula algorithms and some explanation

@@
 \documentclass{llncs}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\pagestyle{headings}
 
@@
 
+\subsection{Solving Bounded Runs}
+
+An abstract game tree can be checked for a winning run by a SAT solver.  The
+tree must be encoded into CNF by making copies of the transition relation for
+each game step in tree. The formula must also check whether the error state has
+been reached in no branch/in all branches for the controller and environment
+respectively. The functions that produce these formulas are shown in Algorithm
+\ref{alg:treeFormula}.
+
+\begin{algorithm}
+    \caption{Tree formulas for Controller and Environment respectively}
+    \label{alg:treeFormula}
+    \begin{algorithmic}
+        \Function{treeFormula}{gt}
+        \If{$\Call{rank}{gt} == 0$}
+        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
+        \Else
+        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \Function{$\overline{treeFormula}$}{gt}
+        \If{$\Call{rank}{gt} == 0$}
+        \State \Return{\Call{E}{$x^{gt}$}}
+        \Else
+        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n})$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+When we produce a formula using $\textsc{treeFormula}$ the SAT solver is
+playing on behalf on the controller. Any unfixed labels, controllable or
+uncontrollable, in the tree will be existentially quantified. This means that
+if there exists a way for the game to be solved when both players cooperate the
+SAT solver will find it.
+
+\subsection{Counterexample Guided Bounded Synthesis}
+
 \end{document}

### 2016-01-20 paper.tex
Added some game tree diagrams and explanation

@@
 \usepackage{algpseudocode}
+\usepackage{subcaption}
+\captionsetup{compatibility=false}
+\usepackage{tikz}
+\usetikzlibrary{trees}
+
 \pagestyle{headings}
@@
 
-\subsection{Solving Bounded Runs}
-
-An abstract game tree can be checked for a winning run by a SAT solver.  The
-tree must be encoded into CNF by making copies of the transition relation for
-each game step in tree. The formula must also check whether the error state has
-been reached in no branch/in all branches for the controller and environment
-respectively. The functions that produce these formulas are shown in Algorithm
-\ref{alg:treeFormula}.
+An abstract game tree can be checked for the existence of a winning run by a
+SAT solver.  The tree must be encoded into CNF by making copies of the
+transition relation for each game step in tree. The formula must also check
+whether the error state has been reached in no branch/in all branches for the
+controller and environment respectively. The functions that produce these
+formulas are shown in Algorithm \ref{alg:treeFormula}.
 
@@
 When we produce a formula using $\textsc{treeFormula}$ the SAT solver is
-playing on behalf on the controller. Any unfixed labels, controllable or
-uncontrollable, in the tree will be existentially quantified. This means that
-if there exists a way for the game to be solved when both players cooperate the
-SAT solver will find it.
+playing on behalf on the controller. Any unfixed labels in the tree,
+controllable or uncontrollable, will be existentially quantified.  This means
+that if there exists a way for the game to be solved when both players
+cooperate the SAT solver will find it. If no winning run exists in an abstract
+game even when the players are cooperating then there is definitely no winning
+run when the opponent is playing adversarily.
 
@@
 
+By checking for the existence of a winning run for the environment in an
+abstract game tree constructed from a partial controller strategy we are
+checking if the partial strategy is winning. If no spoiling run can be found
+then the partial strategy must always win. If a spoiling run is found then we
+then we have a counterexample that proves that the controller strategy is not
+winning. This forms the basis of a counterexample guided abstraction refinement
+framework that operates on candidate strategies.
+
+Synthesis begins with an empty game as seen in Figure \ref{fig:emptytree}.
+Initially we are playing on behalf of the environment because it chooses the
+first move in each step. The empty game is passed to the SAT solver, which
+searches for a candidate environment strategy.  If a candidate is found (Figure
+\ref{fig:candidate}) then it is checked for a spoiling strategy by solving for
+the controller. If no spoiling strategy exists, that means our candidate is a
+winning strategy and the algorithm terminates.  Otherwise we find a
+counterexample (Figure \ref{fig:cex}), which is used to refine the empty game
+tree to include the first move from the controller's spoiling strategy (Figure
+\ref{fig:refined}.  The algorithm continues by finding a new candidate for the
+environment (Figure \ref{fig:candidate2}), and a new counterexample to refine
+the game abstraction once again (Figure \ref{fig:refined2}).
+
+\tikzset{every node/.style={solid}}
+\tikzstyle{fixed}=[solid]
+\tikzstyle{unfixed}=[dash pattern = on 2pt off 2pt]
+\begin{figure}
+    \centering
+    \begin{subfigure}{.3\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {} edge from parent[unfixed]
+                    child {node [circle,draw] {} edge from parent[unfixed]}
+                };
+        \end{tikzpicture}
+        \caption{An empty AGT}
+        \label{fig:emptytree}
+    \end{subfigure}%
+    \begin{subfigure}{.35\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root) {}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {} edge from parent[unfixed]}
+                    edge from parent [fixed] node [left] {$u_1$}
+                };
+        \end{tikzpicture}
+        \caption{Environment candidate}
+        \label{fig:candidate}
+    \end{subfigure}%
+    \begin{subfigure}{.4\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {} edge from parent[unfixed]
+                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+                    edge from parent [fixed] node [left] {$u_1$}
+                };
+        \end{tikzpicture}
+        \caption{Counterexample}
+        \label{fig:cex}
+    \end{subfigure}
+
+    \begin{subfigure}{.3\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {} edge from parent[unfixed]
+                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+                };
+        \end{tikzpicture}
+        \caption{Refined AGT}
+        \label{fig:refined}
+    \end{subfigure}%
+    \begin{subfigure}{.35\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root) {}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {} edge from parent[unfixed]}
+                    edge from parent [fixed] node [left] {$u_2$}
+                };
+        \end{tikzpicture}
+        \caption{New candidate}
+        \label{fig:candidate2}
+    \end{subfigure}%
+    \begin{subfigure}{.4\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {} edge from parent[unfixed]
+                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+                    child {node [circle,draw] {} edge from parent[fixed] node [right] {$c_2$}}
+                };
+        \end{tikzpicture}
+        \caption{Refined AGT}
+        \label{fig:refined2}
+    \end{subfigure}
+    \caption{Abstract game trees}
+\end{figure}
+
 \end{document}

### 2016-01-21 paper.tex
Beginning of the section of unbounded synthesis

@@
 \emph{runs} of the game. Previous work has applied this idea to realizability
-of bounded games \cite{nina} by forming abstract representations of the game.
-In this paper, we extend this idea to unbounded games by constructing
-approximate sets of winning states from abstract trees.
+of bounded games \cite{narodytska2014} by forming abstract representations of
+the game.  In this paper, we extend this idea to unbounded games by
+constructing approximate sets of winning states from abstract trees.
 
@@
 A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
-consists of a set of state variables, sets of controllable and uncontrollable
-label variables, a transition relationship $\delta : (X, L_c, L_u) \to X$, an
-initial state, and an error state. The \emph{controller} and \emph{environment}
-players choose controllable and uncontrollable labels respectively and the game
-proceeds according to $\delta$.
+consists of a set of boolean state variables, sets of controllable and
+uncontrollable label variables, a transition relationship $\delta : (X, L_c,
+L_u) \to X$, an initial state, and an error state. The \emph{controller} and
+\emph{environment} players choose controllable and uncontrollable labels
+respectively and the game proceeds according to $\delta$.
 
@@
 An abstract game tree can be checked for the existence of a winning run by a
-SAT solver.  The tree must be encoded into CNF by making copies of the
-transition relation for each game step in tree. The formula must also check
-whether the error state has been reached in no branch/in all branches for the
-controller and environment respectively. The functions that produce these
-formulas are shown in Algorithm \ref{alg:treeFormula}.
+SAT solver\cite{narodytska2014}.  The tree must be encoded into CNF by making
+copies of the transition relation for each game step in tree. The formula must
+also check whether the error state has been reached in no branch/in all
+branches for the controller and environment respectively. The functions that
+produce these formulas are shown in Algorithm \ref{alg:treeFormula}.
 
@@
 
-Synthesis begins with an empty game as seen in Figure \ref{fig:emptytree}.
-Initially we are playing on behalf of the environment because it chooses the
-first move in each step. The empty game is passed to the SAT solver, which
-searches for a candidate environment strategy.  If a candidate is found (Figure
-\ref{fig:candidate}) then it is checked for a spoiling strategy by solving for
-the controller. If no spoiling strategy exists, that means our candidate is a
-winning strategy and the algorithm terminates.  Otherwise we find a
-counterexample (Figure \ref{fig:cex}), which is used to refine the empty game
-tree to include the first move from the controller's spoiling strategy (Figure
-\ref{fig:refined}.  The algorithm continues by finding a new candidate for the
-environment (Figure \ref{fig:candidate2}), and a new counterexample to refine
-the game abstraction once again (Figure \ref{fig:refined2}).
+The bounded synthesis algorithm described in \cite{narodytska2014} begins with
+an empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing
+on behalf of the environment because it chooses the first move in each step.
+The empty game is passed to the SAT solver, which searches for a candidate
+environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
+then it is checked for a spoiling strategy by solving for the controller.  If
+no spoiling strategy exists, that means our candidate is a winning strategy and
+the algorithm terminates.  Otherwise we find a counterexample (Figure
+\ref{fig:cex}), which is used to refine the empty game tree to include the
+first move from the controller's spoiling strategy (Figure \ref{fig:refined}.
+The algorithm continues by finding a new candidate for the environment (Figure
+\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
+once again (Figure \ref{fig:refined2}).
 
@@
 
+\section{Unbounded Synthesis}
+
+Bounded synthesis can be used to prove the existence of a winning strategy for
+the environment by providing a witness. For the controller, the strongest claim
+that can be made is that the strategy is winning as long as the game does not
+extend longer than the bound.
+
+In model checking the maximum bound is decided based on the states of the game
+\cite{biere1999}. The na\"ive approach is to use size of the state space as the
+bound ($2^{|X|}$). With a bound of this size all states may be explored. One
+optimisation is to use the diameter of the game, which is the smallest number
+$d$ such that for any state $x$ there is a path of length $\leq d$ to all other
+reachable states. However, for large games these bounds are intractable.
+
+When performing model checking or synthesis with BDDs \cite{burch1990} the set
+of states that are winning for the environment is iteratively constructed by
+computing the states from which the environment can force the game into the
+previous winning set. Eventually this process reaches a fixed point and the
+total set of environment winning states is known.
+
+A similar concept can be applied to the bounded synthesis algorithm to
+iteratively increase the bound of the game and terminate when a fixed point is
+reached. When a strategy is found to be winning on an abstract game tree, we
+record as winning the states from which the opponent could find no
+counterexample. To find these states we use interpolation of subformulas of the
+game tree.
+
+\subsection{Learning States with Interpolants}
+
+Given two formulas $A$ and $B$ such that $A \land B$ is unsatisfiable, it is
+possible to construct and interpolant $\mathcal{I}$ such that $A \to
+\mathcal{I}$, $B \land \mathcal{I}$ is unsatisfiable, and $\mathcal{I}$ refers
+only to the intersection of variables in $A$ and $B$. An interpolant can be
+constructed efficiently from a resolution proof of the unsatisfiability of $A
+\land B$ \cite{pudlak1997}.
+
+\bibliographystyle{splncs03}
+\bibliography{paper}
+
 \end{document}

### 2016-01-23 paper.tex
Abstract + some unbounded + some bounded alg + editing

@@
 \usetikzlibrary{trees}
+
+\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
+\algnewcommand{\EndIIf}{}
 
@@
 \begin{abstract}
-    This is the abstract
+
+    Reactive synthesis techniques based on image computations have been shown
+    to work well in many cases but suffer from state explosion in others.
+    A different approach applies SAT solvers in a counterexample guided framework but
+    is limited to bounded synthesis.  We present an extension of this technique
+    to unbounded synthesis by employing interpolation.  Experimental results
+    show this to be a promising alternative that solves some previously
+    intractable instances.
+
 \end{abstract}
@@
 
-A \emph{safety game}, $G = \langle X, L_c, L_u, \delta, I, E, \rangle$,
-consists of a set of boolean state variables, sets of controllable and
-uncontrollable label variables, a transition relationship $\delta : (X, L_c,
-L_u) \to X$, an initial state, and an error state. The \emph{controller} and
+A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E, \rangle$,
+consists of a set of boolean state variables, sets of uncontrollable and
+controllable label variables, a transition relationship $\delta : (X, L_u, L_c)
+\to X$, an initial state, and an error state. The \emph{controller} and
 \emph{environment} players choose controllable and uncontrollable labels
@@
 
-An \emph{run} of a game $(x_0, c_0, u_0), (x_1, c_1, u_1) \dots (x_n, c_n,
-u_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
-\leftarrow \delta(x_k, c_k, u_k)$. A run is winning for the controller if $x_0
+An \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
+c_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
+\leftarrow \delta(x_k, u_k, c_k)$. A run is winning for the controller if $x_0
 = I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
 all runs are restricted to length $n$, whereas unbounded games consider runs of
-infinite length.
+infinite length. Since we consider only deterministic games a run is equivalent
+to a list of assignments to $L_c$ and $L_u$.
 
-A \emph{controller strategy} $\pi^c : X \to L_c$ is a mapping of states to
-controllable labels. A controller strategy is winning in a bounded game of rank
-$n$ if all runs $(x_0, \pi^c(x_0), u_0), (x_1, \pi^c(x_1), u_1) \dots (x_n,
-\pi^c(x_n), u_n)$ are winning. Bounded \emph{realizability} is the problem of
-determining the existence of such a strategy for a bounded game.
+A \emph{controller strategy} $\pi^c : (X, L_u) \to L_c$ is a mapping of states
+and uncontrollable inputs to controllable labels. A controller strategy is
+winning in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)),
+(x_1, u_1, \pi^c(x_1, u_1)) \dots (x_n, u_n, \pi^c(x_n, u_n))$ are winning.
+Bounded \emph{realizability} is the problem of determining the existence of
+such a strategy for a bounded game.
 
-An \emph{environment strategy} $\pi^e : (X, L_c) \to L_u$ is a mapping of
-states and controllable labels to uncontrollable labels. A bounded run is
-winning for the environment if $x_0 = I \land \exists i \in \{1..n\} (x_i = E)$
-and an environment strategy is winning for a bounded game if there exists a run
-$(x_0, c_1, \pi^e(x_1, c_1)), (x_1, c_1, \pi^e(x_1, c_1)) \dots (x_n, c_n,
-\pi^e(x_n, c_n))$ that wins for the environment. Safety games are zero sum,
-therefore the existence of a controller strategy implies the nonexistence of an
-environment strategy and vice versa.
+An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
+uncontrollable labels. A bounded run is winning for the environment if $x_0
+= I \land \exists i \in \{1..n\} (x_i = E)$ and an environment strategy is
+winning for a bounded game if there exists a run $(x_0, \pi^e(x_1), c_1), (x_1,
+\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ that wins for the environment.
+Safety games are zero sum, therefore the existence of a controller strategy
+implies the nonexistence of an environment strategy and vice versa.
 
@@
 
-A set of runs can be symbolically represented by a tree of labels. Each edge of
-the tree is either a \emph{fixed} valuation of controllable or uncontrollable
-variables, or it is \emph{unfixed} denoting any values are possible. Also, the
-tree must alternate between controllable and uncontrollable edges to reflect
-the alternating turns of the game. The entire set of runs of a bounded game of
-rank $n$ is symbolically represented by a tree of depth $2n$ and width 1
-populated by unfixed edges. Reducing the set of runs in a game forms an
-abstract game, which can be represented symbolically by an \emph{abstract game
-tree}.
+A set of runs can be represented symbolically by a tree of valuations to
+controllable and uncontrollable label variables where valuations are allowed to
+be \emph{unfixed}. An unfixed value symbolically represents all possible
+valuations. Formally, we define the tree as a set of lists of label valuations
+and a node in the tree may be identified by the list made by following the path
+from the root.
+
+The entire set of runs of a bounded game of rank $n$ is
+symbolically represented by a tree of depth $2n$ and width 1 populated by
+unfixed edges.  Reducing the set of runs in a game forms an abstract game,
+which can be represented symbolically by an \emph{abstract game tree}.
 
@@
 
+\begin{algorithm}
+    \begin{algorithmic}
+        \Function{solveAbstract}{$\langle s, r \rangle , gt$}
+        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, gt$}
+        \IIf{$r = n - 1$} \Return $cand$ \EndIIf 
+        \State $gt' \gets gt$
+        \Loop
+            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf
+            \State $cex \gets $ \Call{verify}{$\langle s, r \rangle, gt, cand$}
+            \IIf{$cex = NULL$} \Return $cand$ \EndIIf
+            \State $gt' \gets gt' \cup cex$
+%%%            \State $gt' \gets \Call{refine}{gt', cex}$
+            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, gt'$}
+        \EndLoop
+        \EndFunction
+    \end{algorithmic}
+
+%%%    \begin{algorithmic}
+%%%        \Function{refine}{$gt, \langle leaf, spoiling \rangle$}
+%%%            \State \Return $gt \cup \{leaf, spoiling\}$
+%%%        \EndFunction
+%%%    \end{algorithmic}
+
+    \begin{algorithmic}
+        \Function{findCandidate}{$\langle s, r \rangle, T$}
+            \State $\hat{T} \gets \Call{gtExtend}{T}$
+            \State $fml \gets $ \Call{treeFormula}{$\hat{T}$}
+            \State $sol \gets $ \Call{SAT}{$s(X_T) \land fml$}
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \Function{verify}{$\langle r, s\rangle, gt, cand$}
+            \For{$l \in leaves(gt)$}
+                \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle r, s\rangle, gt, l$}
+                \State $spoiling \gets $ \Call{solveAbstract}{$\langle r', s' \rangle, \emptyset$}
+            \EndFor
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
 \section{Unbounded Synthesis}
@@
 
+\begin{figure}
+    \centering
+    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+        \node [] (root){}
+            child {node [circle,draw,label=left:$x_1$] {}
+                child {node [circle,draw] {}
+                    child {node [circle,draw,label=left:$x_0^0$] {} edge from parent[unfixed]}
+                    edge from parent [fixed] node [left] {$u_1$}
+                }
+                child {node [circle,draw] {} edge from parent[unfixed]
+                    child {node [circle,draw,label=right:$x_0^1$] {} edge from parent[unfixed]}
+                    edge from parent [fixed] node [right] {$u_2$}
+                }
+                edge from parent [unfixed]
+            };
+    \end{tikzpicture}
+%%%    \begin{tikzpicture}[overlay]
+%%%        \draw [rotate = -20] (-1.9,0.35) ellipse (0.7 and 1.3);
+%%%    \end{tikzpicture}
+    \caption{A controller-losing game tree}
+    \label{fig:interpolatetree}
+\end{figure}
+
+Consider the snippet of a game tree in Figure \ref{fig:interpolatetree}. The
+tree is losing for the controller, the node labelled $x_1$ is at rank 1, and
+$x_0^0$ and $x_0^1$ are at rank 0. Since the tree is controller-losing we know
+that at least one run represented by the tree contains the error state.  As
+a result, $\textsc{treeFormula}(gt)$ is unsatisfiable. If we take the step
+$x_1$ to $x_0^0$ and cut it from the rest of the tree then
+$\textsc{treeFormula}(step) \land \textsc{treeFormula}(parent)$ must still be
+unsatisfiable.
+
+We can construct an interpolant with $A = \textsc{treeFormula}(parent)$ and $B
+= \textsc{treeFormula}(step)$. The only variables shared between $A$ and $B$
+are the state variables at $x_1$. We know that $B \land \mathcal{I}$ is
+unsatisfiable, therefore all states in $\mathcal{I}$ must lose to the
+uncontrollable label $u_1$. We also know that $A \to \mathcal{I}$, thus
+$\mathcal{I}$ contains all states reachable by the parent tree (on runs that
+avoid the error state.)
+
+\begin{algorithm}
+    \caption{Amended tree formulas for Controller and Environment respectively}
+    \label{alg:treeFormula}
+    \begin{algorithmic}
+        \Function{treeFormula}{gt}
+        \If{$\Call{rank}{gt} == 0$}
+        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}}$ }
+        \Else
+        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \Function{$\overline{treeFormula}$}{gt}
+        \If{$\Call{rank}{gt} == 0$}
+        \State \Return{\Call{E}{$x^{gt}$}}
+        \Else
+        \State \Return{ $\Call{$L^e[rank(gt)]$}{x^{gt}} \land$ \\
+            $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+Now we can consider the step $x_1$ to $x_0^1$, and the parent - now without
+either $x_1 \to x_0^0$ or $x_1 \to x_0^1$. The formula
+$(\textsc{treeFormula}(parent) \land \mathcal{I}) \land
+\textsc{treeFormula}(step)$ must be unsatisfiable. $\mathcal{I}$ contains all
+states that lose to $u_1$ so any other state reachable at $x_1$ must lose to
+$u_2$. Therefore we can compute another interpolant that contains states that
+lose to $u_2$.
+
+This is the foundation for a recursive algorithm that consumes an entire tree
+by removing a single step on each iteration. All learned states from which the
+controller must lose are recorded in a set $L^c$.  This algorithm can also be
+performed on environment-losing trees with the caveat that any state learnt at
+a node of rank $n$ is only known to lose at ranks less than or equal to $n$.
+
+\subsection{Proof of Termination}
+
+\subsection{Algorithm}
+
+\section{Evaluation}
+
+\section{Related Work}
+
+\section{Conclusion}
+
 \bibliographystyle{splncs03}

### 2016-01-24 paper.tex
Cleaned up and commented bounded algorithm
Work on theory sections, copied over Leonid's proof, needs editing still

@@
 \usepackage{algpseudocode}
+\usepackage{fixltx2e}
 \usepackage{subcaption}
@@
 \algnewcommand{\EndIIf}{}
+\MakeRobust{\Call}
+\newcommand{\True}{\texttt{true}}
+\newcommand{\False}{\texttt{false}}
+\newcommand{\II}{\mathcal{I}}
+\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
 
@@
 
-\title{Untitled CAV Paper}
+\title{A SAT-Based Counterexample Guided Method for Unbounded Synthesis}
 
 \author{Alexander Legg\inst{1} 
-    \and Leonid Ryzhyk\inst{2}
-    \and Nina Narodytska\inst{2}}
+    \and Nina Narodytska\inst{2}
+    \and Leonid Ryzhyk\inst{2}}
 
@@
     \email{alexander.legg@nicta.com.au}
-    \and Samsung Research}
+    \and Samsung Research America}
 
@@
 
-    Reactive synthesis techniques based on image computations have been shown
-    to work well in many cases but suffer from state explosion in others.
-    A different approach applies SAT solvers in a counterexample guided framework but
-    is limited to bounded synthesis.  We present an extension of this technique
-    to unbounded synthesis by employing interpolation.  Experimental results
-    show this to be a promising alternative that solves some previously
-    intractable instances.
+    Reactive synthesis techniques based on constructing the winning region of
+    the system have been shown to work well in many cases but suffer from state
+    explosion in others.  A different approach, proposed recently, applies SAT
+    solvers in a counterexample guided framework to solve the synthesis
+    problem.  However, this method is limited to synthesising systems that
+    execute for a bounded number of steps and is incomplete for synthesis with
+    unbounded safety and reachability objectives.  We present an extension of
+    this technique to unbounded synthesis.  Our method applies Craig
+    interpolation to abstract game trees produced by counterexample-guided
+    search in order to construct a monotonic sequence of may-losing regions.
+    Experimental results based on SYNTCOMP 2015 competition benchmarks show
+    this to be a promising alternative that solves some previously intractable
+    instances.
 
@@
     \label{alg:treeFormula}
-    \begin{algorithmic}
+    \begin{algorithmic}[1]
         \Function{treeFormula}{gt}
@@
         \EndFunction
+        \algstore{tf1}
     \end{algorithmic}
 
-    \begin{algorithmic}
-        \Function{$\overline{treeFormula}$}{gt}
+    \begin{algorithmic}[1]
+        \algrestore{tf1}
+        \Function{\textoverline{treeFormula}}{gt}
         \If{$\Call{rank}{gt} == 0$}
@@
         \Else
-        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n})$ }
+        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$ }
         \EndIf
@@
 
-The bounded synthesis algorithm described in \cite{narodytska2014} begins with
-an empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing
-on behalf of the environment because it chooses the first move in each step.
-The empty game is passed to the SAT solver, which searches for a candidate
-environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
-then it is checked for a spoiling strategy by solving for the controller.  If
-no spoiling strategy exists, that means our candidate is a winning strategy and
-the algorithm terminates.  Otherwise we find a counterexample (Figure
-\ref{fig:cex}), which is used to refine the empty game tree to include the
-first move from the controller's spoiling strategy (Figure \ref{fig:refined}.
-The algorithm continues by finding a new candidate for the environment (Figure
+This technique was proposed by Narodytska et al \cite{narodytska2014}. In
+Section \ref{sect:unbounded} we propose an extension to synthesis of unbounded
+games.
+
+The bounded synthesis algorithm begins with an empty game as seen in Figure
+\ref{fig:emptytree}.  Initially we are playing on behalf of the environment
+because it chooses the first move in each step.  The empty game is passed to
+the SAT solver, which searches for a candidate environment strategy.  If a
+candidate is found (Figure \ref{fig:candidate}) then it is checked for a
+spoiling strategy by solving for the controller.  If no spoiling strategy
+exists, that means our candidate is a winning strategy and the algorithm
+terminates.  Otherwise we find a counterexample (Figure \ref{fig:cex}), which
+is used to refine the empty game tree to include the first move from the
+controller's spoiling strategy (Figure \ref{fig:refined}.  The algorithm
+continues by finding a new candidate for the environment (Figure
 \ref{fig:candidate2}), and a new counterexample to refine the game abstraction
-once again (Figure \ref{fig:refined2}).
+once again (Figure \ref{fig:refined2}). The full algorithm is listed in
+Algorithm \ref{alg:bounded}.
 
@@
 \begin{algorithm}
+    \begin{algorithmic}[1]
+        \Function{solveAbstract}{$\langle s, r \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
+        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \State $T' \gets T$
+        \Loop
+            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
+            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
+            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
+            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
+        \EndLoop
+        \EndFunction
+        \algstore{b1}
+    \end{algorithmic}
+
     \begin{algorithmic}
-        \Function{solveAbstract}{$\langle s, r \rangle , gt$}
-        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, gt$}
-        \IIf{$r = n - 1$} \Return $cand$ \EndIIf 
-        \State $gt' \gets gt$
-        \Loop
-            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf
-            \State $cex \gets $ \Call{verify}{$\langle s, r \rangle, gt, cand$}
-            \IIf{$cex = NULL$} \Return $cand$ \EndIIf
-            \State $gt' \gets gt' \cup cex$
-%%%            \State $gt' \gets \Call{refine}{gt', cex}$
-            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, gt'$}
-        \EndLoop
+        \algrestore{b1}
+        \Function{findCandidate}{$\langle s, r \rangle, T$}
+            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
+            \IIf{$sol = \texttt{unsat}$} 
+                \Return $\emptyset$ \Comment{No candidate exists}
+            \EndIIf
+            \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+        \EndFunction
+        \algstore{b2}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{b2}
+        \Function{verify}{$\langle s, r\rangle, gt, cand$}
+            \For{$l \in leaves(gt)$}
+            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
+                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
+            \EndFor
+            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
         \EndFunction
@@
 
-%%%    \begin{algorithmic}
-%%%        \Function{refine}{$gt, \langle leaf, spoiling \rangle$}
-%%%            \State \Return $gt \cup \{leaf, spoiling\}$
-%%%        \EndFunction
-%%%    \end{algorithmic}
-
-    \begin{algorithmic}
-        \Function{findCandidate}{$\langle s, r \rangle, T$}
-            \State $\hat{T} \gets \Call{gtExtend}{T}$
-            \State $fml \gets $ \Call{treeFormula}{$\hat{T}$}
-            \State $sol \gets $ \Call{SAT}{$s(X_T) \land fml$}
-        \EndFunction
-    \end{algorithmic}
-
-    \begin{algorithmic}
-        \Function{verify}{$\langle r, s\rangle, gt, cand$}
-            \For{$l \in leaves(gt)$}
-                \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle r, s\rangle, gt, l$}
-                \State $spoiling \gets $ \Call{solveAbstract}{$\langle r', s' \rangle, \emptyset$}
-            \EndFor
-        \EndFunction
-    \end{algorithmic}
+    \caption{Bounded Synthesis}
+    \label{alg:bounded}
 \end{algorithm}
@@
 \section{Unbounded Synthesis}
+
+\label{sect:unbounded}
 
@@
 \begin{algorithm}
+    \begin{algorithmic}[1]
+        \Require $\sigma(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
+        \Require \emph{Must-invariant} holds
+        \Ensure \emph{Must-invariant} holds
+        \Ensure $\sigma(X_T) \land B^M \equiv \bot$
+        \Function{learn}{$\sigma, T$}
+            \IIf{$T = \emptyset$}
+                \Return
+            \EndIIf
+            \State $n \gets $ non-leaf node with max rank
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
+            \State $B^M \gets B^M \lor \mathcal{I}$
+            \State \Call{learn}{$\sigma, T_1$}
+        \EndFunction
+        \algstore{learn}
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \algrestore{learn}
+        \Require $\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
+        \Require \emph{May-invariant} holds
+        \Ensure \emph{May-invariant} holds
+        \Ensure $\sigma(X_T) \land B^m[$\Call{rank}{$T$}$] \equiv \bot$
+        \Function{\textoverline{learn}}{$\sigma, T$}
+            \IIf{$T = \emptyset$}
+                \Return
+            \EndIIf
+            \State $n \gets $ non-leaf node with max rank
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
+            \For{$i = 1$ to \Call{rank}{$n$}}
+                \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
+            \EndFor
+            \State \Call{\textoverline{learn}}{$\sigma, T_1$}
+        \EndFunction
+    \end{algorithmic}
+    \caption{Learning algorithms}
+    \label{alg:learn}
+\end{algorithm}
+
+\begin{algorithm}
     \caption{Amended tree formulas for Controller and Environment respectively}
     \label{alg:treeFormula}
-    \begin{algorithmic}
+    \begin{algorithmic}[1]
         \Function{treeFormula}{gt}
         \If{$\Call{rank}{gt} == 0$}
-        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}}$ }
+        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}}$ }
         \Else
-        \State \Return{ $\lnot \Call{$L^c$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
         \EndIf
         \EndFunction
+        \algstore{tf2}
     \end{algorithmic}
@@
     \begin{algorithmic}
-        \Function{$\overline{treeFormula}$}{gt}
+        \algrestore{tf2}
+        \Function{\textoverline{treeFormula}}{gt}
         \If{$\Call{rank}{gt} == 0$}
@@
         \Else
-        \State \Return{ $\Call{$L^e[rank(gt)]$}{x^{gt}} \land$ \\
+        \State \Return{ $\Call{$B^m[rank(gt)]$}{x^{gt}} \land$ \\
             $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
@@
 This is the foundation for a recursive algorithm that consumes an entire tree
-by removing a single step on each iteration. All learned states from which the
-controller must lose are recorded in a set $L^c$.  This algorithm can also be
-performed on environment-losing trees with the caveat that any state learnt at
-a node of rank $n$ is only known to lose at ranks less than or equal to $n$.
+by removing a single step on each iteration (Algorithm \ref{alg:learn}). All
+learned states from which the controller \emph{must} lose are recorded in a set
+of \emph{bad} states $B^M$.  This algorithm can also be performed on
+environment-losing trees with the caveat that any state learnt at a node of
+rank $n$ is only known to lose at ranks less than or equal to $n$. We negate
+the interpolants we learn this way and record them in a mapping of ranks to
+sets of states $B^m$ which \emph{may} be bad states for the controller.
+
+\begin{algorithm}
+    \begin{algorithmic}[1]
+        \Function{solveUnbounded}{$T$}
+            \State $B^M \gets E$
+            \State $B^m[0] \gets E$
+            \For{$k = 1 \dots$}
+                \IIf{\Call{SAT}{$I \land B^M$}}
+                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
+                \EndIIf
+                \IIf{$\exists i \  B^m[i] \equiv B^m[i+1]$}
+                    \Return \texttt{realisable} \Comment{Reached fixed point}
+                \EndIIf
+                \State $B^m[k] \gets \top$
+                \State \Call{solveAbstract}{$\langle I, k \rangle, \emptyset$}
+            \EndFor
+        \EndFunction
+        \algstore{u1}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{u1}
+        \Function{solveAbstract}{$\langle s, r \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
+        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \State $T' \gets T$
+        \Loop
+            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
+            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
+            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
+            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
+        \EndLoop
+        \EndFunction
+        \algstore{u2}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{u2}
+        \Function{findCandidate}{$\langle s, r \rangle, T$}
+            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
+            \If{$sol = \texttt{unsat}$} 
+                \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
+                \State \Call{learn}{$\sigma, T$}
+                \State \Return $\emptyset$ \Comment{No candidate exists}
+            \Else
+                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+            \EndIf
+        \EndFunction
+        \algstore{u3}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{u3}
+        \Function{verify}{$\langle s, r\rangle, gt, cand$}
+            \For{$l \in leaves(gt)$}
+            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
+                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
+            \EndFor
+            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
+        \EndFunction
+    \end{algorithmic}
+
+    \caption{Unbounded Synthesis}
+    \label{alg:unbounded}
+\end{algorithm}
+
+\subsection{Correctness}
+
+The unbounded synthesis algorithm (Algorimthm \ref{alg:unbounded}) effectively
+consists of two communicating solvers: the counterexample-guided bounded
+reachability solver and the unbounded solver based on incremental induction.
+Fortunately, correctness of both solvers can be established largely
+independently, which simplifies the proof of correctness of the overall
+algorithm.
+
+We define two global invariants of the algorithm.  The \emph{may-invariant}
+states that sets $B^m[i]$ grows monotonically with $i$ and that each $B^m[i+1]$
+overapproximates the uncontrollable predecessor of $f^m[i]$: $$\forall
+i<k.~B^m[i] \subseteq B^m[i+1], Upre(B^m[i]) \subseteq B^m[i+1].$$
+
+The \emph{must-invariant} guarantees that the must-losing set $B^M$ is an
+underapproximation of the actual losing set $B$: $$B^M \subseteq B$$.
+
+The sets $B^m$ and $B^M$ are only modified by the inductive solver, implemented
+by \textsc{learn} and \textoverline{\textsc{learn}} functions.  Below we prove that these
+functions indeed maintain the invariants.
+
+\subsection{Proof of \textsc{learn}}
+
+We prove that postconditions of \textsc{learn} are satisfied assuming that its
+preconditions hold.
+
+Line~(4) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has
+depth 1.  Consider formulas $f_1=\sigma(X_T) \land \textsc{treeFormula}(T_1)$
+and $F_2 = \textsc{treeFormula}(T_2)$.  These formulas only share variables
+$X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
+any solution of $F_1 \land F_2$ also satisfies $\sigma(X_T) \land
+\textsc{treeFormula}(T)$, which is unsatisfiable (precondition (b)).  Hence the
+interpolation operation is defined for $F_1$ and $F_2$.
+
+Intuitively, the interpolant computed in line~(5) overapproximates the set of 
+states reachable from $\sigma$ by following the tree from the root node to $n$,
+and underapproximates the set of states from which the controller loses against 
+tree $T_2$.
+
+Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is of depth
+1, this means that the controller cannot force the
+game into $W^m[|rank|(n-1)]$ when playing against counterexample moves in $T_2$.  Hence, 
+$\II \cap Cpre(W^m[|rank|(n)]) = \emptyset$.  Furthermore, since the
+may-invariant holds (assumption (c)), $\II \cap Cpre(i) = \emptyset$, for
+all $i<|rank|(n)$.  Hence, removing $\II$ from all $W^m[i], i\leq |rank|(n)$
+in line~(6) preserves the may-invariant, thus satisfying the first post-condition.
+
+Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any assignment to 
+$X_n$ that satisfies $\sigma(X_T) \land |treeFormula|(T_1)$ also satisfies $\II$.  Hence, removing $\II$ 
+from $W_m[|rank|(n)]$ makes $\sigma(X_T) \land |treeFormula|(T_1)$ unsatisfiable, 
+and hence all preconditions of the recursive invocation of \textsf{learn} in line~(7) 
+are satisfied.
+
+At the second las recursive call to \textsf{learn}, tree $T_1$ is empty, $n$ is the root node,
+$|treeFormula|(T_1) \equiv W^m[|rank|(T)](X^T)$; hence 
+$\sigma(X_T) \land |treeFormula|(T_1) \equiv \sigma(X_T) \land W^m[|rank|(T)](X^T) \equiv \bot$.
+Thus the second postcondition of \textsf{learn} holds.
 
@@
 
-\subsection{Algorithm}
+We must prove that $|checkRank|$ terminates and that upon 
+termination its postcondition holds, i.e., state $s$ is added to $W^M$ if 
+its rank is no greater than $k$ or is removed from $W^m[k]$ otherwise.  
+Termination follows from completeness of couterexample-guided search, which 
+terminates after enumerating all possible opponent moves in the worst case.
+
+Assume that $|rank|(s) > k$.  This means that at some point the algorithm
+discovers a counterexample tree of rank $k$ that is losing for the controller 
+from state $s$.  The algorithm then invokes the $|learn|$ method, which 
+eliminates $s$ from $W^m[k]$.  Alternatively, if $|rank|(s) \leq k$, a 
+counterexample tree losing for the environment will be discovered and 
+subsequent call to $\overline{|learn|}$ adds $s$ to $W^M$.
+
+\section{Optimisations}
+
+\subsection{Generalising the initial state}
+
+\subsection{Choosing sensible opponent moves}
 

### 2016-01-25 paper.tex
Sections 2 & 3 are hopefully complete & consistent now

@@
     \begin{algorithmic}[1]
-        \Function{solveAbstract}{$\langle s, r \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
-        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \Function{solveAbstract}{$\langle s, k \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$\langle s, k \rangle, T$} \Comment{Look for a candidate}
+        \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
         \State $T' \gets T$
@@
             \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
-            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, k \rangle, T, cand$} \Comment{Verify candidate}
             \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
             \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
+            \State $cand \gets $ \Call{solveAbstract}{$\langle s, k \rangle, T'$} \Comment{Solve refined AGT}
         \EndLoop
@@
         \algrestore{b1}
-        \Function{findCandidate}{$\langle s, r \rangle, T$}
+        \Function{findCandidate}{$\langle s, k \rangle, T$}
             \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
@@
         \algrestore{b2}
-        \Function{verify}{$\langle s, r\rangle, gt, cand$}
+        \Function{verify}{$\langle s, k\rangle, gt, cand$}
             \For{$l \in leaves(gt)$}
-            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
+                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                 \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
@@
         \Ensure \emph{Must-invariant} holds
-        \Ensure $\sigma(X_T) \land B^M \equiv \bot$
+        \Ensure $\sigma(X_T) \land B^M$
         \Function{learn}{$\sigma, T$}
@@
                 \State $B^m[k] \gets \top$
-                \State \Call{solveAbstract}{$\langle I, k \rangle, \emptyset$}
             \EndFor
@@
         \algrestore{u1}
-        \Function{solveAbstract}{$\langle s, r \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$\langle s, r \rangle, T$} \Comment{Look for a candidate}
-        \IIf{$r = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
-        \State $T' \gets T$
-        \Loop
-            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
-            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, r \rangle, T, cand$} \Comment{Verify candidate}
-            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
-            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$\langle s, r \rangle, T'$} \Comment{Solve refined AGT}
-        \EndLoop
+        \Require \emph{May} and \emph{must} invariants hold
+        \Ensure \emph{May} and \emph{must} invariants hold
+        \Ensure $\lnot s \in B^m[k]$ if there is a winning controller strategy of length $k$ starting at $s$
+        \Ensure $s \in B^M$ if there is a winning environment strategy of length $k$ starting at $s$
+%%%        $\exists u_{k..1} \forall c_{k..1} \  s(x_k) \land (E_k \lor (\delta(x_k, u_k, c_k) \land E(x_{k-1}) \lor ... \land E(x_1) \lor (\delta(x_1, u_1, c_1) \land E(x_0))...)$
+%%%            $\forall u_{k..1} \exists c_{k..1} \  s(x_k) \land \lnot E_k \land \delta(x_k, u_k, c_k) \land \lnot E(x_{k-1}) \land ... \land \delta(x_1, u_1, c_1) \land \lnot E(x_0)$
+        \Function{checkRank}{$s, k$}
+            \State \Return \Call{solveAbstract}{$\langle s, k \rangle, \emptyset$}
         \EndFunction
@@
         \algrestore{u2}
-        \Function{findCandidate}{$\langle s, r \rangle, T$}
+        \Function{solveAbstract}{$\langle s, k \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$\langle s, k \rangle, T$} \Comment{Look for a candidate}
+        \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \State $T' \gets T$
+        \Loop
+            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, k \rangle, T, cand$} \Comment{Verify candidate}
+            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
+            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
+            \State $cand \gets $ \Call{solveAbstract}{$\langle s, k \rangle, T'$} \Comment{Solve refined AGT}
+        \EndLoop
+        \EndFunction
+        \algstore{u3}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{u3}
+        \Function{findCandidate}{$\langle s, k \rangle, T$}
             \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
@@
         \EndFunction
-        \algstore{u3}
+        \algstore{u4}
     \end{algorithmic}
@@
     \begin{algorithmic}
-        \algrestore{u3}
-        \Function{verify}{$\langle s, r\rangle, gt, cand$}
+        \algrestore{u4}
+        \Function{verify}{$\langle s, k\rangle, gt, cand$}
             \For{$l \in leaves(gt)$}
-            \State $\langle r', s'\rangle \gets $ \Call{outcome}{$\langle s, r\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle r', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
+                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                 \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
@@
 
-The unbounded synthesis algorithm (Algorimthm \ref{alg:unbounded}) effectively
+The unbounded synthesis algorithm (Algorithm \ref{alg:unbounded}) effectively
 consists of two communicating solvers: the counterexample-guided bounded
@@
 
-\subsection{Proof of \textsc{learn}}
+\subsection{Proof of \textsc{\textoverline{learn}}}
 
-We prove that postconditions of \textsc{learn} are satisfied assuming that its
-preconditions hold.
+We prove that postconditions of \textsc{\textoverline{learn}} are satisfied
+assuming that its preconditions hold.
 
-Line~(4) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has
-depth 1.  Consider formulas $f_1=\sigma(X_T) \land \textsc{treeFormula}(T_1)$
-and $F_2 = \textsc{treeFormula}(T_2)$.  These formulas only share variables
+Line~(11) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
+1.  Consider formulas $f_1=\sigma(X_T) \land
+\textsc{\textoverline{treeFormula}}(T_1)$ and $F_2 =
+\textsc{\textoverline{treeFormula}}(T_2)$.  These formulas only share variables
 $X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
@@
 
-Intuitively, the interpolant computed in line~(5) overapproximates the set of 
+Intuitively, the interpolant computed in line~(13) overapproximates the set of
 states reachable from $\sigma$ by following the tree from the root node to $n$,
-and underapproximates the set of states from which the controller loses against 
-tree $T_2$.
+and underapproximates the set of states from which the environment loses
+against tree $T_2$.
 
-Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is of depth
-1, this means that the controller cannot force the
-game into $W^m[|rank|(n-1)]$ when playing against counterexample moves in $T_2$.  Hence, 
-$\II \cap Cpre(W^m[|rank|(n)]) = \emptyset$.  Furthermore, since the
-may-invariant holds (assumption (c)), $\II \cap Cpre(i) = \emptyset$, for
-all $i<|rank|(n)$.  Hence, removing $\II$ from all $W^m[i], i\leq |rank|(n)$
-in line~(6) preserves the may-invariant, thus satisfying the first post-condition.
+Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is
+of depth 1, this means that the environment cannot force the game into
+$B^m[\textsc{rank}(n)]$ playing against the counterexample moves in $T_2$.
+Hence, $\II \cap Upre(B^m[\textsc{rank}(n)]) = \emptyset$.  Furthermore, since
+the may-invariant holds, $\II \cap Upre(i) = \emptyset$, for all $i <
+\textsc{rank}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
+\textsc{rank}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
+first post-condition.
 
-Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any assignment to 
-$X_n$ that satisfies $\sigma(X_T) \land |treeFormula|(T_1)$ also satisfies $\II$.  Hence, removing $\II$ 
-from $W_m[|rank|(n)]$ makes $\sigma(X_T) \land |treeFormula|(T_1)$ unsatisfiable, 
-and hence all preconditions of the recursive invocation of \textsf{learn} in line~(7) 
-are satisfied.
+Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any
+assignment to $X_n$ that satisfies $\sigma(X_T) \land |treeFormula|(T_1)$ also
+satisfies $\II$.  Hence, removing $\II$ from $B^m[\textsc{rank}(n)]$ makes
+$\sigma(X_T) \land |treeFormula|(T_1)$ unsatisfiable, and hence all
+preconditions of the recursive invocation of \textsc{\textoverline{learn}} in
+line~(17) are satisfied.
 
-At the second las recursive call to \textsf{learn}, tree $T_1$ is empty, $n$ is the root node,
-$|treeFormula|(T_1) \equiv W^m[|rank|(T)](X^T)$; hence 
-$\sigma(X_T) \land |treeFormula|(T_1) \equiv \sigma(X_T) \land W^m[|rank|(T)](X^T) \equiv \bot$.
-Thus the second postcondition of \textsf{learn} holds.
+At the second last recursive call to \textsc{\textoverline{learn}}, tree $T_1$
+is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
+\equiv B^m[\textsc{rank}(T)](X^T)$; hence $\sigma(X_T) \land
+\textsc{\textoverline{treeFormula}}(T_1) \equiv \sigma(X_T) \land
+B^m[\textsc{rank}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
+\textsc{\textoverline{learn}} holds.
 
@@
 
-We must prove that $|checkRank|$ terminates and that upon 
-termination its postcondition holds, i.e., state $s$ is added to $W^M$ if 
-its rank is no greater than $k$ or is removed from $W^m[k]$ otherwise.  
-Termination follows from completeness of couterexample-guided search, which 
-terminates after enumerating all possible opponent moves in the worst case.
+We must prove that \textsc{checkRank} terminates and that upon termination its
+postcondition holds.
 
-Assume that $|rank|(s) > k$.  This means that at some point the algorithm
-discovers a counterexample tree of rank $k$ that is losing for the controller 
-from state $s$.  The algorithm then invokes the $|learn|$ method, which 
-eliminates $s$ from $W^m[k]$.  Alternatively, if $|rank|(s) \leq k$, a 
-counterexample tree losing for the environment will be discovered and 
-subsequent call to $\overline{|learn|}$ adds $s$ to $W^M$.
+We must prove that $\textsc{checkRank}$ terminates and that upon termination
+its postcondition holds, i.e., state $s$ is removed from $B^m[k]$ if there is a
+winning controller strategy on the bounded safety game of rank $k$ or it is
+added to $B^M$ otherwise. Termination follows from completeness of
+counterexample-guided search, which terminates after enumerating all possible
+opponent moves in the worst case.
+
+Assume that there is a winning strategy for the controller at rank $k$. This
+means that at some point the algorithm discovers a counterexample tree of rank
+$k$ for which the environment cannot force into $E$. The algorithm then invokes
+the \textsc{\textoverline{learn}} method, adds $s$ to $B^M$.  Alternatively, if
+there is a winning strategy for the environment at rank $k$ then a
+counterexample losing for the controller will be found. Subsequently
+\textsc{learn} will be called and $s$ eliminated from $B^m[k]$.
 

### 2016-01-25 paper.tex
Added init generalisation optimisation + more editing

@@
 \algnewcommand{\EndIIf}{}
+\algnewcommand{\IfElse}[3]{\algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}
+\algnewcommand{\EndIfElse}{}
 \MakeRobust{\Call}
@@
         \Function{treeFormula}{gt}
-        \If{$\Call{rank}{gt} == 0$}
+        \If{$\Call{rank}{gt} = 0$}
         \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
@@
         \Function{\textoverline{treeFormula}}{gt}
-        \If{$\Call{rank}{gt} == 0$}
+        \If{$\Call{rank}{gt} = 0$}
         \State \Return{\Call{E}{$x^{gt}$}}
@@
     \begin{algorithmic}[1]
-        \Function{solveAbstract}{$\langle s, k \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$\langle s, k \rangle, T$} \Comment{Look for a candidate}
+        \Function{solveAbstract}{$p, \langle s, k \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$p, \langle s, k \rangle, T$} \Comment{Look for a candidate}
         \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
@@
             \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
-            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$\langle s, k \rangle, T, cand$} \Comment{Verify candidate}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$p, \langle s, k \rangle, T, cand$} \Comment{Verify candidate}
             \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
             \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$\langle s, k \rangle, T'$} \Comment{Solve refined AGT}
+            \State $cand \gets $ \Call{solveAbstract}{$p, \langle s, k \rangle, T'$} \Comment{Solve refined AGT}
         \EndLoop
@@
         \algrestore{b1}
-        \Function{findCandidate}{$\langle s, k \rangle, T$}
-            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
+        \Function{findCandidate}{$p, \langle s, k \rangle, T$}
+            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
+            \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
             \IIf{$sol = \texttt{unsat}$} 
@@
             \EndIIf
-            \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+            \State \Return $\{ \langle n, m \rangle | n \in $ \Call{gtNodes}{T} $, m = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
         \EndFunction
@@
         \algrestore{b2}
-        \Function{verify}{$\langle s, k\rangle, gt, cand$}
+        \Function{verify}{$p, \langle s, k\rangle, gt, cand$}
             \For{$l \in leaves(gt)$}
             \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                 \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
@@
         \Function{treeFormula}{gt}
-        \If{$\Call{rank}{gt} == 0$}
+        \If{$\Call{rank}{gt} = 0$}
         \State \Return{ $\lnot \Call{$B^M$}{x^{gt}}$ }
@@
                 \State $B^m[k] \gets \top$
+                \State \Call{checkRank}{$I, k$}
             \EndFor
@@
         \Function{checkRank}{$s, k$}
-            \State \Return \Call{solveAbstract}{$\langle s, k \rangle, \emptyset$}
+            \State \Return \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
         \EndFunction
@@
         \algrestore{u2}
-        \Function{solveAbstract}{$\langle s, k \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$\langle s, k \rangle, T$} \Comment{Look for a candidate}
+        \Function{solveAbstract}{$p, \langle s, k \rangle , T$}
+        \State $cand \gets $ \Call{findCandidate}{$p, \langle s, k \rangle, T$} \Comment{Look for a candidate}
         \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
@@
         \algrestore{u3}
-        \Function{findCandidate}{$\langle s, k \rangle, T$}
-            \State $sol \gets $ \Call{SAT}{$s(X_T) \land$ \Call{treeFormula}{$T$} }
+        \Function{findCandidate}{$p, \langle s, k \rangle, T$}
+            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
+            \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
             \If{$sol = \texttt{unsat}$} 
@@
                 \State \Call{learn}{$\sigma, T$}
+                \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, T$}}{\Call{\textoverline{learn}}{$\sigma, T$}} \EndIfElse
                 \State \Return $\emptyset$ \Comment{No candidate exists}
@@
         \algrestore{u4}
-        \Function{verify}{$\langle s, k\rangle, gt, cand$}
+        \Function{verify}{$p, \langle s, k\rangle, gt, cand$}
             \For{$l \in leaves(gt)$}
             \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{\textoverline{solveAbstract}}{$\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
                 \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
@@
 
+In this section we present some optimisations to the unbounded synthesis algorithm.
+
 \subsection{Generalising the initial state}
+
+If more states are removed from $B^m$ on each iteration of \textsc{checkRank}
+the algorithm will converge and terminate sooner. Additionally, any states
+removed from $B^M$ will reduce the number of states to be considered by the
+controller in future iterations.
+
+As it is written, the algorithm only considers an overapproximations of states
+reachable from $I$ for learning. Assuming that the algorithm does not
+terminate, then the addition of an extra step into the game on each iteration
+has the potential to greatly increase the reachable states. Considering some of
+these states earlier than they become reachable may lead to earlier
+termination.
+
+\begin{algorithm}
+    \begin{algorithmic}
+        \Function{checkRank}{$s, k$}
+            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
+            \IIf{$r \neq \emptyset$} \Return $r$ \EndIIf
+            \State $s' \gets s$
+            \For{$x \in s$}
+                \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, \langle s' \setminus \{x\}, k \rangle, \emptyset$}
+                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf
+            \EndFor
+            \State \Return $\emptyset$
+        \EndFunction
+    \end{algorithmic}
+    \caption{Generalise $I$ optimisation}
+    \label{alg:opt1}
+\end{algorithm}
+
+The optimisation that allows this is relatively simple and is inspired by a
+common greedy heuristic for minimising $\texttt{unsat}$ cores. $I$ is a value
+assignment to each variable in $X$. If the environment does not win for
+$\langle I, k \rangle$ then we attempt to solve for a generalised version of
+$I$ by removing one assignment at a time. If the controller can win from the
+larger set of states then we continue generalising without it. In this way we
+learn more states by increasing the reachable set. In our benchmarks we have
+observed that this optimisation is beneficial on the first few iterations of
+\textsc{checkRank}.
 

### 2016-01-25 paper.tex
more editing + optimisations section

@@
 
-The entire set of runs of a bounded game of rank $n$ is
-symbolically represented by a tree of depth $2n$ and width 1 populated by
-unfixed edges.  Reducing the set of runs in a game forms an abstract game,
-which can be represented symbolically by an \emph{abstract game tree}.
+The entire set of runs of a bounded game of rank $n$ is symbolically
+represented by a tree of depth $2n$ (a step is two moves - environment then
+controller) and width 1 populated by unfixed edges.  Fixing some moves and
+thereby reducing the set of runs forms an \emph{abstract game tree} that
+symbolically represents a smaller version of the game.
 
@@
 copies of the transition relation for each game step in tree. The formula must
-also check whether the error state has been reached in no branch/in all
-branches for the controller and environment respectively. The functions that
-produce these formulas are shown in Algorithm \ref{alg:treeFormula}.
+also check whether the error state has been reached in no branch
+(\emph{respectively}: in all branches) for the controller (\emph{resp}:
+environment.) The functions that produce these formulas are shown in Algorithm
+\ref{alg:treeFormula}.
 
@@
 that if there exists a way for the game to be solved when both players
-cooperate the SAT solver will find it. If no winning run exists in an abstract
-game even when the players are cooperating then there is definitely no winning
-run when the opponent is playing adversarily.
+cooperate on unfixed labels the SAT solver will find it. If no winning run
+exists in an abstract game even when the players are cooperating then there is
+no winning run when the opponent is playing adversarily.
 
@@
 abstract game tree constructed from a partial controller strategy we are
-checking if the partial strategy is winning. If no spoiling run can be found
-then the partial strategy must always win. If a spoiling run is found then we
-then we have a counterexample that proves that the controller strategy is not
-winning. This forms the basis of a counterexample guided abstraction refinement
-framework that operates on candidate strategies.
+checking if the partial strategy is winning. If no environment-winning run is
+exists within the partial strategy then it is a winning partial strategy,
+otherwise the spoiling run is a counterexample that proves that the partial
+strategy is not winning. This forms the basis of a counterexample guided
+abstraction refinement framework that operates on candidate strategies.
 
@@
 Section \ref{sect:unbounded} we propose an extension to synthesis of unbounded
-games.
+games but the original algorithm is outlined here for completeness.
 
@@
 is used to refine the empty game tree to include the first move from the
-controller's spoiling strategy (Figure \ref{fig:refined}.  The algorithm
+controller's spoiling strategy (Figure \ref{fig:refined}).  The algorithm
 continues by finding a new candidate for the environment (Figure
 \ref{fig:candidate2}), and a new counterexample to refine the game abstraction
-once again (Figure \ref{fig:refined2}). The full algorithm is listed in
+once again (Figure \ref{fig:refined2}). The full procedure is listed in
 Algorithm \ref{alg:bounded}.
@@
 Bounded synthesis can be used to prove the existence of a winning strategy for
-the environment by providing a witness. For the controller, the strongest claim
-that can be made is that the strategy is winning as long as the game does not
-extend longer than the bound.
+the environment on the unbounded game by providing a witness. For the
+controller, the strongest claim that can be made is that the strategy is
+winning as long as the game does not extend beyond the bound.
 
-In model checking the maximum bound is decided based on the states of the game
-\cite{biere1999}. The na\"ive approach is to use size of the state space as the
-bound ($2^{|X|}$). With a bound of this size all states may be explored. One
-optimisation is to use the diameter of the game, which is the smallest number
-$d$ such that for any state $x$ there is a path of length $\leq d$ to all other
-reachable states. However, for large games these bounds are intractable.
+It is possible to set a bound such that all runs in the unbounded game will be
+considered. The na\"ive approach is to use size of the state space as the bound
+($2^{|X|}$) so that all states may be explored by the algorithm. A more nuanced
+approach is to use the diameter of the game \cite{biere1999}, which is the
+smallest number $d$ such that for any state $x$ there is a path of length $\leq
+d$ to all other reachable states. This approach is common in bounded model
+checking but for synthesis it quickly becomes intractable to consider such long
+runs.
 
@@
 Given two formulas $A$ and $B$ such that $A \land B$ is unsatisfiable, it is
-possible to construct and interpolant $\mathcal{I}$ such that $A \to
-\mathcal{I}$, $B \land \mathcal{I}$ is unsatisfiable, and $\mathcal{I}$ refers
-only to the intersection of variables in $A$ and $B$. An interpolant can be
-constructed efficiently from a resolution proof of the unsatisfiability of $A
-\land B$ \cite{pudlak1997}.
+possible to construct a Craig interpolant\cite{craig1957} $\mathcal{I}$ such
+that $A \to \mathcal{I}$, $B \land \mathcal{I}$ is unsatisfiable, and
+$\mathcal{I}$ refers only to the intersection of variables in $A$ and $B$.  An
+interpolant can be constructed efficiently from a resolution proof of the
+unsatisfiability of $A \land B$ \cite{pudlak1997}.
 
@@
 $x_1$ to $x_0^0$ and cut it from the rest of the tree then
-$\textsc{treeFormula}(step) \land \textsc{treeFormula}(parent)$ must still be
+$\textsc{treeFormula}(step) \land \textsc{treeFormula}(parent)$ must too be
 unsatisfiable.
@@
     \caption{Amended tree formulas for Controller and Environment respectively}
-    \label{alg:treeFormula}
+    \label{alg:unboundedTreeFormula}
     \begin{algorithmic}[1]
@@
 
-In this section we present some optimisations to the unbounded synthesis algorithm.
+%%%In this section two optimisations to the unbounded synthesis algorithm are
+%%%described.
 
@@
 
+One aspect of the algorithm proposed by Narodytska et al\cite{narodytska2014}
+is that it mimics real-world uses of synthesis by allowing cooperation between
+system and environment. Unfortunately, allowing the opponents to pick moves for
+one another can frequently backfire. It is common when modelling real-world
+systems to abstract over out of scope failures with transitions that
+immediately determine the outcome of the game. For example, in a network driver
+a request to send a packet may fail due to a disconnected wire, which might be
+modelled as an environment controlled transition to a system-winning state.
+
+This slows down the synthesis process as all erroneous transition are
+accumulated into the abstract game tree. To mitigate this effect we have a
+heuristic to guess non-erroneous labels as \emph{default moves} that are used
+instead of allowing the opponent to choose. This does not effect the
+correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
+is \texttt{unsat} with default moves then those moves would eventually have
+been added to the abstraction in a refinement step. If the formula is
+\texttt{unsat} even with the opponent choosing moves then it must be
+\texttt{unsat} with default moves.
+
 \section{Evaluation}

### 2016-01-26 paper.tex
Some more related work

@@
 to synthesis as an alternative to BDDs. One approach is to find sets of states
-in CNF \cite{demiurge}. Another approach is to eschew states and focus on
+in CNF \cite{bloem2014}. Another approach is to eschew states and focus on
 \emph{runs} of the game. Previous work has applied this idea to realizability
@@
 strategy is not winning. This forms the basis of a counterexample guided
-abstraction refinement framework that operates on candidate strategies.
+abstraction refinement framework \cite{clarke2000} that operates on candidate strategies.
 
@@
 
+Synthesis of safety games is a thoroughly explored area of research with most
+efforts directed toward solving games with BDDs \cite{burch1990} and abstract
+interpretation \cite{walker2014,brenguier2014}. Satisfiability solving has been used
+previously for synthesis in a suite of methods proposed by Bloem et al
+\cite{bloem2014}. The authors propose a propose employing competing SAT solvers
+to learn clauses, which is similar to our approach but does not unroll the
+game. They also suggest QBF solver, template-based, and Effectively
+Propositional Logic (EPL) approaches.
+
+There are different approaches to bounded synthesis than the one described
+here. The authors of \cite{finkbeiner2013} suggest a methodology directly
+inspired by bounded model checking and it has been adapted to symbolic
+synthesis \cite{ehlers2010}. Lazy synthesis \cite{finkbeiner2012} is a
+counterexample-guided approach to bounded synthesis that refines an
+implementation for the game instead of an abstraction of it.
+
+SAT-based bounded model checking approaches that unroll the transition relation
+have been extended to unbounded by using conflicts in the solver
+\cite{mcmillan2002}, or by interpolation \cite{mcmillan2003}. However, there
+are no corresponding adaptations to synthesis. Incremental induction
+\cite{bradley2011} is another technique for unbounded model checking with an
+equivalent synthesis method \cite{morgenstern2013}, which computes the set of
+safe states incrementally.
+
 \section{Conclusion}

### 2016-01-26 paper.tex
Edits based on Leonid's feedback

@@
 \usetikzlibrary{trees}
+\usetikzlibrary{positioning}
 
@@
 
-The bounded synthesis algorithm begins with an empty game as seen in Figure
-\ref{fig:emptytree}.  Initially we are playing on behalf of the environment
-because it chooses the first move in each step.  The empty game is passed to
-the SAT solver, which searches for a candidate environment strategy.  If a
-candidate is found (Figure \ref{fig:candidate}) then it is checked for a
-spoiling strategy by solving for the controller.  If no spoiling strategy
-exists, that means our candidate is a winning strategy and the algorithm
-terminates.  Otherwise we find a counterexample (Figure \ref{fig:cex}), which
-is used to refine the empty game tree to include the first move from the
-controller's spoiling strategy (Figure \ref{fig:refined}).  The algorithm
-continues by finding a new candidate for the environment (Figure
+The bounded synthesis algorithm (Algorithm \ref{alg:bounded}) begins with an
+empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing on
+behalf of the environment because it chooses the first move in each step.  The
+empty game is passed to the SAT solver, which searches for a candidate
+environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
+then it is checked for a spoiling strategy by solving for the controller.  If
+no spoiling strategy exists, that means our candidate is a winning strategy and
+the algorithm terminates.  Otherwise we find a counterexample (Figure
+\ref{fig:cex}), which is used to refine the empty game tree to include the
+first move from the controller's spoiling strategy (Figure \ref{fig:refined}).
+The algorithm continues by finding a new candidate for the environment (Figure
 \ref{fig:candidate2}), and a new counterexample to refine the game abstraction
@@
             \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
-            \IIf{$sol = \texttt{unsat}$} 
-                \Return $\emptyset$ \Comment{No candidate exists}
-            \EndIIf
-            \State \Return $\{ \langle n, m \rangle | n \in $ \Call{gtNodes}{T} $, m = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+            \If{$sol = \texttt{unsat}$} 
+                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
+                    \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
+                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, T$}}{\Call{\textoverline{learn}}{$\sigma, T$}} \EndIfElse
+                \EndIf
+                \State \Return $\emptyset$ \Comment{No candidate exists}
+            \Else
+                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+            \EndIf
         \EndFunction
@@
 
+\subsection{Extending Bounded Synthesis to Unbounded Games}
+
+\begin{algorithm}[h]
+    \begin{algorithmic}[1]
+        \Function{solveUnbounded}{$T$}
+            \State $B^M \gets E$
+            \State $B^m[0] \gets E$
+            \For{$k = 1 \dots$}
+                \IIf{\Call{SAT}{$I \land B^M$}}
+                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
+                \EndIIf
+                \IIf{$\exists i \  B^m[i] \equiv B^m[i+1]$}
+                    \Return \texttt{realisable} \Comment{Reached fixed point}
+                \EndIIf
+                \State $B^m[k] \gets \top$
+                \State \Call{checkRank}{$I, k$}
+            \EndFor
+        \EndFunction
+        \algstore{u1}
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \algrestore{u1}
+        \Require \emph{May} and \emph{must} invariants hold
+        \Ensure \emph{May} and \emph{must} invariants hold
+        \Ensure $\lnot s \in B^m[k]$ if there is a winning controller strategy of length $k$ starting at $s$
+        \Ensure $s \in B^M$ if there is a winning environment strategy of length $k$ starting at $s$
+%%%        $\exists u_{k..1} \forall c_{k..1} \  s(x_k) \land (E_k \lor (\delta(x_k, u_k, c_k) \land E(x_{k-1}) \lor ... \land E(x_1) \lor (\delta(x_1, u_1, c_1) \land E(x_0))...)$
+%%%            $\forall u_{k..1} \exists c_{k..1} \  s(x_k) \land \lnot E_k \land \delta(x_k, u_k, c_k) \land \lnot E(x_{k-1}) \land ... \land \delta(x_1, u_1, c_1) \land \lnot E(x_0)$
+        \Function{checkRank}{$s, k$}
+            \State \Return \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
+        \EndFunction
+    \end{algorithmic}
+    \caption{Unbounded Synthesis}
+    \label{alg:unbounded}
+\end{algorithm}
+
+The unbounded synthesis algorithm (Algorithm \ref{alg:unbounded}) effectively
+consists of two communicating solvers: the counterexample-guided bounded
+reachability solver and the unbounded solver based on incremental induction.
+The unbounded solver calls the bounded solver with an increasing bound (line 8)
+while learning states from abstract game trees that either must or may lose for
+the controller. When the initial state is contained within the must-losing
+states the game is known to be unrealisable (line 5).
+
+May-losing states are learned by removing controller-winning states. We
+maintain a set of states for every rank up the current bound. The states we
+learn while executing the bounded algorithm can only said to be
+controller-winning for a particular rank or below. However, we construct each
+set carefully such that it is monotonically larger than the set of the rank
+below (see Figure \ref{fig:maylosing}) and so that the environment is unable to
+force play from one set to the next. Due to this careful construction, when two
+adjacent sets become equivalent we know that the algorithm has reached a fixed
+point and the controller is winning in the unbounded game (line 6).
+
+\begin{figure}
+    \centering
+    \begin{tikzpicture}
+        \node [minimum size=2cm,circle,draw] (a){$B^m[n]$};
+        \node [minimum size=1.5cm,circle,draw,right=1.8cm of a] (b){$B^m[n-1]$};
+        \node [minimum size=1cm,right=1.5cm of b] (dots){$\dots$};
+        \node [minimum size=0.5cm,circle,draw,right=1.5cm of dots] (c){$B^m[0]$};
+        \draw[-latex] (a.east) -- node [above] {$\supset$} (b.west);
+        \draw[-latex] (b.east) -- node [above] {$\supset$} (dots);
+        \draw[-latex] (dots.east) -- node [above] {$\supset$} (c.west);
+    \end{tikzpicture}
+    \caption{May-losing states}
+    \label{fig:maylosing}
+\end{figure}
+
+When executing the unbounded solver, lines 18 and 19 become active in the
+bounded solver. These lines call the learning procedures when the solver fails
+to find a candidate for an abstract game tree. The states symbolically
+represented by nodes in the tree are losing for whichever player could not find
+a winning candidate and can be extracted from the tree using interpolation.
+
 \subsection{Learning States with Interpolants}
@@
 \begin{algorithm}
-    \begin{algorithmic}[1]
-        \Require $\sigma(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
-        \Require \emph{Must-invariant} holds
-        \Ensure \emph{Must-invariant} holds
-        \Ensure $\sigma(X_T) \land B^M$
-        \Function{learn}{$\sigma, T$}
-            \IIf{$T = \emptyset$}
-                \Return
-            \EndIIf
-            \State $n \gets $ non-leaf node with max rank
-            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
-            \State $B^M \gets B^M \lor \mathcal{I}$
-            \State \Call{learn}{$\sigma, T_1$}
-        \EndFunction
-        \algstore{learn}
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \algrestore{learn}
-        \Require $\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
-        \Require \emph{May-invariant} holds
-        \Ensure \emph{May-invariant} holds
-        \Ensure $\sigma(X_T) \land B^m[$\Call{rank}{$T$}$] \equiv \bot$
-        \Function{\textoverline{learn}}{$\sigma, T$}
-            \IIf{$T = \emptyset$}
-                \Return
-            \EndIIf
-            \State $n \gets $ non-leaf node with max rank
-            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
-            \For{$i = 1$ to \Call{rank}{$n$}}
-                \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
-            \EndFor
-            \State \Call{\textoverline{learn}}{$\sigma, T_1$}
-        \EndFunction
-    \end{algorithmic}
-    \caption{Learning algorithms}
-    \label{alg:learn}
-\end{algorithm}
-
-\begin{algorithm}
     \caption{Amended tree formulas for Controller and Environment respectively}
@@
 
+\subsection{Correctness}
+
 \begin{algorithm}
     \begin{algorithmic}[1]
-        \Function{solveUnbounded}{$T$}
-            \State $B^M \gets E$
-            \State $B^m[0] \gets E$
-            \For{$k = 1 \dots$}
-                \IIf{\Call{SAT}{$I \land B^M$}}
-                    \Return \texttt{unrealisable} \Comment{Losing in the initial state}
+        \Require $\sigma(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
+        \Require \emph{Must-invariant} holds
+        \Ensure \emph{Must-invariant} holds
+        \Ensure $\sigma(X_T) \land B^M$
+        \Function{learn}{$\sigma, T$}
+            \IIf{$T = \emptyset$}
+                \Return
             \EndIIf
-                \IIf{$\exists i \  B^m[i] \equiv B^m[i+1]$}
-                    \Return \texttt{realisable} \Comment{Reached fixed point}
-                \EndIIf
-                \State $B^m[k] \gets \top$
-                \State \Call{checkRank}{$I, k$}
-            \EndFor
+            \State $n \gets $ non-leaf node with max rank
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
+            \State $B^M \gets B^M \lor \mathcal{I}$
+            \State \Call{learn}{$\sigma, T_1$}
         \EndFunction
-        \algstore{u1}
+        \algstore{learn}
     \end{algorithmic}
 
-    \begin{algorithmic}
-        \algrestore{u1}
-        \Require \emph{May} and \emph{must} invariants hold
-        \Ensure \emph{May} and \emph{must} invariants hold
-        \Ensure $\lnot s \in B^m[k]$ if there is a winning controller strategy of length $k$ starting at $s$
-        \Ensure $s \in B^M$ if there is a winning environment strategy of length $k$ starting at $s$
-%%%        $\exists u_{k..1} \forall c_{k..1} \  s(x_k) \land (E_k \lor (\delta(x_k, u_k, c_k) \land E(x_{k-1}) \lor ... \land E(x_1) \lor (\delta(x_1, u_1, c_1) \land E(x_0))...)$
-%%%            $\forall u_{k..1} \exists c_{k..1} \  s(x_k) \land \lnot E_k \land \delta(x_k, u_k, c_k) \land \lnot E(x_{k-1}) \land ... \land \delta(x_1, u_1, c_1) \land \lnot E(x_0)$
-        \Function{checkRank}{$s, k$}
-            \State \Return \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
-        \EndFunction
-        \algstore{u2}
-    \end{algorithmic}
-
-    \begin{algorithmic}
-        \algrestore{u2}
-        \Function{solveAbstract}{$p, \langle s, k \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$p, \langle s, k \rangle, T$} \Comment{Look for a candidate}
-        \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
-        \State $T' \gets T$
-        \Loop
-            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
-            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$p, \langle s, k \rangle, T, cand$} \Comment{Verify candidate}
-            \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
-            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$p, \langle s, k \rangle, T'$} \Comment{Solve refined AGT}
-        \EndLoop
-        \EndFunction
-        \algstore{u3}
-    \end{algorithmic}
-
-    \begin{algorithmic}
-        \algrestore{u3}
-        \Function{findCandidate}{$p, \langle s, k \rangle, T$}
-            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
-            \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
-            \If{$sol = \texttt{unsat}$} 
-                \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
-                \State \Call{learn}{$\sigma, T$}
-                \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, T$}}{\Call{\textoverline{learn}}{$\sigma, T$}} \EndIfElse
-                \State \Return $\emptyset$ \Comment{No candidate exists}
-            \Else
-                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
-            \EndIf
-        \EndFunction
-        \algstore{u4}
-    \end{algorithmic}
-
-    \begin{algorithmic}
-        \algrestore{u4}
-        \Function{verify}{$p, \langle s, k\rangle, gt, cand$}
-            \For{$l \in leaves(gt)$}
-            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
-                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
+    \begin{algorithmic}[1]
+        \algrestore{learn}
+        \Require $\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
+        \Require \emph{May-invariant} holds
+        \Ensure \emph{May-invariant} holds
+        \Ensure $\sigma(X_T) \land B^m[$\Call{rank}{$T$}$] \equiv \bot$
+        \Function{\textoverline{learn}}{$\sigma, T$}
+            \IIf{$T = \emptyset$}
+                \Return
+            \EndIIf
+            \State $n \gets $ non-leaf node with max rank
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
+            \For{$i = 1$ to \Call{rank}{$n$}}
+                \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
             \EndFor
-            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
+            \State \Call{\textoverline{learn}}{$\sigma, T_1$}
         \EndFunction
     \end{algorithmic}
-
-    \caption{Unbounded Synthesis}
-    \label{alg:unbounded}
+    \caption{Learning algorithms}
+    \label{alg:learn}
 \end{algorithm}
 
-\subsection{Correctness}
-
-The unbounded synthesis algorithm (Algorithm \ref{alg:unbounded}) effectively
-consists of two communicating solvers: the counterexample-guided bounded
-reachability solver and the unbounded solver based on incremental induction.
-Fortunately, correctness of both solvers can be established largely
-independently, which simplifies the proof of correctness of the overall
-algorithm.
+The correctness of the unbounded synthesis algorithm can be established
+independently from that of the bounded algorithm. Fortunately, the correctness
+of the bounded solver has been established in \cite{narodytska2014}.
 
@@
 
+We evaluate our approach on the benchmarks of the 2015 synthesis competition
+(SYNTCOMP'15).
+
 \section{Related Work}

### 2016-01-26 paper.tex
commented out authors for double-blind reviewing
made a pass over background definitions

@@
 \captionsetup{compatibility=false}
+\usepackage{amsmath}
 \usepackage{tikz}
@@
 
-\author{Alexander Legg\inst{1} 
-    \and Nina Narodytska\inst{2}
-    \and Leonid Ryzhyk\inst{2}}
+\author{}
+\institute{}
 
-\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
-    Communications and the Digital Economy and the Australian Research Council through the ICT
-    Centre of Excellence program.} and UNSW \\
-    \email{alexander.legg@nicta.com.au}
-    \and Samsung Research America}
+%\author{Alexander Legg\inst{1} 
+%    \and Nina Narodytska\inst{2}
+%    \and Leonid Ryzhyk\inst{2}}
+%
+%\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
+%    Communications and the Digital Economy and the Australian Research Council through the ICT
+%    Centre of Excellence program.} and UNSW \\
+%    \email{alexander.legg@nicta.com.au}
+%    \and Samsung Research America}
 
 \maketitle
+
+\sloppy
 
@@
 
-\section{Reactive Synthesis}
+\section{Background}
 
-A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E, \rangle$,
-consists of a set of boolean state variables, sets of uncontrollable and
-controllable label variables, a transition relationship $\delta : (X, L_u, L_c)
-\to X$, an initial state, and an error state. The \emph{controller} and
-\emph{environment} players choose controllable and uncontrollable labels
-respectively and the game proceeds according to $\delta$.
+A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E \rangle$,
+is defined over boolean state variables $X$, uncontrollable label variables $L_u$, and
+controllable label variables $L_c$.  $I$ is the initial state of the game given as a 
+valuation of state variables.  $E(X)$ is the set of error states represented by its 
+characteristic formula.  The transition relation $\delta(X, L_u, L_c, X')$ of the game 
+is a boolean formula that relates current state and label to the set of possible next 
+states of the game.  We assume deterministic games, where 
+$\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies (x'_1=x'_2)$.
 
-An \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
-c_n)$ is a chain of state and label pairs of length $n$ s.t.  $x_{k+1}
-\leftarrow \delta(x_k, u_k, c_k)$. A run is winning for the controller if $x_0
-= I \land \forall i \in \{1..n\} (x_i \neq E)$. In a bounded game of rank $n$
-all runs are restricted to length $n$, whereas unbounded games consider runs of
-infinite length. Since we consider only deterministic games a run is equivalent
-to a list of assignments to $L_c$ and $L_u$.
+At every round of the game, the \emph{environment} picks an
+uncontrollable label, the \emph{controller} responds by choosing a controllable 
+label and the game transitions into a new state according to $\delta$. 
+A \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
+c_n)$ is a chain of state and label pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$. 
+A run is winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot E(x_i))$. 
+In a \emph{bounded game} of rank $n$ all runs are restricted to length $n$, whereas unbounded 
+games consider runs of infinite length. Since we consider only deterministic games, a run 
+is uniquely described by a list of assignments to $L_c$ and $L_u$.
 
@@
 uncontrollable labels. A bounded run is winning for the environment if $x_0
-= I \land \exists i \in \{1..n\} (x_i = E)$ and an environment strategy is
-winning for a bounded game if there exists a run $(x_0, \pi^e(x_1), c_1), (x_1,
-\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ that wins for the environment.
-Safety games are zero sum, therefore the existence of a controller strategy
-implies the nonexistence of an environment strategy and vice versa.
+= I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is
+winning for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1,
+\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.
+Safety games are zero sum, therefore the existence of a winning controller strategy
+implies the nonexistence of a winning environment strategy and vice versa.
 
-\subsection{Abstract Game Trees}
+\subsection{Counterexample-guided bounded synthesis}
+
+We review the bounded synthesis algorithm by Narodytska et al.~\cite{narodytska2014},
+which is the main building block for our unbounded algorithm.
 
@@
 
-A strategy is equivalent to the set of all runs with the player's labels
+A strategy is equivalent to the set of all runs with the controller's labels
 obeying the strategy mapping $\pi$. Therefore, a strategy can also be
@@
 
-\subsection{Counterexample Guided Bounded Synthesis}
-
 By checking for the existence of a winning run for the environment in an
@@
 abstraction refinement framework \cite{clarke2000} that operates on candidate strategies.
-
-This technique was proposed by Narodytska et al \cite{narodytska2014}. In
-Section \ref{sect:unbounded} we propose an extension to synthesis of unbounded
-games but the original algorithm is outlined here for completeness.
 
@@
 
-One aspect of the algorithm proposed by Narodytska et al\cite{narodytska2014}
+One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
 is that it mimics real-world uses of synthesis by allowing cooperation between

### 2016-01-27 paper.tex
Edits to intro, related work, section 2

--- 2016-01-26 paper.tex
@@
 specification.  However, the reactive synthesis problem is 2EXPTIME-complete so
-naive algorithms are infeasible on even simple systems.
+naive algorithms are infeasible on anything but simple systems.
 
@@
 controller must prevent the environment from forcing the game into an error
-state.  Much of the complexity of reactive synthesis stems from tracking the
-set of states in which a player is winning.
+state. In this way the environment is equivalent to an existential search of
+the game and the controller is universal. Much of the complexity of reactive
+synthesis stems from managing the interactions of the alternating quantifiers
+on the state space of the game.
 
@@
 represent a relation on a set of game variables but in the worst case the
-representation may be exponential. This means that BDDs are not a
-one-size-fits-all solution for all reactive synthesis specifications.
+representation may be exponential in the number of variables. This means that
+BDDs are not a one-size-fits-all solution for all reactive synthesis
+specifications.
 
@@
 to synthesis as an alternative to BDDs. One approach is to find sets of states
-in CNF \cite{bloem2014}. Another approach is to eschew states and focus on
-\emph{runs} of the game. Previous work has applied this idea to realizability
-of bounded games \cite{narodytska2014} by forming abstract representations of
-the game.  In this paper, we extend this idea to unbounded games by
-constructing approximate sets of winning states from abstract trees.
+in CNF \cite{bloem2014,morgenstern2013}. Another approach is to eschew states
+and focus on \emph{runs} of the game. Previous work has applied this idea to
+realizability of bounded games \cite{narodytska2014} by forming abstractions of
+the game and refining in a counterexample-guided framework. This has been shown
+to outperform BDDs on certain classes of game specifications but is only able
+to find counterexamples of a certain length to the safety property.  In this
+paper, we extend this idea to unbounded games by approximating sets of unsafe
+states from abstract games. Careful construction ensures that a fixed point in
+these sets guarantees completeness.
+
+Section 2 outlines the original bounded synthesis algorithm. In Section 3 we
+describe and prove the correctness of our extension of the algorithm to
+unbounded games. In the following sections we discuss optimisations to the
+algorithm, evaluate our methodology, and compare our approach to other
+synthesis techniques.
 
@@
 winning in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)),
-(x_1, u_1, \pi^c(x_1, u_1)) \dots (x_n, u_n, \pi^c(x_n, u_n))$ are winning.
+(x_1, u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
 Bounded \emph{realizability} is the problem of determining the existence of
@@
 also check whether the error state has been reached in no branch
-(\emph{respectively}: in all branches) for the controller (\emph{resp}:
-environment.) The functions that produce these formulas are shown in Algorithm
+(\emph{respectively}: in all branches) for the controller (environment.) The
+functions that produce these formulas are shown in Algorithm
 \ref{alg:treeFormula}.
@@
 The \emph{must-invariant} guarantees that the must-losing set $B^M$ is an
-underapproximation of the actual losing set $B$: $$B^M \subseteq B$$.
+underapproximation of the actual losing set $B$: $$B^M \subseteq B.$$
 
@@
 Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any
-assignment to $X_n$ that satisfies $\sigma(X_T) \land |treeFormula|(T_1)$ also
-satisfies $\II$.  Hence, removing $\II$ from $B^m[\textsc{rank}(n)]$ makes
-$\sigma(X_T) \land |treeFormula|(T_1)$ unsatisfiable, and hence all
+assignment to $X_n$ that satisfies $\sigma(X_T) \land
+\textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
+removing $\II$ from $B^m[\textsc{rank}(n)]$ makes $\sigma(X_T) \land
+\textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
 preconditions of the recursive invocation of \textsc{\textoverline{learn}} in
@@
 
-There are different approaches to bounded synthesis than the one described
-here. The authors of \cite{finkbeiner2013} suggest a methodology directly
-inspired by bounded model checking and it has been adapted to symbolic
-synthesis \cite{ehlers2010}. Lazy synthesis \cite{finkbeiner2012} is a
-counterexample-guided approach to bounded synthesis that refines an
-implementation for the game instead of an abstraction of it.
-
 SAT-based bounded model checking approaches that unroll the transition relation
@@
 
+There are different approaches to bounded synthesis than the one described
+here. The authors of \cite{finkbeiner2013} suggest a methodology directly
+inspired by bounded model checking and it has been adapted to symbolic
+synthesis \cite{ehlers2010}. Lazy synthesis \cite{finkbeiner2012} is a
+counterexample-guided approach to bounded synthesis that refines an
+implementation for the game instead of an abstraction of it.
+
+The algorithm presented in this paper solves realisability for safety games.
+There is a method for extracting strategies from abstract game trees that is
+compatible with our method \cite{een2015}. It involves a similar interpolation
+approach for discovering states in game tree nodes.
+
 \section{Conclusion}

### 2016-01-27 paper.tex
Bounded section from cav'14 paper

@@
 
-\subsection{Counterexample-guided bounded synthesis}
-
-We review the bounded synthesis algorithm by Narodytska et al.~\cite{narodytska2014},
-which is the main building block for our unbounded algorithm.
-
-A set of runs can be represented symbolically by a tree of valuations to
-controllable and uncontrollable label variables where valuations are allowed to
-be \emph{unfixed}. An unfixed value symbolically represents all possible
-valuations. Formally, we define the tree as a set of lists of label valuations
-and a node in the tree may be identified by the list made by following the path
-from the root.
-
-The entire set of runs of a bounded game of rank $n$ is symbolically
-represented by a tree of depth $2n$ (a step is two moves - environment then
-controller) and width 1 populated by unfixed edges.  Fixing some moves and
-thereby reducing the set of runs forms an \emph{abstract game tree} that
-symbolically represents a smaller version of the game.
-
-A strategy is equivalent to the set of all runs with the controller's labels
-obeying the strategy mapping $\pi$. Therefore, a strategy can also be
-represented by a tree. Relaxing the restriction of the strategy mapping allows
-for a \emph{partial strategy} in which multiple labels are now possible for a
-single state.
-
-A strategy or partial strategy can also be thought of as an abstract game. A
-partial strategy for the controller is a restriction only on the controllable
-labels in the game. So if the environment can not win in the abstract game
-equivalent to the controller's partial strategy, then all strategies allowable
-by that partial strategy must be winning.
-
-An abstract game tree can be checked for the existence of a winning run by a
-SAT solver\cite{narodytska2014}.  The tree must be encoded into CNF by making
-copies of the transition relation for each game step in tree. The formula must
-also check whether the error state has been reached in no branch
-(\emph{respectively}: in all branches) for the controller (environment.) The
-functions that produce these formulas are shown in Algorithm
-\ref{alg:treeFormula}.
-
-\begin{algorithm}
-    \caption{Tree formulas for Controller and Environment respectively}
-    \label{alg:treeFormula}
-    \begin{algorithmic}[1]
-        \Function{treeFormula}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
-        \Else
-        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
-        \EndIf
-        \EndFunction
-        \algstore{tf1}
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \algrestore{tf1}
-        \Function{\textoverline{treeFormula}}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{\Call{E}{$x^{gt}$}}
-        \Else
-        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$ }
-        \EndIf
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-When we produce a formula using $\textsc{treeFormula}$ the SAT solver is
-playing on behalf on the controller. Any unfixed labels in the tree,
-controllable or uncontrollable, will be existentially quantified.  This means
-that if there exists a way for the game to be solved when both players
-cooperate on unfixed labels the SAT solver will find it. If no winning run
-exists in an abstract game even when the players are cooperating then there is
-no winning run when the opponent is playing adversarily.
-
-By checking for the existence of a winning run for the environment in an
-abstract game tree constructed from a partial controller strategy we are
-checking if the partial strategy is winning. If no environment-winning run is
-exists within the partial strategy then it is a winning partial strategy,
-otherwise the spoiling run is a counterexample that proves that the partial
-strategy is not winning. This forms the basis of a counterexample guided
-abstraction refinement framework \cite{clarke2000} that operates on candidate strategies.
-
-The bounded synthesis algorithm (Algorithm \ref{alg:bounded}) begins with an
-empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing on
-behalf of the environment because it chooses the first move in each step.  The
-empty game is passed to the SAT solver, which searches for a candidate
-environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
-then it is checked for a spoiling strategy by solving for the controller.  If
-no spoiling strategy exists, that means our candidate is a winning strategy and
-the algorithm terminates.  Otherwise we find a counterexample (Figure
-\ref{fig:cex}), which is used to refine the empty game tree to include the
-first move from the controller's spoiling strategy (Figure \ref{fig:refined}).
-The algorithm continues by finding a new candidate for the environment (Figure
-\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
-once again (Figure \ref{fig:refined2}). The full procedure is listed in
-Algorithm \ref{alg:bounded}.
+\subsection{Abstract Game Trees}
 
@@
     \centering
-    \begin{subfigure}{.3\textwidth}
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
             \node [circle,draw] (root){}
-                child {node [circle,draw] {} edge from parent[unfixed]
-                    child {node [circle,draw] {} edge from parent[unfixed]}
-                };
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
-        \caption{An empty AGT}
-        \label{fig:emptytree}
+        \captionsetup{width=.8\textwidth}
+        \caption{AGT}
+        \label{fig:agt}
     \end{subfigure}%
-    \begin{subfigure}{.35\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root) {}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {} edge from parent[unfixed]}
-                    edge from parent [fixed] node [left] {$u_1$}
-                };
-        \end{tikzpicture}
-        \caption{Environment candidate}
-        \label{fig:candidate}
-    \end{subfigure}%
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
             \node [circle,draw] (root){}
-                child {node [circle,draw] {} edge from parent[unfixed]
-                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-                    edge from parent [fixed] node [left] {$u_1$}
-                };
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        node [left=4pt] {$\epsilon$}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    node [left=4pt] {$\beta$}
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    node [left=4pt] {$\gamma$}
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [right=4pt] {$\alpha$}
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
-        \caption{Counterexample}
-        \label{fig:cex}
-    \end{subfigure}
-
-    \begin{subfigure}{.3\textwidth}
+        \captionsetup{width=.8\textwidth}
+        \caption{AGT with strategy}
+        \label{fig:strategy}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
             \node [circle,draw] (root){}
-                child {node [circle,draw] {} edge from parent[unfixed]
-                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-                };
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        node [left=4pt] {$\epsilon$}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    node [left=4pt] {$\beta$}
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] (sroot) {}
+                    child {node [below left=0 and -10pt] (sa) {}
+                        edge from parent [draw=none]
+                    }
+                    child {node [below right=0 and -10pt] (sb) {}
+                        edge from parent [draw=none]
+                    }
+                    node [left=4pt] {$\gamma$}
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [right=4pt] {$\alpha$}
+                node [above=4pt] {$\langle s, k \rangle$};
+                \draw[unfixed] (sa.center) -- (sb.center);
+                \draw[unfixed] (sroot.south) -- (sa.center);
+                \draw[unfixed] (sroot.south) -- (sb.center);
         \end{tikzpicture}
-        \caption{Refined AGT}
-        \label{fig:refined}
+        \captionsetup{width=.8\textwidth}
+        \caption{AGT with spoiling strategy}
+        \label{fig:spoiling}
     \end{subfigure}%
-    \begin{subfigure}{.35\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root) {}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {} edge from parent[unfixed]}
-                    edge from parent [fixed] node [left] {$u_2$}
-                };
-        \end{tikzpicture}
-        \caption{New candidate}
-        \label{fig:candidate2}
-    \end{subfigure}%
-    \begin{subfigure}{.4\textwidth}
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
             \node [circle,draw] (root){}
-                child {node [circle,draw] {} edge from parent[unfixed]
-                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-                    child {node [circle,draw] {} edge from parent[fixed] node [right] {$c_2$}}
-                };
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {$e$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
+        \captionsetup{width=.8\textwidth}
         \caption{Refined AGT}
-        \label{fig:refined2}
+        \label{fig:refined}
     \end{subfigure}
+%%%    \begin{subfigure}{.3\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root){}
+%%%                child {node [circle,draw] {} edge from parent[unfixed]
+%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{An empty AGT}
+%%%        \label{fig:emptytree}
+%%%    \end{subfigure}%
+%%%    \begin{subfigure}{.35\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root) {}
+%%%                child {node [circle,draw] {}
+%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
+%%%                    edge from parent [fixed] node [left] {$u_1$}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{Environment candidate}
+%%%        \label{fig:candidate}
+%%%    \end{subfigure}
+
+%%%    \begin{subfigure}{.4\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root){}
+%%%                child {node [circle,draw] {} edge from parent[unfixed]
+%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+%%%                    edge from parent [fixed] node [left] {$u_1$}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{Counterexample}
+%%%        \label{fig:cex}
+%%%    \end{subfigure}%
+%%%    \begin{subfigure}{.3\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root){}
+%%%                child {node [circle,draw] {} edge from parent[unfixed]
+%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{Refined AGT}
+%%%        \label{fig:refined}
+%%%    \end{subfigure}%
+%%%    \begin{subfigure}{.35\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root) {}
+%%%                child {node [circle,draw] {}
+%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
+%%%                    edge from parent [fixed] node [left] {$u_2$}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{New candidate}
+%%%        \label{fig:candidate2}
+%%%    \end{subfigure}%
+%%%    \begin{subfigure}{.4\textwidth}
+%%%        \centering
+%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%            \node [circle,draw] (root){}
+%%%                child {node [circle,draw] {} edge from parent[unfixed]
+%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
+%%%                    child {node [circle,draw] {} edge from parent[fixed] node [right] {$c_2$}}
+%%%                };
+%%%        \end{tikzpicture}
+%%%        \caption{Refined AGT}
+%%%        \label{fig:refined2}
+%%%    \end{subfigure}
     \caption{Abstract game trees}
+    \label{fig:alltrees}
 \end{figure}
+
+We review the bounded synthesis algorithm by Narodytska et
+al.~\cite{narodytska2014}, which is the main building block for our unbounded
+algorithm.  The algorithm constructs a series of abstractions of the input
+game.  An abstraction restricts actions available to one of the players.
+Specifically, we consider abstractions represented as trees of actions,
+referred to as \emph{abstract game trees}.  Together with a state-round pair
+$\langle s,k\rangle$, an abstract game tree defines an \emph{abstract game}
+played from this state.  Figure~\ref{fig:agt} shows an example abstract game
+tree.  In the abstract game, the environment player is required to pick actions
+from the tree, starting from the root node.  After reaching a leaf, it
+continues playing unrestricted.  The tree in Figure~\ref{fig:agt} restricts the
+initial environment action to the set $\{a, d\}$.  After choosing action $d$,
+the environment reaches a leaf of the tree and continues playing unrestricted.
+Alternatively, after choosing $a$, the environment is required to play action
+$b$ in the next round.
+
+Nodes of an abstract game tree are uniquely identified by the list of edge
+labels along the path from the root to the node.  We identify an abstract game
+tree with the set of its nodes.  For example, the tree in Figure~\ref{fig:agt}
+can be written as $\{(),(d),(a),(a,b)\}$.  We denote $\textsc{leaves}(T)$ the
+subset of leaf nodes of a tree $T$.
+
+A \emph{partial strategy} $Strat: T \rightarrow L_c $ assigns a controllable
+action to be played in each node of the abstract game tree.
+Figure~\ref{fig:strategy} shows an example partial strategy.  The controller
+starts by choosing action $\alpha$.  If the environment plays $a$, the
+controller responds with $\beta$ in the next round, and so on. Given a partial
+strategy $Strat$, we can map each leaf $l$ of the abstract game tree to
+$\langle s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained
+by playing all controllable and uncontrollable actions on the path from the
+root to the leaf.
+
+%%%A set of runs can be represented symbolically by a tree of valuations to
+%%%controllable and uncontrollable label variables where valuations are allowed to
+%%%be \emph{unfixed}. An unfixed value symbolically represents all possible
+%%%valuations. Formally, we define the tree as a set of lists of label valuations
+%%%and a node in the tree may be identified by the list made by following the path
+%%%from the root.
+
+%%%The entire set of runs of a bounded game of rank $n$ is symbolically
+%%%represented by a tree of depth $2n$ (a step is two moves - environment then
+%%%controller) and width 1 populated by unfixed edges.  Fixing some moves and
+%%%thereby reducing the set of runs forms an \emph{abstract game tree} that
+%%%symbolically represents a smaller version of the game.
+
+%%%A strategy is equivalent to the set of all runs with the controller's labels
+%%%obeying the strategy mapping $\pi$. Therefore, a strategy can also be
+%%%represented by a tree. Relaxing the restriction of the strategy mapping allows
+%%%for a \emph{partial strategy} in which multiple labels are now possible for a
+%%%single state.
+
+%%%A strategy or partial strategy can also be thought of as an abstract game. A
+%%%partial strategy for the controller is a restriction only on the controllable
+%%%labels in the game. So if the environment can not win in the abstract game
+%%%equivalent to the controller's partial strategy, then all strategies allowable
+%%%by that partial strategy must be winning.
+
+%%%An abstract game tree can be checked for the existence of a winning run by a
+%%%SAT solver\cite{narodytska2014}.  The tree must be encoded into CNF by making
+%%%copies of the transition relation for each game step in tree. The formula must
+%%%also check whether the error state has been reached in no branch
+%%%(\emph{respectively}: in all branches) for the controller (environment.) The
+%%%functions that produce these formulas are shown in Algorithm
+%%%\ref{alg:treeFormula}.
+
+%%%When we produce a formula using $\textsc{treeFormula}$ the SAT solver is
+%%%playing on behalf on the controller. Any unfixed labels in the tree,
+%%%controllable or uncontrollable, will be existentially quantified.  This means
+%%%that if there exists a way for the game to be solved when both players
+%%%cooperate on unfixed labels the SAT solver will find it. If no winning run
+%%%exists in an abstract game even when the players are cooperating then there is
+%%%no winning run when the opponent is playing adversarily.
+
+%%%By checking for the existence of a winning run for the environment in an
+%%%abstract game tree constructed from a partial controller strategy we are
+%%%checking if the partial strategy is winning. If no environment-winning run
+%%%exists within the partial strategy then it is a winning partial strategy,
+%%%otherwise the spoiling run is a counterexample that proves that the partial
+%%%strategy is not winning. This forms the basis of a counterexample guided
+%%%abstraction refinement framework \cite{clarke2000} that operates on candidate strategies.
+
+\subsection{Counterexample-guided bounded synthesis}
 
@@
 \end{algorithm}
+
+Algorithm~\ref{alg:bounded} illustrates the bounded synthesis algorithm. The
+algorithm takes a concrete game $G$ as an implicit argument. In addition, it
+takes a player (controller or environment), state-round pair $\langle s,
+k\rangle$ and an abstract game tree $T$ and returns a winning partial strategy
+for it, if one exists.  The initial invocation of the algorithm takes the
+initial state $\langle I, 0\rangle$ and an empty abstract game tree
+$\emptyset$. Initially the solver is playing on behalf of the environment since
+that player takes the first move in every step.  The empty game tree does not
+constrain opponent moves, hence solving such an abstraction is equivalent to
+solving the original concrete game.  The algorithm is organised as a
+counterexample-guided abstraction refinement (CEGAR) loop.  The first step of
+the algorithm uses the \textsc{findCandidate} function, described in detail
+below, to come up with a candidate partial strategy for $T$. If it fails to
+find a strategy, this means that no winning partial strategy exists for $T$.
+If, on the other hand, a candidate partial strategy is found, we need to verify
+if it is indeed winning for $T$.
+
+The \textsc{verify} procedure searches for a \emph{spoiling} counterexample
+strategy in each leaf of the candidate partial strategy by calling
+\textsc{solveAbstract} for the opponent. The dual solver solves games on behalf
+of the controller player.  {Figure~\ref{fig:spoiling}} shows a spoiling
+strategy discovered one of the leaves of the partial strategy.
+
+If the dual solver can find no spoiling strategy at any of the leaves, then the
+candidate partial strategy is a winning one. Otherwise, \textsc{verify} returns
+the move used by the opponent to defeat a leaf of the partial strategy, which
+is appended to the corresponding node in $T$ in order to refine it in line~(9)
+as shown in Figure~(\ref{fig:refined}).
+
+We solve the refined game by recursively invoking \textsc{solveAbstract} on it.
+If no partial winning strategy is found for the refined game then there is also
+no partial winning strategy for the original abstract game, and the algorithm
+returns a failure.  Otherwise, the partial strategy for the refined game is
+\emph{projected} on the original abstract game by removing the leaves
+introduced by refinements. The resulting partial strategy becomes a candidate
+strategy to be verified at the next iteration of the loop. In the worst case
+the loop terminates after all actions in the game are refined into the abstract
+game.
+
+The way game trees are refined is determined by the formula passed to the SAT
+solver in $\textsc{findCandidate}$. This formula encodes an abstract game tree
+into CNF by making copies of the transition relation for each game step in the
+tree.  The formula must also check whether the error state has been reached in
+no branch (\emph{respectively}: in all branches) for the controller
+(environment.) This procedure is defined in Algorithm \ref{alg:treeFormula} for
+both the controller in $\textsc{treeFormula}$ and the environment in
+$\textsc{\textoverline{treeFormula}}$.
+
+Since the game tree formulation is passed to a SAT solver, any unfixed labels
+in the tree - both controllable and uncontrollable - will be existentially
+quantified. This means that the SAT solver will find any way to win the game
+while both players are cooperating. If no winning run exists in an abstract
+game even when the players are cooperating then there is no winning run when
+the opponent is playing adversarily. When a winning run is found, the actions
+chosen by the SAT solver are used to refine the game tree. This is advantageous
+for many synthesis problems where the game must be formalised as adversarial
+for correctness but the final implementation will be cooperating with its
+environment in the real world. An example of such a system is a device driver
+that cooperates with the device and OS to interface between the two.
+
+\begin{algorithm}
+    \caption{Tree formulas for Controller and Environment respectively}
+    \label{alg:treeFormula}
+    \begin{algorithmic}[1]
+        \Function{treeFormula}{gt}
+        \If{$\Call{rank}{gt} = 0$}
+        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
+        \Else
+        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \EndIf
+        \EndFunction
+        \algstore{tf1}
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \algrestore{tf1}
+        \Function{\textoverline{treeFormula}}{gt}
+        \If{$\Call{rank}{gt} = 0$}
+        \State \Return{\Call{E}{$x^{gt}$}}
+        \Else
+        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+%%%The bounded synthesis algorithm (Algorithm \ref{alg:bounded}) begins with an
+%%%empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing on
+%%%behalf of the environment because it chooses the first move in each step.  The
+%%%empty game is passed to the SAT solver, which searches for a candidate
+%%%environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
+%%%then it is checked for a spoiling strategy by solving for the controller.  If
+%%%no spoiling strategy exists, that means our candidate is a winning strategy and
+%%%the algorithm terminates.  Otherwise we find a counterexample (Figure
+%%%\ref{fig:cex}), which is used to refine the empty game tree to include the
+%%%first move from the controller's spoiling strategy (Figure \ref{fig:refined}).
+%%%The algorithm continues by finding a new candidate for the environment (Figure
+%%%\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
+%%%once again (Figure \ref{fig:refined2}). The full procedure is listed in
+%%%Algorithm \ref{alg:bounded}.
 

### 2016-01-27 paper.tex
Wrote most of the prose for evaluation, still needs results

@@
 \usepackage{amsmath}
+\usepackage{mathtools}
 \usepackage{tikz}
@@
     \begin{algorithmic}[1]
-        \Function{treeFormula}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{ $\lnot \Call{E}{x^{gt}}$ }
+        \Function{treeFormula}{$T$}
+        \If{$\Call{rank}{T} = 0$}
+        \State \Return{ $\lnot \Call{E}{X_{T}}$ }
         \Else
-        \State \Return{ $\lnot \Call{E}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
+            $$\bigwedge_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{treeFormula}{n})$$
+        }
         \EndIf
@@
         \algrestore{tf1}
-        \Function{\textoverline{treeFormula}}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{\Call{E}{$x^{gt}$}}
+        \Function{\textoverline{treeFormula}}{$T$}
+        \If{$\Call{rank}{T} = 0$}
+        \State \Return{\Call{E}{$X_{T}$}}
         \Else
-        \State \Return{ $\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$ }
+        \State \Return{ $\Call{E}{X_{T}} \lor$ \\
+        $$\bigvee_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$$ }
         \EndIf
@@
 We evaluate our approach on the benchmarks of the 2015 synthesis competition
-(SYNTCOMP'15).
+(SYNTCOMP'15). Each benchmark comprises of controllable and uncontrollable
+inputs to a circuit that assigns values to latches. One latch is configured as
+the error bit that determines the winner of the safety game. The benchmark
+suite is a collection of both real-world and toy specifications including
+generalised buffers, AMBA bus controllers, device drivers, and converted LTL
+formulas.  Descriptions of many of the benchmark families used can be found in
+the 2014 competition report~\cite{jacobs2015}.
+
+The implementation of our algorithm uses \textsc{Glucose}~\cite{audemard2014}
+for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
+generation. We compared our implementation to two solvers that competed in the
+sequential realisability track of SYNTCOMP'15.  \textsc{Simple BDD
+Solver}~\cite{walker2014} because it had the best results in the track, and
+\textsc{Demiurge}~\cite{bloem2014} because it was the only SAT-based tool to
+compete. The benchmarks were run on a cluster of Intel Quad Core Xeon E5405
+CPUs with 16GB of memory. Each solver was allowed exclusive access to a node
+for one hour to solve an instance.
+
+INSERT GRAPHS
+
+Our implementation was able to solver 103 out of the 250 specification in the
+alloted time, including 10 instances that were unable to be solved by any other
+solver including all other solvers in SYNTCOMP'15. \textsc{Simple BDD Solver}
+and \textsc{Demiurge} were able to solve X and Y respectively. Something about
+us vs BDDs and us vs demiurge with numbers of specs.
+
+Four of the previously unsolved instances are device driver instances and
+another five are toy examples. This supports the hypothesis that different game
+solving methodologies perform better on certain classes of specifications.
 

### 2016-01-27 paper.tex
tweaked the bounded section

@@
 A run is winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot E(x_i))$. 
-In a \emph{bounded game} of rank $n$ all runs are restricted to length $n$, whereas unbounded 
+In a \emph{bounded game} with bound $n$ all runs are restricted to length $n$, whereas unbounded 
 games consider runs of infinite length. Since we consider only deterministic games, a run 
@@
 
-\subsection{Abstract Game Trees}
+\subsection{Counterexample-guided bounded synthesis}
 
@@
         \captionsetup{width=.8\textwidth}
-        \caption{AGT with strategy}
+        \caption{AGT with partial strategy}
         \label{fig:strategy}
     \end{subfigure}%
-    \begin{subfigure}[t]{.25\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        node [left=4pt] {$\epsilon$}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    node [left=4pt] {$\beta$}
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] (sroot) {}
-                    child {node [below left=0 and -10pt] (sa) {}
-                        edge from parent [draw=none]
-                    }
-                    child {node [below right=0 and -10pt] (sb) {}
-                        edge from parent [draw=none]
-                    }
-                    node [left=4pt] {$\gamma$}
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [right=4pt] {$\alpha$}
-                node [above=4pt] {$\langle s, k \rangle$};
-                \draw[unfixed] (sa.center) -- (sb.center);
-                \draw[unfixed] (sroot.south) -- (sa.center);
-                \draw[unfixed] (sroot.south) -- (sb.center);
-        \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
-        \caption{AGT with spoiling strategy}
-        \label{fig:spoiling}
-    \end{subfigure}%
+%    \begin{subfigure}[t]{.25\textwidth}
+%        \centering
+%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%            \node [circle,draw] (root){}
+%                child {node [circle,draw] {}
+%                    child {node [circle,draw] {}
+%                        node [left=4pt] {$\epsilon$}
+%                        edge from parent [fixed] node [left] {$b$}
+%                    }
+%                    node [left=4pt] {$\beta$}
+%                    edge from parent [fixed] node [left] {$a$}
+%                }
+%                child {node [circle,draw] (sroot) {}
+%                    child {node [below left=0 and -10pt] (sa) {}
+%                        edge from parent [draw=none]
+%                    }
+%                    child {node [below right=0 and -10pt] (sb) {}
+%                        edge from parent [draw=none]
+%                    }
+%                    node [left=4pt] {$\gamma$}
+%                    edge from parent [fixed] node [right] {$d$}
+%                }
+%                node [right=4pt] {$\alpha$}
+%                node [above=4pt] {$\langle s, k \rangle$};
+%                \draw[unfixed] (sa.center) -- (sb.center);
+%                \draw[unfixed] (sroot.south) -- (sa.center);
+%                \draw[unfixed] (sroot.south) -- (sb.center);
+%        \end{tikzpicture}
+%        \captionsetup{width=.8\textwidth}
+%        \caption{AGT with spoiling strategy}
+%        \label{fig:spoiling}
+%    \end{subfigure}%
     \begin{subfigure}[t]{.25\textwidth}
@@
 We review the bounded synthesis algorithm by Narodytska et
-al.~\cite{narodytska2014}, which is the main building block for our unbounded
+al.~\cite{narodytska2014}, which is the main building block of our unbounded
 algorithm.  The algorithm constructs a series of abstractions of the input
-game.  An abstraction restricts actions available to one of the players.
+game.  
+An abstraction restricts actions available to one of the players. 
+%If the player can win the abstract game, then it can also win the original 
+%concrete game; however the opposite is not generally true.
 Specifically, we consider abstractions represented as trees of actions,
-referred to as \emph{abstract game trees}.  Together with a state-round pair
-$\langle s,k\rangle$, an abstract game tree defines an \emph{abstract game}
-played from this state.  Figure~\ref{fig:agt} shows an example abstract game
-tree.  In the abstract game, the environment player is required to pick actions
+referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an example abstract game
+tree for the controller (abstract game trees for the environment are similar).  In the abstract game, 
+the controller is required to pick actions
 from the tree, starting from the root node.  After reaching a leaf, it
 continues playing unrestricted.  The tree in Figure~\ref{fig:agt} restricts the
-initial environment action to the set $\{a, d\}$.  After choosing action $d$,
-the environment reaches a leaf of the tree and continues playing unrestricted.
-Alternatively, after choosing $a$, the environment is required to play action
-$b$ in the next round.
-
-Nodes of an abstract game tree are uniquely identified by the list of edge
-labels along the path from the root to the node.  We identify an abstract game
-tree with the set of its nodes.  For example, the tree in Figure~\ref{fig:agt}
-can be written as $\{(),(d),(a),(a,b)\}$.  We denote $\textsc{leaves}(T)$ the
-subset of leaf nodes of a tree $T$.
-
-A \emph{partial strategy} $Strat: T \rightarrow L_c $ assigns a controllable
-action to be played in each node of the abstract game tree.
-Figure~\ref{fig:strategy} shows an example partial strategy.  The controller
-starts by choosing action $\alpha$.  If the environment plays $a$, the
-controller responds with $\beta$ in the next round, and so on. Given a partial
+initial controller action to the set $\{a, d\}$.  After choosing action $a$,
+the controller is required to play action $b$ in the next round.  It then reaches a leaf 
+of the tree and continues playing unrestricted.
+The root of the tree is annotated by the initial state $s$ of the abstract game and the 
+bound $k$ on the number of rounds.
+%Together with a state-round pair
+%$\langle s,k\rangle$, an abstract game tree defines an \emph{abstract game}
+%played from this state.
+%Nodes of an abstract game tree are uniquely identified by the list of edge
+%labels along the path from the root to the node.  
+%We identify an abstract game tree with the set of its nodes.  
+We denote $\textsc{nodes}(T)$ the set of all nodes of a tree $T$ and
+$\textsc{leaves}(T)$ the subset of leaf nodes.
+
+Given a controller (environment)
+abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow L_u$ 
+($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with the opponent's
+action to be played in this node.   
+Figure~\ref{fig:strategy} shows an example partial strategy.  The environment 
+starts by choosing action $\alpha$.  If the controller plays $a$, the
+environment responds with $\beta$ in the next round, and so on. Given a partial
 strategy $Strat$, we can map each leaf $l$ of the abstract game tree to
-$\langle s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained
+$\langle s',k'\rangle=\textsc{outcome}(s, k, Strat, l)$, where $s'$ is the state obtained
 by playing all controllable and uncontrollable actions on the path from the
-root to the leaf.
+root to the leaf, and $i'$ is the number of rounds left in the game after reaching the leaf.
+An opponent's partial strategy is winning if for all its outcomes $\langle s',k'\rangle$,
+$s'$ is a winning state for the opponent in the concrete game with bound $k'$.
+
+%The abstract game tree restricts moves available to the player, thus making the game easier
+%to win for the opponent.  Hence, if the opponent does not have a winning partial strategy in
+%the abstract game, it also does not have a winning strategy in the concrete game.
 
@@
 
-\subsection{Counterexample-guided bounded synthesis}
-
 \begin{algorithm}
     \begin{algorithmic}[1]
-        \Function{solveAbstract}{$p, \langle s, k \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$p, \langle s, k \rangle, T$} \Comment{Look for a candidate}
-        \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \Function{solveAbstract}{$p, s, k, T$}
+        \State $cand \gets $ \Call{findCandidate}{$p, s, k, T$} \Comment{Look for a candidate}
+        \IIf{$k = 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
         \State $T' \gets T$
@@
             \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
-            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$p, \langle s, k \rangle, T, cand$} \Comment{Verify candidate}
+            \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$p, s, k, T, cand$} \Comment{Verify candidate}
             \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
-            \State $T' \gets $ \Call{gtAppend}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$p, \langle s, k \rangle, T'$} \Comment{Solve refined AGT}
+            \State $T' \gets $ \Call{append}{$T', l, u$} \Comment{Refine AGT with counterexample}
+            \State $cand \gets $ \Call{solveAbstract}{$p, s, k, T'$} \Comment{Solve refined AGT}
         \EndLoop
@@
         \algrestore{b1}
-        \Function{findCandidate}{$p, \langle s, k \rangle, T$}
+        \Function{findCandidate}{$p, s, k, T$}
             \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
@@
             \Else
-                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{T} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
             \EndIf
@@
         \algrestore{b2}
-        \Function{verify}{$p, \langle s, k\rangle, gt, cand$}
+        \Function{verify}{$p, s, k, T, cand$}
             \For{$l \in leaves(gt)$}
-            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$\langle s, k\rangle, gt, l$} \Comment{Get rank and state at leaf}
-                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $\langle k', s' \rangle, \emptyset$} \Comment{Solve for the opponent}
+            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$s, k, cand, l$} \Comment{Get rank and state at leaf}
+                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $k'$, $s'$, $\emptyset$} \Comment{Solve for the opponent}
                 \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
@@
 Algorithm~\ref{alg:bounded} illustrates the bounded synthesis algorithm. The
-algorithm takes a concrete game $G$ as an implicit argument. In addition, it
-takes a player (controller or environment), state-round pair $\langle s,
-k\rangle$ and an abstract game tree $T$ and returns a winning partial strategy
-for it, if one exists.  The initial invocation of the algorithm takes the
-initial state $\langle I, 0\rangle$ and an empty abstract game tree
+algorithm takes a concrete game $G$ with bound $n$ as an implicit argument. In addition, it
+takes a player $p$ (controller or environment), state $s$, bound
+$k$ and an abstract game tree $T$ and returns a winning partial strategy
+for $opponent(p)$, if one exists.  The initial invocation of the algorithm takes the
+initial state $\langle I, n\rangle$ and an empty abstract game tree
 $\emptyset$.  Initially the solver is playing on behalf of the environment since
@@
 counterexample-guided abstraction refinement (CEGAR) loop.  The first step of
-the algorithm uses the \textsc{findCandidate} function, described in detail
-below, to come up with a candidate partial strategy for $T$. If it fails to
+the algorithm uses the \textsc{findCandidate} function, described below, to come 
+up with a candidate partial strategy for $T$. If it fails to
 find a strategy, this means that no winning partial strategy exists for $T$.
@@
 \textsc{solveAbstract} for the opponent. The dual solver solves games on behalf
-of the controller player.  {Figure~\ref{fig:spoiling}} shows a spoiling
-strategy discovered one of the leaves of the partial strategy.
+of the controller player.  
+%{Figure~\ref{fig:spoiling}} shows a spoiling
+%strategy discovered one of the leaves of the partial strategy.
 

### 2016-01-28 paper.tex
Started adding high level explanations

--- 2016-01-27 paper.tex
@@
 games consider runs of infinite length. Since we consider only deterministic games, a run 
-is uniquely described by a list of assignments to $L_c$ and $L_u$.
+is uniquely described by a list of assignments to $L_u$ and $L_c$.
 
@@
 implies the nonexistence of a winning environment strategy and vice versa.
+
+\subsection{Realisability of Bounded Safety Games}
+
+We review the bounded synthesis algorithm by Narodytska et
+al.~\cite{narodytska2014}, which is the main building block for our unbounded
+algorithm.
+
+Realisability for a bounded safety game can be solved by checking an
+quantified formula consisting of the game's transition relation \emph{unrolled}
+to the length of the bounded game. The quantifiers represent the two players of
+the game: universal for the controller and existential for the environment.
+Just as the players alternate in choosing actions every step of the game, there
+must be quantifier alternations in the formula for each unrolling.
+
+Instead of solving both quantifiers at once, two competing solvers are used to
+solve abstractions of the game. Each solver builds a candidate strategy for its
+corresponding player and uses its opponent's candidate to restrict its own
+search. The solvers communicate to refine one another's candidate strategies by
+finding counterexamples. Eventually one solver will be unable to find a
+counterexample indicating that its opponent's candidate strategy is a winning
+strategy for the game.
 
@@
     \end{subfigure}
-%%%    \begin{subfigure}{.3\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root){}
-%%%                child {node [circle,draw] {} edge from parent[unfixed]
-%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{An empty AGT}
-%%%        \label{fig:emptytree}
-%%%    \end{subfigure}%
-%%%    \begin{subfigure}{.35\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root) {}
-%%%                child {node [circle,draw] {}
-%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
-%%%                    edge from parent [fixed] node [left] {$u_1$}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{Environment candidate}
-%%%        \label{fig:candidate}
-%%%    \end{subfigure}
-
-%%%    \begin{subfigure}{.4\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root){}
-%%%                child {node [circle,draw] {} edge from parent[unfixed]
-%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-%%%                    edge from parent [fixed] node [left] {$u_1$}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{Counterexample}
-%%%        \label{fig:cex}
-%%%    \end{subfigure}%
-%%%    \begin{subfigure}{.3\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root){}
-%%%                child {node [circle,draw] {} edge from parent[unfixed]
-%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{Refined AGT}
-%%%        \label{fig:refined}
-%%%    \end{subfigure}%
-%%%    \begin{subfigure}{.35\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root) {}
-%%%                child {node [circle,draw] {}
-%%%                    child {node [circle,draw] {} edge from parent[unfixed]}
-%%%                    edge from parent [fixed] node [left] {$u_2$}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{New candidate}
-%%%        \label{fig:candidate2}
-%%%    \end{subfigure}%
-%%%    \begin{subfigure}{.4\textwidth}
-%%%        \centering
-%%%        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%            \node [circle,draw] (root){}
-%%%                child {node [circle,draw] {} edge from parent[unfixed]
-%%%                    child {node [circle,draw] {} edge from parent[fixed] node [left] {$c_1$}}
-%%%                    child {node [circle,draw] {} edge from parent[fixed] node [right] {$c_2$}}
-%%%                };
-%%%        \end{tikzpicture}
-%%%        \caption{Refined AGT}
-%%%        \label{fig:refined2}
-%%%    \end{subfigure}
     \caption{Abstract game trees}
@@
 
-We review the bounded synthesis algorithm by Narodytska et
-al.~\cite{narodytska2014}, which is the main building block for our unbounded
-algorithm.  The algorithm constructs a series of abstractions of the input
-game.  An abstraction restricts actions available to one of the players.
+An abstraction of the game restricts actions available to one of the players.
 Specifically, we consider abstractions represented as trees of actions,
@@
 $\langle s,k\rangle$, an abstract game tree defines an \emph{abstract game}
-played from this state.  Figure~\ref{fig:agt} shows an example abstract game
-tree.  In the abstract game, the environment player is required to pick actions
-from the tree, starting from the root node.  After reaching a leaf, it
-continues playing unrestricted.  The tree in Figure~\ref{fig:agt} restricts the
-initial environment action to the set $\{a, d\}$.  After choosing action $d$,
-the environment reaches a leaf of the tree and continues playing unrestricted.
-Alternatively, after choosing $a$, the environment is required to play action
-$b$ in the next round.
+played from this state.  In the abstract game, the environment player is
+required to pick actions from the tree, starting from the root node.  After
+reaching a leaf, it continues playing unrestricted.  The tree in
+Figure~\ref{fig:agt} restricts the initial environment action to the set $\{a,
+d\}$.  After choosing action $d$, the environment reaches a leaf of the tree
+and continues playing unrestricted.  Alternatively, after choosing $a$, the
+environment is required to play action $b$ in the next round.
 
@@
 
-%%%A set of runs can be represented symbolically by a tree of valuations to
-%%%controllable and uncontrollable label variables where valuations are allowed to
-%%%be \emph{unfixed}. An unfixed value symbolically represents all possible
-%%%valuations. Formally, we define the tree as a set of lists of label valuations
-%%%and a node in the tree may be identified by the list made by following the path
-%%%from the root.
+\subsection{Counterexample-guided bounded synthesis}
 
-%%%The entire set of runs of a bounded game of rank $n$ is symbolically
-%%%represented by a tree of depth $2n$ (a step is two moves - environment then
-%%%controller) and width 1 populated by unfixed edges.  Fixing some moves and
-%%%thereby reducing the set of runs forms an \emph{abstract game tree} that
-%%%symbolically represents a smaller version of the game.
-
-%%%A strategy is equivalent to the set of all runs with the controller's labels
-%%%obeying the strategy mapping $\pi$. Therefore, a strategy can also be
-%%%represented by a tree. Relaxing the restriction of the strategy mapping allows
-%%%for a \emph{partial strategy} in which multiple labels are now possible for a
-%%%single state.
-
-%%%A strategy or partial strategy can also be thought of as an abstract game. A
-%%%partial strategy for the controller is a restriction only on the controllable
-%%%labels in the game. So if the environment can not win in the abstract game
-%%%equivalent to the controller's partial strategy, then all strategies allowable
-%%%by that partial strategy must be winning.
-
-%%%An abstract game tree can be checked for the existence of a winning run by a
-%%%SAT solver\cite{narodytska2014}.  The tree must be encoded into CNF by making
-%%%copies of the transition relation for each game step in tree. The formula must
-%%%also check whether the error state has been reached in no branch
-%%%(\emph{respectively}: in all branches) for the controller (environment.) The
-%%%functions that produce these formulas are shown in Algorithm
-%%%\ref{alg:treeFormula}.
-
-%%%When we produce a formula using $\textsc{treeFormula}$ the SAT solver is
-%%%playing on behalf on the controller. Any unfixed labels in the tree,
-%%%controllable or uncontrollable, will be existentially quantified.  This means
-%%%that if there exists a way for the game to be solved when both players
-%%%cooperate on unfixed labels the SAT solver will find it. If no winning run
-%%%exists in an abstract game even when the players are cooperating then there is
-%%%no winning run when the opponent is playing adversarily.
-
-%%%By checking for the existence of a winning run for the environment in an
-%%%abstract game tree constructed from a partial controller strategy we are
-%%%checking if the partial strategy is winning. If no environment-winning run
-%%%exists within the partial strategy then it is a winning partial strategy,
-%%%otherwise the spoiling run is a counterexample that proves that the partial
-%%%strategy is not winning. This forms the basis of a counterexample guided
-%%%abstraction refinement framework \cite{clarke2000} that operates on candidate strategies.
-
-\subsection{Counterexample-guided bounded synthesis}
+The bounded synthesis algorithm solves abstract game trees by invoking a SAT
+solver to search for candidate partial strategies. Candidate strategies are
+checked and by a call to the opponent solver, which also refines the abstract
+game tree. The refinements are then solved recursively.  The full procedure is
+illustrated in Algorithm~\ref{alg:bounded}.
 
@@
 
-Algorithm~\ref{alg:bounded} illustrates the bounded synthesis algorithm. The
-algorithm takes a concrete game $G$ as an implicit argument. In addition, it
+The algorithm takes a concrete game $G$ as an implicit argument. In addition it
 takes a player (controller or environment), state-round pair $\langle s,
 k\rangle$ and an abstract game tree $T$ and returns a winning partial strategy
-for it, if one exists.  The initial invocation of the algorithm takes the
-initial state $\langle I, 0\rangle$ and an empty abstract game tree
-$\emptyset$. Initially the solver is playing on behalf of the environment since
-that player takes the first move in every step.  The empty game tree does not
-constrain opponent moves, hence solving such an abstraction is equivalent to
-solving the original concrete game.  The algorithm is organised as a
-counterexample-guided abstraction refinement (CEGAR) loop.  The first step of
-the algorithm uses the \textsc{findCandidate} function, described in detail
-below, to come up with a candidate partial strategy for $T$. If it fails to
-find a strategy, this means that no winning partial strategy exists for $T$.
-If, on the other hand, a candidate partial strategy is found, we need to verify
-if it is indeed winning for $T$.
+for $T$, if one exists.  The environment takes the first move in each step of
+the game so the algorithm initially plays on the environment's behalf. The
+initial invocation takes the initial state $\langle I, n\rangle$ and an empty
+abstract game tree $\emptyset$. The empty game tree does not constrain opponent
+moves, hence solving such an abstraction is equivalent to solving the original
+concrete game.
+
+The algorithm is organised as a counterexample-guided abstraction refinement
+(CEGAR) loop.  The first step of the algorithm uses the \textsc{findCandidate}
+function, described in detail below, to come up with a candidate partial
+strategy for $T$. If it fails to find a strategy, this means that no winning
+partial strategy exists for $T$.  If, on the other hand, a candidate partial
+strategy is found, we need to verify if it is indeed winning for $T$.
 
@@
 The way game trees are refined is determined by the formula passed to the SAT
-solver in $\textsc{findCandidate}$. This formula encodes an abstract game tree
-into CNF by making copies of the transition relation for each game step in the
-tree.  The formula must also check whether the error state has been reached in
-no branch (\emph{respectively}: in all branches) for the controller
-(environment.) This procedure is defined in Algorithm \ref{alg:treeFormula} for
-both the controller in $\textsc{treeFormula}$ and the environment in
-$\textsc{\textoverline{treeFormula}}$.
+solver in $\textsc{findCandidate}$. This formula contains CNF encodings of all
+of the unrolled runs represented by an abstract game tree and the winning
+condition of the current player. Runs are encoded by copying the transition
+relation for every step in the abstract game. When playing for the controller,
+the SAT solver searches for a satisfying assignment to the unfixed label
+variables in tree so that none of the runs reaches the error state. The
+environment formulation is satisfiable if any run does reach the error state.
+The formula is constructed recursively from the root of a tree by
+$\textsc{treeFormula}$.
 
-Since the game tree formulation is passed to a SAT solver, any unfixed labels
-in the tree - both controllable and uncontrollable - will be existentially
-quantified. This means that the SAT solver will find any way to win the game
-while both players are cooperating. If no winning run exists in an abstract
-game even when the players are cooperating then there is no winning run when
-the opponent is playing adversarily. When a winning run is found, the actions
-chosen by the SAT solver are used to refine the game tree. This is advantageous
-for many synthesis problems where the game must be formalised as adversarial
-for correctness but the final implementation will be cooperating with its
-environment in the real world. An example of such a system is a device driver
-that cooperates with the device and OS to interface between the two.
+Since the game tree formulation is passed to a SAT solver, both controllable
+and uncontrollable unfixed labels will be existentially quantified. This means
+that the SAT solver will find any way to win the game while both players are
+cooperating. If no winning run exists in an abstract game even when the players
+are cooperating then there is no winning run when the opponent is playing
+adversarily. When a winning run is found, the actions chosen by the SAT solver
+are used to refine the game tree. This is advantageous for many synthesis
+problems where the game must be formalised as adversarial for correctness but
+the final implementation will be cooperating with its environment in the real
+world. An example of such a system is a device driver that cooperates with the
+device and OS to interface between the two.
 

### 2016-01-28 paper.tex

@@
 
-\subsection{Realisability of Bounded Safety Games}
+\subsection{Counterexample-guided bounded synthesis}
 
@@
 strategy for the game.
-
-\subsection{Abstract Game Trees}
 
@@
         \captionsetup{width=.8\textwidth}
-        \caption{AGT with strategy}
+        \caption{AGT with partial strategy}
         \label{fig:strategy}
-    \end{subfigure}%
-    \begin{subfigure}[t]{.25\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        node [left=4pt] {$\epsilon$}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    node [left=4pt] {$\beta$}
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] (sroot) {}
-                    child {node [below left=0 and -10pt] (sa) {}
-                        edge from parent [draw=none]
-                    }
-                    child {node [below right=0 and -10pt] (sb) {}
-                        edge from parent [draw=none]
-                    }
-                    node [left=4pt] {$\gamma$}
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [right=4pt] {$\alpha$}
-                node [above=4pt] {$\langle s, k \rangle$};
-                \draw[unfixed] (sa.center) -- (sb.center);
-                \draw[unfixed] (sroot.south) -- (sa.center);
-                \draw[unfixed] (sroot.south) -- (sb.center);
-        \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
-        \caption{AGT with spoiling strategy}
-        \label{fig:spoiling}
     \end{subfigure}%
@@
 Specifically, we consider abstractions represented as trees of actions,
-referred to as \emph{abstract game trees}.  Together with a state-round pair
-$\langle s,k\rangle$, an abstract game tree defines an \emph{abstract game}
-played from this state.  In the abstract game, the environment player is
-required to pick actions from the tree, starting from the root node.  After
-reaching a leaf, it continues playing unrestricted.  The tree in
-Figure~\ref{fig:agt} restricts the initial environment action to the set $\{a,
-d\}$.  After choosing action $d$, the environment reaches a leaf of the tree
-and continues playing unrestricted.  Alternatively, after choosing $a$, the
-environment is required to play action $b$ in the next round.
+referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an example abstract game
+tree for the controller (abstract game trees for the environment are similar).  In the abstract game, 
+the controller is required to pick actions
+from the tree, starting from the root node.  After reaching a leaf, it
+continues playing unrestricted.  The tree in Figure~\ref{fig:agt} restricts the
+initial controller action to the set $\{a, d\}$.  After choosing action $a$,
+the controller is required to play action $b$ in the next round.  It then reaches a leaf 
+of the tree and continues playing unrestricted.
+The root of the tree is annotated by the initial state $s$ of the abstract game and the 
+bound $k$ on the number of rounds. 
+We denote $\textsc{nodes}(T)$ the set of all nodes of a tree $T$ and
+$\textsc{leaves}(T)$ the subset of leaf nodes.
 
-Nodes of an abstract game tree are uniquely identified by the list of edge
-labels along the path from the root to the node.  We identify an abstract game
-tree with the set of its nodes.  For example, the tree in Figure~\ref{fig:agt}
-can be written as $\{(),(d),(a),(a,b)\}$.  We denote $\textsc{leaves}(T)$ the
-subset of leaf nodes of a tree $T$.
-
-A \emph{partial strategy} $Strat: T \rightarrow L_c $ assigns a controllable
-action to be played in each node of the abstract game tree.
-Figure~\ref{fig:strategy} shows an example partial strategy.  The controller
-starts by choosing action $\alpha$.  If the environment plays $a$, the
-controller responds with $\beta$ in the next round, and so on. Given a partial
+Given a controller (environment)
+abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow L_u$ 
+($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with the opponent's
+action to be played in this node.   
+Figure~\ref{fig:strategy} shows an example partial strategy.  The environment 
+starts by choosing action $\alpha$.  If the controller plays $a$, the
+environment responds with $\beta$ in the next round, and so on. Given a partial
 strategy $Strat$, we can map each leaf $l$ of the abstract game tree to
@@
 root to the leaf.
-
-\subsection{Counterexample-guided bounded synthesis}
 
@@
     \begin{algorithmic}[1]
-        \Function{solveAbstract}{$p, \langle s, k \rangle , T$}
-        \State $cand \gets $ \Call{findCandidate}{$p, \langle s, k \rangle, T$} \Comment{Look for a candidate}
-        \IIf{$k = n - 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
+        \Function{solveAbstract}{$p, s, k, T$}
+        \State $cand \gets $ \Call{findCandidate}{$p, s, k, T$} \Comment{Look for a candidate}
+        \IIf{$k = 1$} \Return $cand$ \EndIIf \Comment{Reached the bound}
         \State $T' \gets T$
@@
 
-The algorithm takes a concrete game $G$ as an implicit argument. In addition it
-takes a player (controller or environment), state-round pair $\langle s,
-k\rangle$ and an abstract game tree $T$ and returns a winning partial strategy
-for $T$, if one exists.  The environment takes the first move in each step of
-the game so the algorithm initially plays on the environment's behalf. The
-initial invocation takes the initial state $\langle I, n\rangle$ and an empty
-abstract game tree $\emptyset$. The empty game tree does not constrain opponent
-moves, hence solving such an abstraction is equivalent to solving the original
-concrete game.
+The algorithm takes a concrete game $G$ with bound $n$ as an implicit argument.
+In addition, it takes a player $p$ (controller or environment), state $s$,
+bound $k$ and an abstract game tree $T$ and returns a winning partial strategy
+for $opponent(p)$, if one exists.  The initial invocation of the algorithm
+takes the initial state $\langle I, n\rangle$ and an empty abstract game tree
+$\emptyset$.  Initially the solver is playing on behalf of the environment
+since that player takes the first move in every step.  The empty game tree does
+not constrain opponent moves, hence solving such an abstraction is equivalent
+to solving the original concrete game.
 
@@
 (CEGAR) loop.  The first step of the algorithm uses the \textsc{findCandidate}
-function, described in detail below, to come up with a candidate partial
-strategy for $T$. If it fails to find a strategy, this means that no winning
-partial strategy exists for $T$.  If, on the other hand, a candidate partial
-strategy is found, we need to verify if it is indeed winning for $T$.
+function, described below, to come up with a candidate partial strategy for
+$T$. If it fails to find a strategy, this means that no winning partial
+strategy exists for $T$.  If, on the other hand, a candidate partial strategy
+is found, we need to verify if it is indeed winning for $T$.
 
@@
 \textsc{solveAbstract} for the opponent. The dual solver solves games on behalf
-of the controller player.  {Figure~\ref{fig:spoiling}} shows a spoiling
-strategy discovered one of the leaves of the partial strategy.
+of the controller player.
 
@@
 The formula is constructed recursively from the root of a tree by
-$\textsc{treeFormula}$.
+$\textsc{treeFormula}$ (see Algorithm~\ref{alg:treeFormula}).
 
@@
 \end{algorithm}
-
-%%%The bounded synthesis algorithm (Algorithm \ref{alg:bounded}) begins with an
-%%%empty game as seen in Figure \ref{fig:emptytree}.  Initially we are playing on
-%%%behalf of the environment because it chooses the first move in each step.  The
-%%%empty game is passed to the SAT solver, which searches for a candidate
-%%%environment strategy.  If a candidate is found (Figure \ref{fig:candidate})
-%%%then it is checked for a spoiling strategy by solving for the controller.  If
-%%%no spoiling strategy exists, that means our candidate is a winning strategy and
-%%%the algorithm terminates.  Otherwise we find a counterexample (Figure
-%%%\ref{fig:cex}), which is used to refine the empty game tree to include the
-%%%first move from the controller's spoiling strategy (Figure \ref{fig:refined}).
-%%%The algorithm continues by finding a new candidate for the environment (Figure
-%%%\ref{fig:candidate2}), and a new counterexample to refine the game abstraction
-%%%once again (Figure \ref{fig:refined2}). The full procedure is listed in
-%%%Algorithm \ref{alg:bounded}.
 

### 2016-01-28 paper.tex
Fixed consistency of game trees in unbounded section

@@
     \centering
-    \begin{subfigure}[t]{.25\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
         \centering
@@
     \end{subfigure}%
-    \begin{subfigure}[t]{.25\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
         \centering
@@
 
-The unbounded synthesis algorithm (Algorithm \ref{alg:unbounded}) effectively
-consists of two communicating solvers: the counterexample-guided bounded
-reachability solver and the unbounded solver based on incremental induction.
-The unbounded solver calls the bounded solver with an increasing bound (line 8)
-while learning states from abstract game trees that either must or may lose for
-the controller. When the initial state is contained within the must-losing
-states the game is known to be unrealisable (line 5).
+The unbounded synthesis algorithm enhances the bounded method with
+computational learning. States are carefully recorded from unsatisfiable
+abstract games while maintaining an invariant ensuring the monotonicity of
+learning.
 
-May-losing states are learned by removing controller-winning states. We
-maintain a set of states for every rank up the current bound. The states we
-learn while executing the bounded algorithm can only said to be
-controller-winning for a particular rank or below. However, we construct each
-set carefully such that it is monotonically larger than the set of the rank
-below (see Figure \ref{fig:maylosing}) and so that the environment is unable to
-force play from one set to the next. Due to this careful construction, when two
-adjacent sets become equivalent we know that the algorithm has reached a fixed
-point and the controller is winning in the unbounded game (line 6).
-
-\begin{figure}
-    \centering
-    \begin{tikzpicture}
-        \node [minimum size=2cm,circle,draw] (a){$B^m[n]$};
-        \node [minimum size=1.5cm,circle,draw,right=1.8cm of a] (b){$B^m[n-1]$};
-        \node [minimum size=1cm,right=1.5cm of b] (dots){$\dots$};
-        \node [minimum size=0.5cm,circle,draw,right=1.5cm of dots] (c){$B^m[0]$};
-        \draw[-latex] (a.east) -- node [above] {$\supset$} (b.west);
-        \draw[-latex] (b.east) -- node [above] {$\supset$} (dots);
-        \draw[-latex] (dots.east) -- node [above] {$\supset$} (c.west);
-    \end{tikzpicture}
-    \caption{May-losing states}
-    \label{fig:maylosing}
-\end{figure}
+During the CEGAR loop of the bounded algorithm, each of the competing solvers
+searches for satisfying assignments of labels to abstract game trees. When the
+abstract game is unsatisfiable this indicates that the states represented by
+nodes in the abstract game tree are losing for the current player. We extract
+these states from the game tree using interpolation.
 
@@
 a winning candidate and can be extracted from the tree using interpolation.
+
+The states in an abstract game with no controller candidate are
+\emph{must-losing}. The controller will never be able to avoid the error state
+from these states.
+
+From abstractions with no environment candidate we record the complement states
+as \emph{may-losing}. We maintain a set of states for every rank up the current
+bound. The states we learn while executing the bounded algorithm can only said
+to be controller-winning for a particular rank or below. However, we construct
+each set carefully such that it is monotonically larger than the set of the
+rank below
+%(see Figure \ref{fig:maylosing})
+and so that the environment is unable to force play from one set to the next.
+Due to this careful construction, when two adjacent sets become equivalent we
+know that the algorithm has reached a fixed point and the controller is winning
+in the unbounded game (line 6).
+
+%%%\begin{figure}
+%%%    \centering
+%%%    \begin{tikzpicture}
+%%%        \node [minimum size=2cm,circle,draw] (a){$B^m[n]$};
+%%%        \node [minimum size=1.5cm,circle,draw,right=1.8cm of a] (b){$B^m[n-1]$};
+%%%        \node [minimum size=1cm,right=1.5cm of b] (dots){$\dots$};
+%%%        \node [minimum size=0.5cm,circle,draw,right=1.5cm of dots] (c){$B^m[0]$};
+%%%        \draw[-latex] (a.east) -- node [above] {$\supset$} (b.west);
+%%%        \draw[-latex] (b.east) -- node [above] {$\supset$} (dots);
+%%%        \draw[-latex] (dots.east) -- node [above] {$\supset$} (c.west);
+%%%    \end{tikzpicture}
+%%%    \caption{May-losing states}
+%%%    \label{fig:maylosing}
+%%%\end{figure}
 
@@
     \centering
+%%%    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+%%%        \node [circle,draw] (root){}
+%%%            child {node [circle,draw] {}
+%%%                child {node [circle,draw] {}
+%%%                    edge from parent [fixed] node [left] {$b$}
+%%%                }
+%%%                edge from parent [fixed] node [left] {$a$}
+%%%            }
+%%%            child {node [circle,draw] {}
+%%%                edge from parent [fixed] node [right] {$d$}
+%%%            }
+%%%            node [above=4pt] {$\langle s, k \rangle$};
+%%%    \end{tikzpicture}
     \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
         \node [] (root){}
-            child {node [circle,draw,label=left:$x_1$] {}
         child {node [circle,draw] {}
-                    child {node [circle,draw,label=left:$x_0^0$] {} edge from parent[unfixed]}
-                    edge from parent [fixed] node [left] {$u_1$}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$a$}
                 }
                 child {node [circle,draw] {} edge from parent[unfixed]
-                    child {node [circle,draw,label=right:$x_0^1$] {} edge from parent[unfixed]}
-                    edge from parent [fixed] node [right] {$u_2$}
+                    edge from parent [fixed] node [right] {$b$}
                 }
-                edge from parent [unfixed]
             };
@@
 Consider the snippet of a game tree in Figure \ref{fig:interpolatetree}. The
-tree is losing for the controller, the node labelled $x_1$ is at rank 1, and
-$x_0^0$ and $x_0^1$ are at rank 0. Since the tree is controller-losing we know
-that at least one run represented by the tree contains the error state.  As
-a result, $\textsc{treeFormula}(gt)$ is unsatisfiable. If we take the step
-$x_1$ to $x_0^0$ and cut it from the rest of the tree then
-$\textsc{treeFormula}(step) \land \textsc{treeFormula}(parent)$ must too be
-unsatisfiable.
+tree is losing for the controller, the topmost node is at rank 1, and the $a$ and
+$b$ nodes are at rank 0. Since the tree is controller-losing we know that at
+least one state in the tree will uncontrollably lead to the error state.  As a
+result, $\textsc{treeFormula}(T)$ is unsatisfiable. If we take a step of the
+tree, such as the step where the environment plays $a$, and cut it from the
+rest of the tree then $\textsc{treeFormula}(step) \land
+\textsc{treeFormula}(parent)$ must too be unsatisfiable.
 
-We can construct an interpolant with $A = \textsc{treeFormula}(parent)$ and $B
-= \textsc{treeFormula}(step)$. The only variables shared between $A$ and $B$
-are the state variables at $x_1$. We know that $B \land \mathcal{I}$ is
-unsatisfiable, therefore all states in $\mathcal{I}$ must lose to the
-uncontrollable label $u_1$. We also know that $A \to \mathcal{I}$, thus
+We can construct an interpolant with $F_1 = \textsc{treeFormula}(parent)$ and
+$F_2 = \textsc{treeFormula}(step)$. The only variables shared between $F_1$ and
+$F_2$ are the state variables at rank 1. We know that $F_2 \land \mathcal{I}$
+is unsatisfiable, therefore all states in $\mathcal{I}$ must lose to the
+uncontrollable action $a$. We also know that $F_1 \to \mathcal{I}$, thus
 $\mathcal{I}$ contains all states reachable by the parent tree (on runs that
@@
 
-Now we can consider the step $x_1$ to $x_0^1$, and the parent - now without
-either $x_1 \to x_0^0$ or $x_1 \to x_0^1$. The formula
+Now we can consider the step where the environment plays $b$, and the parent
+now without either step in the snippet. The formula
 $(\textsc{treeFormula}(parent) \land \mathcal{I}) \land
 \textsc{treeFormula}(step)$ must be unsatisfiable.  $\mathcal{I}$ contains all
-states that lose to $u_1$ so any other state reachable at $x_1$ must lose to
-$u_2$. Therefore we can compute another interpolant that contains states that
-lose to $u_2$.
+states that lose to $a$ so any other state reachable at rank 1 in this subtree
+must lose to $b$.  Therefore we can compute another interpolant that contains
+states that lose to $b$.
 
@@
 states that sets $B^m[i]$ grows monotonically with $i$ and that each $B^m[i+1]$
-overapproximates the uncontrollable predecessor of $f^m[i]$: $$\forall
+overapproximates the uncontrollable predecessor of $B^m[i]$: $$\forall
 i<k.~B^m[i] \subseteq B^m[i+1], Upre(B^m[i]) \subseteq B^m[i+1].$$

### 2016-01-28 paper.tex
Added benchmark plots for us and demiurge

@@
 
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[y=0.04cm, x=0.15cm]
+        %axis
+        \draw (0,0) -- coordinate (x axis mid) (60,0);
+            \draw (0,0) -- coordinate (y axis mid) (0,150);
+            %ticks
+            \foreach \x in {0,10,...,60}
+                \draw (\x,1pt) -- (\x,-3pt)
+                node[anchor=north] {\x};
+            \foreach \y in {0,25,...,150}
+                \draw (1pt,\y) -- (-3pt,\y) 
+                    node[anchor=east] {\y}; 
+        %labels      
+        \node[below=0.8cm] at (x axis mid) {Time (min)};
+        \node[rotate=90, above left=1cm and 0.7cm] at (y axis mid) {Instances solved};
+
+        %Plots
+        \draw plot[mark=x, mark options={fill=white}] file {bench_termite.dat};
+        \draw plot[mark=+, mark options={fill=white}] file {bench_demiurge.dat};
+
+        \begin{scope}[shift={(30,20)}] 
+            \draw (0,0) -- 
+                plot[mark=x, mark options={fill=white}] (0.25,0) -- (0.5,0) 
+                node[right]{Unbounded synthesis};
+            \draw[yshift=\baselineskip] (0,0) -- 
+                plot[mark=+, mark options={fill=white}] (0.25,0) -- (0.5,0)
+                node[right]{Demiurge};
+        \end{scope}
+
+    \end{tikzpicture}
+\end{figure}
+
 The implementation of our algorithm uses \textsc{Glucose}~\cite{audemard2014}
@@
 for one hour to solve an instance.
-
-INSERT GRAPHS
 

### 2016-01-28 paper.tex
Editing pass

@@
 representing states symbolically.  Historically the most successful technique
-has been to use \emph{Binary Decision Diagrams} (BDDs).  BDDs efficiently
+has been to use \emph{Binary Decision Diagrams}.  BDDs efficiently
 represent a relation on a set of game variables but in the worst case the
-representation may be exponential in the number of variables. This means that
+representation may be exponential in the number of variables. As a result
 BDDs are not a one-size-fits-all solution for all reactive synthesis
@@
 to outperform BDDs on certain classes of game specifications but is only able
-to find counterexamples of a certain length to the safety property.  In this
+to find counterexamples to the safety property of a certain length.  In this
 paper, we extend this idea to unbounded games by approximating sets of unsafe
@@
 
-Realisability for a bounded safety game can be solved by checking an
-quantified formula consisting of the game's transition relation \emph{unrolled}
-to the length of the bounded game. The quantifiers represent the two players of
-the game: universal for the controller and existential for the environment.
-Just as the players alternate in choosing actions every step of the game, there
-must be quantifier alternations in the formula for each unrolling.
+Realisability for a bounded safety game can be solved by checking a quantified
+formula consisting of the game's transition relation \emph{unrolled} to the
+length of the bounded game. The quantifiers represent the two players of the
+game: universal for the controller and existential for the environment.  Just
+as the players alternate in choosing actions every step of the game, the
+formula is prefixed by a quantifier alternation over $L_u$ and $L_c$ for each
+step.
 
@@
 finding counterexamples. Eventually one solver will be unable to find a
-counterexample indicating that its opponent's candidate strategy is a winning
+counterexample, indicating that its opponent's candidate strategy is a winning
 strategy for the game.
@@
 abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow L_u$ 
-($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with the opponent's
-action to be played in this node.   
-Figure~\ref{fig:strategy} shows an example partial strategy.  The environment 
-starts by choosing action $\alpha$.  If the controller plays $a$, the
-environment responds with $\beta$ in the next round, and so on. Given a partial
-strategy $Strat$, we can map each leaf $l$ of the abstract game tree to
-$\langle s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained
-by playing all controllable and uncontrollable actions on the path from the
-root to the leaf.
+($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with
+the opponent's action to be played in that node.   Figure~\ref{fig:strategy}
+shows an example partial strategy.  The environment starts by choosing action
+$\alpha$.  If the controller plays $a$, the environment responds with $\beta$
+in the next round, and so on. Given a partial strategy $Strat$, we can map each
+leaf $l$ of the abstract game tree to $\langle
+s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
+playing all controllable and uncontrollable actions on the path from the root
+to the leaf.
 
@@
 
-The algorithm takes a concrete game $G$ with bound $n$ as an implicit argument.
-In addition, it takes a player $p$ (controller or environment), state $s$,
-bound $k$ and an abstract game tree $T$ and returns a winning partial strategy
-for $opponent(p)$, if one exists.  The initial invocation of the algorithm
-takes the initial state $\langle I, n\rangle$ and an empty abstract game tree
-$\emptyset$.  Initially the solver is playing on behalf of the environment
-since that player takes the first move in every step.  The empty game tree does
-not constrain opponent moves, hence solving such an abstraction is equivalent
-to solving the original concrete game.
+The algorithm takes a concrete game $G$ with maximum bound $n$ as an implicit
+argument.  In addition, it takes a player $p$ (controller or environment),
+state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
+partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
+invocation of the algorithm takes the initial state $\langle I, n\rangle$ and
+an empty abstract game tree $\emptyset$.  Initially the solver is playing on
+behalf of the environment since that player takes the first move in every step.
+The empty game tree does not constrain opponent moves, hence solving such an
+abstraction is equivalent to solving the original concrete game.
 
@@
 
-The way game trees are refined is determined by the formula passed to the SAT
-solver in $\textsc{findCandidate}$. This formula contains CNF encodings of all
-of the unrolled runs represented by an abstract game tree and the winning
-condition of the current player. Runs are encoded by copying the transition
-relation for every step in the abstract game. When playing for the controller,
-the SAT solver searches for a satisfying assignment to the unfixed label
-variables in tree so that none of the runs reaches the error state. The
-environment formulation is satisfiable if any run does reach the error state.
-The formula is constructed recursively from the root of a tree by
-$\textsc{treeFormula}$ (see Algorithm~\ref{alg:treeFormula}).
+Candidates are discovered by passing a formulation of the abstract game tree to
+a SAT solver in $\textsc{findCandidate}$. This formula contains CNF encodings
+of all of the unrolled runs represented by the tree and the winning condition
+of the current player.  Runs are encoded by copying the transition relation for
+every step in the abstract game. When playing for the controller, the SAT
+solver searches for a satisfying assignment to the unfixed label variables in
+tree so that none of the runs reaches the error state. The environment
+formulation is satisfiable if any run does reach the error state.  The formula
+is constructed recursively from the root of a tree by $\textsc{treeFormula}$
+(see Algorithm~\ref{alg:treeFormula}).
 
@@
 problems where the game must be formalised as adversarial for correctness but
-the final implementation will be cooperating with its environment in the real
-world. An example of such a system is a device driver that cooperates with the
-device and OS to interface between the two.
+the final implementation will cooperate with its environment in the real world.
+An example of such a system is a device driver that cooperates with the device
+and OS to provide the interface between the two.
 
@@
 
-A similar concept can be applied to the bounded synthesis algorithm to
-iteratively increase the bound of the game and terminate when a fixed point is
-reached. When a strategy is found to be winning on an abstract game tree, we
-record as winning the states from which the opponent could find no
-counterexample. To find these states we use interpolation of subformulas of the
-game tree.
+We apply a similar fixed point computation on game states to the bounded
+synthesis algorithm in order to promote it to unbounded games. Our unbounded
+procedure iteratively executes the bounded algorithm while recording states
+from abstract game trees. By carefully managing the way states are recorded we
+guarantee that a fixed point proves the controller's strategy correct.
 
@@
 The states in an abstract game with no controller candidate are
-\emph{must-losing}. The controller will never be able to avoid the error state
-from these states.
+\emph{must-losing}. The environment can always force the game into the error
+states from these states.
 
-From abstractions with no environment candidate we record the complement states
-as \emph{may-losing}. We maintain a set of states for every rank up the current
-bound. The states we learn while executing the bounded algorithm can only said
-to be controller-winning for a particular rank or below. However, we construct
-each set carefully such that it is monotonically larger than the set of the
-rank below
-%(see Figure \ref{fig:maylosing})
-and so that the environment is unable to force play from one set to the next.
-Due to this careful construction, when two adjacent sets become equivalent we
-know that the algorithm has reached a fixed point and the controller is winning
-in the unbounded game (line 6).
+From abstractions with no environment candidate we record the complement of
+states in the tree as \emph{may-losing}. The environment cannot reach the error
+state in a number of steps equal to the distance to the bottom of the tree.
+
+We maintain a set of states for each rank up to the current bound. We maintain
+an invariant over these sets via careful construction so that they are
+monotonically increasing by rank. We also ensure that the environment is unable
+to force play from one set to the next. Due to these invariants, when two
+adjacent sets become equivalent we know that the algorithm has reached a fixed
+point and the controller is winning in the unbounded game (line 6).
 
@@
 
-Given two formulas $A$ and $B$ such that $A \land B$ is unsatisfiable, it is
-possible to construct a Craig interpolant\cite{craig1957} $\mathcal{I}$ such
-that $A \to \mathcal{I}$, $B \land \mathcal{I}$ is unsatisfiable, and
-$\mathcal{I}$ refers only to the intersection of variables in $A$ and $B$.  An
-interpolant can be constructed efficiently from a resolution proof of the
-unsatisfiability of $A \land B$ \cite{pudlak1997}.
+Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
+it is possible to construct a Craig interpolant\cite{craig1957} $\mathcal{I}$
+such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
+$\mathcal{I}$ refers only to the intersection of variables in $F_1$ and $F_2$.
+An interpolant can be constructed efficiently from a resolution proof of the
+unsatisfiability of $F_1 \land F_2$ \cite{pudlak1997}.
 
@@
 by removing a single step on each iteration (Algorithm \ref{alg:learn}). All
-learned states from which the controller \emph{must} lose are recorded in a set
-of \emph{bad} states $B^M$.  This algorithm can also be performed on
-environment-losing trees with the caveat that any state learnt at a node of
-rank $n$ is only known to lose at ranks less than or equal to $n$. We negate
-the interpolants we learn this way and record them in a mapping of ranks to
-sets of states $B^m$ which \emph{may} be bad states for the controller.
+learned states from which the controller must lose are recorded in a set of bad
+states $B^M$. Environment-losing are learned in the same way but are recorded
+by removing the state from a set of sets of controller may-losing states $B^m$.
+An environment-losing state of rank $n$ is removed from every set at rank $n$
+and lower.
 
@@
 Line~(11) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
-1.  Consider formulas $f_1=\sigma(X_T) \land
+1.  Consider formulas $F_1=\sigma(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ and $F_2 =
@@
 
-\subsection{Choosing sensible opponent moves}
+\subsection{Choosing opponent moves}
 

### 2016-01-28 paper.tex
Added syntcomp results to evaluation and wrote some analysis

@@
         %labels      
-        \node[below=0.8cm] at (x axis mid) {Time (min)};
+        \node[below=0.8cm] at (x axis mid) {Time (minutes)};
         \node[rotate=90, above left=1cm and 0.7cm] at (y axis mid) {Instances solved};
@@
 for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
-generation. We compared our implementation to two solvers that competed in the
-sequential realisability track of SYNTCOMP'15.  \textsc{Simple BDD
-Solver}~\cite{walker2014} because it had the best results in the track, and
-\textsc{Demiurge}~\cite{bloem2014} because it was the only SAT-based tool to
-compete. The benchmarks were run on a cluster of Intel Quad Core Xeon E5405
-CPUs with 16GB of memory. Each solver was allowed exclusive access to a node
-for one hour to solve an instance.
+generation. We compared our implementation to
+\textsc{Demiurge}~\cite{bloem2014}, the only SAT-based solver to compete in the
+sequential realisability track of SYNTCOMP'15. The benchmarks were run on a
+cluster of Intel Quad Core Xeon E5405 CPUs with 16GB of memory.  Each solver
+was allowed exclusive access to a node for one hour to solve an instance.
 
 Our implementation was able to solver 103 out of the 250 specification in the
-alloted time, including 10 instances that were unable to be solved by any other
-solver including all other solvers in SYNTCOMP'15. \textsc{Simple BDD Solver}
-and \textsc{Demiurge} were able to solve X and Y respectively. Something about
-us vs BDDs and us vs demiurge with numbers of specs.
+alloted time, including 12 instances that were unable to be solved by any other
+solver in SYNTCOMP'15. \textsc{Demiurge} was able to solve 136 specifications.
+77 of those instances were also solved by our implementation, so
+\textsc{Demiurge} solved 59 instances that our solver did not. However, our
+method was able to solve 26 specifications not solved by \textsc{Demiurge}.
 
-Four of the previously unsolved instances are device driver instances and
-another five are toy examples. This supports the hypothesis that different game
-solving methodologies perform better on certain classes of specifications.
+For reference, the results of the synthesis competition are shown in
+Figure~\ref{fig:syntcomp}. The competition ran on better hardware with a longer
+time period, which explains why \textsc{Demiurge} performed better.
+
+\begin{figure}
+    \centering
+    \setlength{\tabcolsep}{8pt}
+    \begin{tabular}{l | r | r }
+        \textbf{Solver} & \textbf{Solved} & \textbf{Unique} \\
+        \hline
+        Simple BDD Solver (2) & 195 & 10 \\
+        AbsSynthe (seq2) & 187 & 2 \\
+        Simple BDD Solver (1) & 185 & \\
+        AbsSynthe (seq3) & 179 & \\
+        Realizer (sequential) & 179 & \\
+        AbsSynthe (seq1) & 173 & 1 \\
+        Demiurge (D1real) & 139 & 5 \\
+        Aisy & 98 & \\
+        \hline
+        Unbounded Synthesis & 103 & 12 \\
+    \end{tabular}
+    \caption{Synthesis Competition 2015 Results}
+    \label{fig:syntcomp}
+\end{figure}
+
+Five of the instances unique to our solver are device driver instances and
+another five are from the \texttt{stay} family, which are classified as toy
+instances.  Out of the 26 instances that \textsc{Demiurge} could not solve,
+there were only 6 different benchmark classes. This supports the hypothesis
+that different game solving methodologies perform better on certain classes of
+specifications.
+
+Whilst our solver is unable to solve as many instances as other tools yet, it
+was able to solve more unique instances than any solver in the competition.
+This shows that the method has potential. In the parallel track,
+\textsc{Demiurge} competes with a suite of 3 separate but communicating
+solvers. The solvers relay unsafe states to one another, which is compatible
+with the set $B^M$ in our solver. It remains future work to explore the
+possibility of combining these solvers.
 

### 2016-01-28 paper.tex
Added a conclusion

@@
 
+We presented an extension to an existing bounded synthesis technique that
+promotes it to unbounded safety games. The approach taken as whole differs
+other synthesis techniques by operating in a counterexample guided abstraction
+refinement loop instead of attempting to construct a winning region. The
+results show that our approach is useful for solving some classes of games that
+have never previously been solved.
+
+Future work includes incorporating the solver into a parallel suite of
+communicating solver~\cite{bloem2014}. We also aim experiment with
+optimisations by focusing on heuristics for avoiding unhelpful actions as well
+as constructing more succinct representations of interpolants.
+
 \bibliographystyle{splncs03}

### 2016-01-28 paper.tex
worked on the intro

--- 2016-01-28 paper.tex
@@
 \emph{correct by construction} controller for a reactive system from a
-specification.  However, the reactive synthesis problem is 2EXPTIME-complete so
-naive algorithms are infeasible on anything but simple systems.
-
-Reactive synthesis is formalised as a game between the \emph{controller} and
+specification.  Reactive synthesis is formalised as a game between the \emph{controller} and
 its \emph{environment}. In this work we focus on safety games, in which the
 controller must prevent the environment from forcing the game into an error
-state. In this way the environment is equivalent to an existential search of
-the game and the controller is universal. Much of the complexity of reactive
-synthesis stems from managing the interactions of the alternating quantifiers
-on the state space of the game.
+state.
 
+The reactive synthesis problem is 2EXPTIME-complete so
+naive algorithms are infeasible on anything but simple systems.
+% In this way the environment is equivalent to an existential search of
+%the game and the controller is universal. Much of the complexity of reactive
+%synthesis stems from managing the interactions of the alternating quantifiers
+%on the state space of the game.
 There are several techniques that aim to mitigate this complexity by
-representing states symbolically.  Historically the most successful technique
-has been to use \emph{Binary Decision Diagrams}.  BDDs efficiently
-represent a relation on a set of game variables but in the worst case the
-representation may be exponential in the number of variables. As a result
-BDDs are not a one-size-fits-all solution for all reactive synthesis
-specifications.
+representing states and transitions of the system symbolically~\cite{Piterman_PS_06,bloem2014,morgenstern2013}. 
+These techniques incrementally construct a symbolic representation of the set of 
+discovered winning states of the game.  The downside of this approach is that 
+keeping track of all discovered winning states can lead to a space explosion even
+when using efficient symbolic representation such as BDDs or CNFs.
 
-Advances in SAT solving technology has prompted research into its applicability
-to synthesis as an alternative to BDDs. One approach is to find sets of states
-in CNF \cite{bloem2014,morgenstern2013}. Another approach is to eschew states
-and focus on \emph{runs} of the game. Previous work has applied this idea to
-realizability of bounded games \cite{narodytska2014} by forming abstractions of
-the game and refining in a counterexample-guided framework. This has been shown
-to outperform BDDs on certain classes of game specifications but is only able
-to find counterexamples to the safety property of a certain length.  In this
-paper, we extend this idea to unbounded games by approximating sets of unsafe
-states from abstract games. Careful construction ensures that a fixed point in
-these sets guarantees completeness.
+%Historically the most successful technique
+%has been to use \emph{Binary Decision Diagrams}.  BDDs efficiently
+%represent a relation on a set of game variables but in the worst case the
+%representation may be exponential in the number of variables. As a result
+%BDDs are not a one-size-fits-all solution for all reactive synthesis
+%specifications.
+
+%Advances in SAT solving technology has prompted research into its applicability
+%to synthesis as an alternative to BDDs. One approach is to find sets of states
+%in CNF \cite{bloem2014,morgenstern2013}.
+
+An alternative approach, proposed by Narodytska et al.~\cite{narodytska2014},
+is to eschew states and focus on \emph{runs} of the game.  The method works by
+exploring a subset of the concrete runs of the game and proving that these runs can be
+generalised into a winning strategy on behalf of one of the players. In contrast to 
+other existing synthesis methods, it does not store,
+in either symbolic or explicit form, the set of winning states.
+Instead, it uses counterexample-guided backtracking search to identify a small subset
+of runs that are sufficient to solve the game.
+
+This method has been shown to outperform BDDs on certain classes of games; however it
+suffers from an important limitation: it is only able to solve games with a bounded number
+of rounds.  In case of safety games, this means proving that the controller 
+can keep the game within the safe region for a bounded number of steps.  
+This is insufficient in most practical situations that require unbounded realizability.
+
+%Previous work has applied this idea to
+%realizability of bounded games \cite{narodytska2014} by forming abstractions of
+%the game and refining in a counterexample-guided framework.
+
+In this paper, we extend the method by Narodytska et al.~\cite{narodytska2014} to unbounded 
+safety games.  To this end we enhance the method with computational learning: every time
+the search algorithm discovers a new counterexample, the learning procedure analyzes the 
+counterexample, extracting a subset of states winning for one of the players from it.  
+The learning procedure ensures that reaching a fixed point in these sets is sufficient to establish 
+unbounded realizability.
+Our method can be seen as a hybrid between the counterexample-guided algorithm by
+Narodytska et al. and methods based on winning set computation: we use the former to guide 
+the search, while we rely on the latter to ensure convergence.
+
+We evaluate our method on the benchmarks of the 2015 synthesis competition
+(SYNTCOMP'15).  While our solver solves fewer total instances than competitors,
+it solves the largest number of unique instances, i.e., instances that could not be solved by 
+any other solver.  These results confirm that there exist classes of problems that are
+hard for traditional synthesis techniques, but can be efficiently solved by our method.
+While further performance improvements are clearly needed, in its current state the method 
+can be used in a portfolio solver to increase the number of tractable instances.
 
@@
             \Else
-                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{gtNodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
             \EndIf

### 2016-01-28 paper.tex
Small edits

--- 2016-01-28 paper.tex
@@
 
+We compile interpolants into DNF cubes by constructing a temporary BDD. This
+smaller representation is necessary for maintaining the size of formulas passed
+to the SAT solver. However, it can occaisionally lead to state explosion.
+
 \subsection{Correctness}
@@
 
+We presented an extension to an existing bounded synthesis technique that
+promotes it to unbounded safety games. The approach taken as whole differs from
+other synthesis techniques by combining counterexample guided game solving with
+winning set computation. The results show that our approach is able to solve
+more unique instances than other solvers by performing well on certain classes
+of games that are hard for other methodologies.
+
+Future work includes incorporating the solver into a parallel suite of
+communicating solvers~\cite{bloem2014}. There is evidence that different
+solvers perform well on different classes of games. Thus we hypothesise that
+the way forward for synthesis tools is to combine the efforts of many different
+techniques.
+
+We also aim to experiment with optimisations by focusing on heuristics for
+avoiding unhelpful actions as well as constructing succinct representations of
+interpolants without resorting to BDDs.
+
 \bibliographystyle{splncs03}

### 2016-01-29 paper.tex
Edited SAT formula explanation

@@
         \Function{findCandidate}{$p, s, k, T$}
-            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$T$}}{\Call{\textoverline{treeFormula}}{$T$}} \EndIfElse
-            \State $sol \gets $ \Call{SAT}{$s(X_T) \land f$}
+        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
+            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
+            \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
             \If{$sol = \texttt{unsat}$} 
@@
                     \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
-                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, T$}}{\Call{\textoverline{learn}}{$\sigma, T$}} \EndIfElse
+                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, \hat{T}$}}{\Call{\textoverline{learn}}{$\sigma, \hat{T}$}} \EndIfElse
                 \EndIf
@@
 
-Candidates are discovered by passing a formulation of the abstract game tree to
-a SAT solver in $\textsc{findCandidate}$. This formula contains CNF encodings
-of all of the unrolled runs represented by the tree and the winning condition
-of the current player.  Runs are encoded by copying the transition relation for
-every step in the abstract game. When playing for the controller, the SAT
-solver searches for a satisfying assignment to the unfixed label variables in
-tree so that none of the runs reaches the error state. The environment
-formulation is satisfiable if any run does reach the error state.  The formula
-is constructed recursively from the root of a tree by $\textsc{treeFormula}$
-(see Algorithm~\ref{alg:treeFormula}).
-
-Since the game tree formulation is passed to a SAT solver, both controllable
-and uncontrollable unfixed labels will be existentially quantified. This means
-that the SAT solver will find any way to win the game while both players are
-cooperating. If no winning run exists in an abstract game even when the players
-are cooperating then there is no winning run when the opponent is playing
-adversarily. When a winning run is found, the actions chosen by the SAT solver
-are used to refine the game tree. This is advantageous for many synthesis
-problems where the game must be formalised as adversarial for correctness but
-the final implementation will cooperate with its environment in the real world.
-An example of such a system is a device driver that cooperates with the device
-and OS to provide the interface between the two.
+The CEGAR loop depends on the ability to guess candidate partial strategies in
+\textsc{findCandidate}. For this purpose we use the heuristic that a partial
+strategy may be winning if for each \textsc{outcome} in the tree there exists a
+play of controller and environment actions that is winning for the current
+player. We can formulate this heuristic as a SAT query, which is constructed
+recursively by $\textsc{treeFormula}$ ($\textsc{\textoverline{treeFormula}}$)
+in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
+bound with \emph{unfixed} actions that allow existential choice by the SAT
+solver. Then each game round in the tree is encoded in CNF by making copies of
+the transition relation along with the winning condition of the current player.
+
+The heuristic is based on the observation that for many synthesis problems in
+the real world the final implementation will cooperate with its environment.
+The existential search of the SAT solver simulates cooperation between the two
+players.
+
+%%%Candidates are discovered by passing a formulation of the abstract game tree to
+%%%a SAT solver in $\textsc{findCandidate}$. This formula contains CNF encodings
+%%%of all of the unrolled runs represented by the tree and the winning condition
+%%%of the current player.  Runs are encoded by copying the transition relation for
+%%%every step in the abstract game. When playing for the controller, the SAT
+%%%solver searches for a satisfying assignment to the unfixed label variables in
+%%%tree so that none of the runs reaches the error state. The environment
+%%%formulation is satisfiable if any run does reach the error state.  The formula
+%%%is constructed recursively from the root of a tree by $\textsc{treeFormula}$
+%%%(see Algorithm~\ref{alg:treeFormula}).
+
+%%%Since the game tree formulation is passed to a SAT solver, both controllable
+%%%and uncontrollable unfixed labels will be existentially quantified. This means
+%%%that the SAT solver will find any way to win the game while both players are
+%%%cooperating. If no winning run exists in an abstract game even when the players
+%%%are cooperating then there is no winning run when the opponent is playing
+%%%adversarily. When a winning run is found, the actions chosen by the SAT solver
+%%%are used to refine the game tree. This is advantageous for many synthesis
+%%%problems where the game must be formalised as adversarial for correctness but
+%%%the final implementation will cooperate with its environment in the real world.
+%%%An example of such a system is a device driver that cooperates with the device
+%%%and OS to provide the interface between the two.
 
@@
 
-When executing the unbounded solver, lines 18 and 19 become active in the
+When executing the unbounded solver, lines 19 and 20 become active in the
 bounded solver. These lines call the learning procedures when the solver fails
@@
 
-\subsection{Choosing opponent moves}
-
-One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
-is that it mimics real-world uses of synthesis by allowing cooperation between
-system and environment. Unfortunately, allowing the opponents to pick moves for
-one another can frequently backfire. It is common when modelling real-world
-systems to abstract over out of scope failures with transitions that
-immediately determine the outcome of the game. For example, in a network driver
-a request to send a packet may fail due to a disconnected wire, which might be
-modelled as an environment controlled transition to a system-winning state.
-
-This slows down the synthesis process as all erroneous transition are
-accumulated into the abstract game tree. To mitigate this effect we have a
-heuristic to guess non-erroneous labels as \emph{default moves} that are used
-instead of allowing the opponent to choose. This does not effect the
-correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
-is \texttt{unsat} with default moves then those moves would eventually have
-been added to the abstraction in a refinement step. If the formula is
-\texttt{unsat} even with the opponent choosing moves then it must be
-\texttt{unsat} with default moves.
+%%%\subsection{Choosing opponent moves}
+
+%%%One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
+%%%is that it mimics real-world uses of synthesis by allowing cooperation between
+%%%system and environment. Unfortunately, allowing the opponents to pick moves for
+%%%one another can frequently backfire. It is common when modelling real-world
+%%%systems to abstract over out of scope failures with transitions that
+%%%immediately determine the outcome of the game. For example, in a network driver
+%%%a request to send a packet may fail due to a disconnected wire, which might be
+%%%modelled as an environment controlled transition to a system-winning state.
+
+%%%This slows down the synthesis process as all erroneous transition are
+%%%accumulated into the abstract game tree. To mitigate this effect we have a
+%%%heuristic to guess non-erroneous labels as \emph{default moves} that are used
+%%%instead of allowing the opponent to choose. This does not effect the
+%%%correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
+%%%is \texttt{unsat} with default moves then those moves would eventually have
+%%%been added to the abstraction in a refinement step. If the formula is
+%%%\texttt{unsat} even with the opponent choosing moves then it must be
+%%%\texttt{unsat} with default moves.
 

### 2016-01-29 paper.tex
tried to explain treeformula's better

@@
 \textsc{findCandidate}. For this purpose we use the heuristic that a partial
-strategy may be winning if for each \textsc{outcome} in the tree there exists a
-play of controller and environment actions that is winning for the current
-player. We can formulate this heuristic as a SAT query, which is constructed
-recursively by $\textsc{treeFormula}$ ($\textsc{\textoverline{treeFormula}}$)
-in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
-bound with \emph{unfixed} actions that allow existential choice by the SAT
-solver. Then each game round in the tree is encoded in CNF by making copies of
-the transition relation along with the winning condition of the current player.
-
-The heuristic is based on the observation that for many synthesis problems in
-the real world the final implementation will cooperate with its environment.
-The existential search of the SAT solver simulates cooperation between the two
-players.
+strategy may be winning if each \textsc{outcome} of the strategy can be extended to a
+run of the game that is winning for the current player.  Clearly, if such a partial 
+strategy does not exist then no winning partial strategy can exist for the abstract 
+game tree.  
+We can formulate this heuristic as a SAT query, which is constructed
+recursively by $\textsc{treeFormula}$ (for the controller) or $\textsc{\textoverline{treeFormula}}$
+(for the environment) in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
+bound with edges that are labeled with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown in Figure~\ref{}.  
+Then each edge in the tree is encoded in CNF by making copies of the transition relation along with the 
+winning condition.
+
+%The heuristic is based on the observation that for many synthesis problems in
+%the real world the final implementation will cooperate with its environment.
+%The existential search of the SAT solver simulates cooperation between the two
+%players.
 

### 2016-01-29 paper.tex
More edits to section 2

@@
 
-A \emph{safety game}, $G = \langle X, L_u, L_c, \delta, I, E \rangle$,
-is defined over boolean state variables $X$, uncontrollable label variables $L_u$, and
-controllable label variables $L_c$.  $I$ is the initial state of the game given as a 
-valuation of state variables.  $E(X)$ is the set of error states represented by its 
-characteristic formula.  The transition relation $\delta(X, L_u, L_c, X')$ of the game 
-is a boolean formula that relates current state and label to the set of possible next 
-states of the game.  We assume deterministic games, where 
-$\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies (x'_1=x'_2)$.
-
-At every round of the game, the \emph{environment} picks an
-uncontrollable label, the \emph{controller} responds by choosing a controllable 
-label and the game transitions into a new state according to $\delta$. 
-A \emph{run} of a game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n,
-c_n)$ is a chain of state and label pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$. 
-A run is winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot E(x_i))$. 
-In a \emph{bounded game} with bound $n$ all runs are restricted to length $n$, whereas unbounded 
-games consider runs of infinite length. Since we consider only deterministic games, a run 
-is uniquely described by a list of assignments to $L_u$ and $L_c$.
-
-A \emph{controller strategy} $\pi^c : (X, L_u) \to L_c$ is a mapping of states
-and uncontrollable inputs to controllable labels. A controller strategy is
-winning in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)),
-(x_1, u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
+A \emph{safety game}, $G = \langle X, U, C, \delta, I, E \rangle$, is defined
+over boolean state variables $X$, uncontrollable action variables $U$, and
+controllable action variables $C$.  $I$ is the initial state of the game given
+as a valuation of state variables.  $E(X)$ is the set of error states
+represented by its characteristic formula.  The transition relation $\delta(X,
+U, C, X')$ of the game is a boolean formula that relates current state and
+action to the set of possible next states of the game.  We assume deterministic
+games, where $\delta(x,u,c,x'_1) \land \delta(x,u,c,x'_2) \implies
+(x'_1=x'_2)$.
+
+At every round of the game, the \emph{environment} picks an uncontrollable
+action, the \emph{controller} responds by choosing a controllable action and
+the game transitions into a new state according to $\delta$.  A \emph{run} of a
+game $(x_0, u_0, c_0), (x_1, u_1, c_1) \dots (x_n, u_n, c_n)$ is a chain of
+state and action pairs s.t.\,$\delta(x_k, u_k, c_k, x_{k+1})$.  A run is
+winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot
+E(x_i))$.  In a \emph{bounded game} with bound $\kappa$ all runs are restricted
+to length $\kappa$, whereas unbounded games consider runs of infinite length.
+Since we consider only deterministic games, a run is uniquely described by a
+list of assignments to $U$ and $C$.
+
+A \emph{controller strategy} $\pi^c : (X, U) \to C$ is a mapping of states and
+uncontrollable inputs to controllable actions. A controller strategy is winning
+in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)), (x_1,
+u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
 Bounded \emph{realizability} is the problem of determining the existence of
@@
 
-An \emph{environment strategy} $\pi^e : X \to L_u$ is a mapping of states to
-uncontrollable labels. A bounded run is winning for the environment if $x_0
-= I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is
-winning for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1,
-\pi^e(x_1), c_1) \dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.
-Safety games are zero sum, therefore the existence of a winning controller strategy
-implies the nonexistence of a winning environment strategy and vice versa.
+An \emph{environment strategy} $\pi^e : X \to U$ is a mapping of states to
+uncontrollable actions. A bounded run is winning for the environment if $x_0 =
+I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is winning
+for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1, \pi^e(x_1), c_1)
+\dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.  Safety games
+are zero sum, therefore the existence of a winning controller strategy implies
+the nonexistence of a winning environment strategy and vice versa.
 
@@
 as the players alternate in choosing actions every step of the game, the
-formula is prefixed by a quantifier alternation over $L_u$ and $L_c$ for each
+formula is prefixed by a quantifier alternation over $U$ and $C$ for each
 step.
@@
     \centering
-    \begin{subfigure}[t]{.32\textwidth}
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
     \end{subfigure}%
-    \begin{subfigure}[t]{.32\textwidth}
+    \begin{subfigure}[t]{.25\textwidth}
         \centering
@@
         \label{fig:refined}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.25\textwidth}
+        \centering
+        \begin{tikzpicture}[scale=0.75,dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [left] {$c$}
+                        }
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [left] {$f$}
+                        }
+                        edge from parent [fixed] node [right] {$e$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \captionsetup{width=.8\textwidth}
+        \caption{Extended AGT}
+        \label{fig:extended}
     \end{subfigure}
@@
 Given a controller (environment)
-abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow L_u$ 
-($Strat: \textsc{nodes}(T) \rightarrow L_c$) labels each node of the tree with
+abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ 
+($Strat: \textsc{nodes}(T) \rightarrow C$) labels each node of the tree with
 the opponent's action to be played in that node.   Figure~\ref{fig:strategy}
@@
 
-The algorithm takes a concrete game $G$ with maximum bound $n$ as an implicit
+The algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
 argument.  In addition, it takes a player $p$ (controller or environment),
@@
 partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
-invocation of the algorithm takes the initial state $\langle I, n\rangle$ and
-an empty abstract game tree $\emptyset$.  Initially the solver is playing on
+invocation of the algorithm takes the initial state $I$, bound $\kappa$ and an
+empty abstract game tree $\emptyset$.  Initially the solver is playing on
 behalf of the environment since that player takes the first move in every step.
@@
 \textsc{findCandidate}. For this purpose we use the heuristic that a partial
-strategy may be winning if each \textsc{outcome} of the strategy can be extended to a
-run of the game that is winning for the current player.  Clearly, if such a partial 
-strategy does not exist then no winning partial strategy can exist for the abstract 
-game tree.  
-We can formulate this heuristic as a SAT query, which is constructed
-recursively by $\textsc{treeFormula}$ (for the controller) or $\textsc{\textoverline{treeFormula}}$
-(for the environment) in Algorithm~\ref{alg:treeFormula}. The tree is first extended to the maximum
-bound with edges that are labeled with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown in Figure~\ref{}.  
-Then each edge in the tree is encoded in CNF by making copies of the transition relation along with the 
-winning condition.
-
-%The heuristic is based on the observation that for many synthesis problems in
-%the real world the final implementation will cooperate with its environment.
-%The existential search of the SAT solver simulates cooperation between the two
-%players.
+strategy may be winning if each \textsc{outcome} of the strategy can be
+extended to a run of the game that is winning for the current player.  Clearly,
+if such a partial strategy does not exist then no winning partial strategy can
+exist for the abstract game tree. We can formulate this heuristic as a SAT
+query, which is constructed recursively by $\textsc{treeFormula}$ (for the
+controller) or $\textsc{\textoverline{treeFormula}}$ (for the environment) in
+Algorithm~\ref{alg:treeFormula}.
+
+The tree is first extended to the maximum bound with edges that are labeled
+with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
+in Figure~\ref{}. Then copies of state and action variables are introduced for
+each node in tree and opponent action variables for each edge.  Opponent
+actions are set to $\textsc{action}(e)$, the action that labels the edge of the
+tree $e$. The new state and action variables are connected to the successor
+nodes (\textsc{succ}) by an encoding transition relation and constrained to the
+winning condition of the player.
 
@@
         \Function{treeFormula}{$T$}
-        \If{$\Call{rank}{T} = 0$}
+        \If{$\Call{bound}{T} = 0$}
         \State \Return{ $\lnot \Call{E}{X_{T}}$ }
@@
         \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
-            $$\bigwedge_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{treeFormula}{n})$$
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
         }
@@
         \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{rank}{T} = 0$}
+        \If{$\Call{bound}{T} = 0$}
         \State \Return{\Call{E}{$X_{T}$}}
@@
         \State \Return{ $\Call{E}{X_{T}} \lor$ \\
-        $$\bigvee_{n \in \Call{succ}{T}}(\Call{$\delta$}{X_n, U_n, C_n, X'_n} \land \Call{label}{n} \land \Call{\textoverline{treeFormula}}{n})$$ }
+        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
         \EndIf
@@
 controller, the strongest claim that can be made is that the strategy is
-winning as long as the game does not extend beyond the bound.
-
-It is possible to set a bound such that all runs in the unbounded game will be
+winning as long as the game does not extend beyond the maximum bound.
+
+It is possible to set a maximum bound such that all runs in the unbounded game will be
 considered. The na\"ive approach is to use size of the state space as the bound
@@
 smallest number $d$ such that for any state $x$ there is a path of length $\leq
-d$ to all other reachable states. This approach is common in bounded model
-checking but for synthesis it quickly becomes intractable to consider such long
-runs.
-
-When performing model checking or synthesis with BDDs \cite{burch1990} the set
-of states that are winning for the environment is iteratively constructed by
-computing the states from which the environment can force the game into the
-previous winning set. Eventually this process reaches a fixed point and the
-total set of environment winning states is known.
-
-We apply a similar fixed point computation on game states to the bounded
-synthesis algorithm in order to promote it to unbounded games. Our unbounded
-procedure iteratively executes the bounded algorithm while recording states
-from abstract game trees. By carefully managing the way states are recorded we
-guarantee that a fixed point proves the controller's strategy correct.
-
-\subsection{Extending Bounded Synthesis to Unbounded Games}
+d$ to all other reachable states. However, the diameter is difficult to
+estimate and can still lead to infeasibly long games.
+
+We instead present an approach that iteratively solves games of increasing
+bound while learning states. We show that reaching a fixed point in learned
+states is sufficient for completeness.
+
+%%%\subsection{Extending Bounded Synthesis to Unbounded Games}
 
@@
 
-The unbounded synthesis algorithm enhances the bounded method with
-computational learning. States are carefully recorded from unsatisfiable
-abstract games while maintaining an invariant ensuring the monotonicity of
-learning.
-
-During the CEGAR loop of the bounded algorithm, each of the competing solvers
-searches for satisfying assignments of labels to abstract game trees. When the
-abstract game is unsatisfiable this indicates that the states represented by
-nodes in the abstract game tree are losing for the current player. We extract
-these states from the game tree using interpolation.
-
-When executing the unbounded solver, lines 19 and 20 become active in the
-bounded solver. These lines call the learning procedures when the solver fails
-to find a candidate for an abstract game tree. The states symbolically
-represented by nodes in the tree are losing for whichever player could not find
-a winning candidate and can be extracted from the tree using interpolation.
-
-The states in an abstract game with no controller candidate are
-\emph{must-losing}. The environment can always force the game into the error
-states from these states.
-
-From abstractions with no environment candidate we record the complement of
-states in the tree as \emph{may-losing}. The environment cannot reach the error
-state in a number of steps equal to the distance to the bottom of the tree.
-
-We maintain a set of states for each rank up to the current bound. We maintain
-an invariant over these sets via careful construction so that they are
-monotonically increasing by rank. We also ensure that the environment is unable
-to force play from one set to the next. Due to these invariants, when two
-adjacent sets become equivalent we know that the algorithm has reached a fixed
-point and the controller is winning in the unbounded game (line 6).
-
-%%%\begin{figure}
-%%%    \centering
-%%%    \begin{tikzpicture}
-%%%        \node [minimum size=2cm,circle,draw] (a){$B^m[n]$};
-%%%        \node [minimum size=1.5cm,circle,draw,right=1.8cm of a] (b){$B^m[n-1]$};
-%%%        \node [minimum size=1cm,right=1.5cm of b] (dots){$\dots$};
-%%%        \node [minimum size=0.5cm,circle,draw,right=1.5cm of dots] (c){$B^m[0]$};
-%%%        \draw[-latex] (a.east) -- node [above] {$\supset$} (b.west);
-%%%        \draw[-latex] (b.east) -- node [above] {$\supset$} (dots);
-%%%        \draw[-latex] (dots.east) -- node [above] {$\supset$} (c.west);
-%%%    \end{tikzpicture}
-%%%    \caption{May-losing states}
-%%%    \label{fig:maylosing}
-%%%\end{figure}
+%%%The unbounded synthesis algorithm enhances the bounded method with
+%%%computational learning. States are carefully recorded from unsatisfiable
+%%%abstract games while maintaining an invariant ensuring the monotonicity of
+%%%learning.
+
+%%%During the CEGAR loop of the bounded algorithm, each of the competing solvers
+%%%searches for satisfying assignments of labels to abstract game trees. When the
+%%%abstract game is unsatisfiable this indicates that the states represented by
+%%%nodes in the abstract game tree are losing for the current player. We extract
+%%%these states from the game tree using interpolation.
+
+%%%When executing the unbounded solver, lines 19 and 20 become active in the
+%%%bounded solver. These lines call the learning procedures when the solver fails
+%%%to find a candidate for an abstract game tree. The states symbolically
+%%%represented by nodes in the tree are losing for whichever player could not find
+%%%a winning candidate and can be extracted from the tree using interpolation.
+
+%%%The states in an abstract game with no controller candidate are
+%%%\emph{must-losing}. The environment can always force the game into the error
+%%%states from these states.
+
+%%%From abstractions with no environment candidate we record the complement of
+%%%states in the tree as \emph{may-losing}. The environment cannot reach the error
+%%%state in a number of steps equal to the distance to the bottom of the tree.
+
+%%%We maintain a set of states for each rank up to the current bound. We maintain
+%%%an invariant over these sets via careful construction so that they are
+%%%monotonically increasing by rank. We also ensure that the environment is unable
+%%%to force play from one set to the next. Due to these invariants, when two
+%%%adjacent sets become equivalent we know that the algorithm has reached a fixed
+%%%point and the controller is winning in the unbounded game (line 6).
 
 \subsection{Learning States with Interpolants}
+
+We modify the bounded synthesis algorithm to learn a subset of states from
+failed attempts to find candidate strategies.
+
+When the abstract game is unsatisfiable this indicates that the states
+represented by nodes in the abstract game tree are losing for the current
+player. We extract these states from the game tree using interpolation.
 
@@
     \centering
-%%%    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-%%%        \node [circle,draw] (root){}
-%%%            child {node [circle,draw] {}
-%%%                child {node [circle,draw] {}
-%%%                    edge from parent [fixed] node [left] {$b$}
-%%%                }
-%%%                edge from parent [fixed] node [left] {$a$}
-%%%            }
-%%%            child {node [circle,draw] {}
-%%%                edge from parent [fixed] node [right] {$d$}
-%%%            }
-%%%            node [above=4pt] {$\langle s, k \rangle$};
-%%%    \end{tikzpicture}
     \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-        \node [] (root){}
-        child {node [circle,draw] {}
-                child {node [circle,draw] {}
+        \node [circle,draw] (root){}
+            child {node [circle,draw] {}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$b$}
+                }
                 edge from parent [fixed] node [left] {$a$}
             }
-                child {node [circle,draw] {} edge from parent[unfixed]
-                    edge from parent [fixed] node [right] {$b$}
-                }
-            };
+            child {node [circle,draw] {}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [right] {$e$}
+                }
+                edge from parent [fixed] node [right] {$d$}
+            }
+            node [above=4pt] {$\langle s, 2 \rangle$};
     \end{tikzpicture}
-%%%    \begin{tikzpicture}[overlay]
-%%%        \draw [rotate = -20] (-1.9,0.35) ellipse (0.7 and 1.3);
-%%%    \end{tikzpicture}
     \caption{A controller-losing game tree}
@@
 
-Consider the snippet of a game tree in Figure \ref{fig:interpolatetree}. The
+Consider the game tree in Figure \ref{fig:interpolatetree}. The
 tree is losing for the controller, the topmost node is at rank 1, and the $a$ and

### 2016-01-29 paper.tex
reference to fig:extended

@@
 with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
-in Figure~\ref{}. Then copies of state and action variables are introduced for
+in Figure~\ref{fig:extended}. Then copies of state and action variables are introduced for
 each node in tree and opponent action variables for each edge.  Opponent

### 2016-01-29 paper.tex
Lots of editing

@@
 \usetikzlibrary{positioning}
+\usetikzlibrary{plotmarks}
 
@@
 unbounded realizability.
-Our method can be seen as a hybrid between the counterexample-guided algorithm by
-Narodytska et al. and methods based on winning set computation: we use the former to guide 
-the search, while we rely on the latter to ensure convergence.
+Our method can be seen as a hybrid between the counterexample-guided algorithm
+by Narodytska et al. and methods based on losing set computation: we use the
+former to guide the search, while we rely on the latter to ensure convergence.
 
@@
 describe and prove the correctness of our extension of the algorithm to
-unbounded games. In the following sections we discuss optimisations to the
-algorithm, evaluate our methodology, and compare our approach to other
-synthesis techniques.
+unbounded games. In the following sections we evaluate our methodology, and
+compare our approach to other synthesis techniques.
 
@@
 winning for the controller if $x_0 = I \land \forall i \in \{1..n\} (\lnot
-E(x_i))$.  In a \emph{bounded game} with bound $\kappa$ all runs are restricted
-to length $\kappa$, whereas unbounded games consider runs of infinite length.
-Since we consider only deterministic games, a run is uniquely described by a
-list of assignments to $U$ and $C$.
+E(x_i))$.  In a \emph{bounded game} with maximum bound $\kappa$ all runs are
+restricted to length $\kappa$, whereas unbounded games consider runs of
+infinite length.  Since we consider only deterministic games, a run is uniquely
+described by a list of assignments to $U$ and $C$.
 
@@
 uncontrollable inputs to controllable actions. A controller strategy is winning
-in a bounded game of rank $n$ if all runs $(x_0, u_0, \pi^c(x_0, u_0)), (x_1,
-u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are winning.
-Bounded \emph{realizability} is the problem of determining the existence of
-such a strategy for a bounded game.
+in a bounded game of maximum bound $\kappa$ if all runs $(x_0, u_0, \pi^c(x_0,
+u_0)), (x_1, u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are
+winning.  Bounded \emph{realizability} is the problem of determining the
+existence of such a strategy for a bounded game.
 
@@
 game: universal for the controller and existential for the environment.  Just
-as the players alternate in choosing actions every step of the game, the
+as the players alternate in choosing actions every round of the game, the
 formula is prefixed by a quantifier alternation over $U$ and $C$ for each
-step.
+round.
 
@@
 strategy for the game.
+
+An abstraction of the game restricts actions available to one of the players.
+Specifically, we consider abstractions represented as trees of actions,
+referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an
+example abstract game tree for the controller (abstract game trees for the
+environment are similar).  In the abstract game, the controller is required to
+pick actions from the tree, starting from the root node.  After reaching a
+leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
+restricts the initial controller action to the set $\{a, d\}$.  After choosing
+action $a$, the controller is required to play action $b$ in the next round.
+It then reaches a leaf of the tree and continues playing unrestricted.  The
+root of the tree is annotated by the initial state $s$ of the abstract game and
+the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
+of all nodes of a tree $T$ and $\textsc{leaves}(T)$ the subset of leaf nodes.
 
@@
 
-An abstraction of the game restricts actions available to one of the players.
-Specifically, we consider abstractions represented as trees of actions,
-referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an example abstract game
-tree for the controller (abstract game trees for the environment are similar).  In the abstract game, 
-the controller is required to pick actions
-from the tree, starting from the root node.  After reaching a leaf, it
-continues playing unrestricted.  The tree in Figure~\ref{fig:agt} restricts the
-initial controller action to the set $\{a, d\}$.  After choosing action $a$,
-the controller is required to play action $b$ in the next round.  It then reaches a leaf 
-of the tree and continues playing unrestricted.
-The root of the tree is annotated by the initial state $s$ of the abstract game and the 
-bound $k$ on the number of rounds. 
-We denote $\textsc{nodes}(T)$ the set of all nodes of a tree $T$ and
-$\textsc{leaves}(T)$ the subset of leaf nodes.
-
-Given a controller (environment)
-abstract game tree $T$ a \emph{partial strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ 
-($Strat: \textsc{nodes}(T) \rightarrow C$) labels each node of the tree with
-the opponent's action to be played in that node.   Figure~\ref{fig:strategy}
-shows an example partial strategy.  The environment starts by choosing action
-$\alpha$.  If the controller plays $a$, the environment responds with $\beta$
-in the next round, and so on. Given a partial strategy $Strat$, we can map each
-leaf $l$ of the abstract game tree to $\langle
-s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
-playing all controllable and uncontrollable actions on the path from the root
-to the leaf.
+Given a controller (environment) abstract game tree $T$ a \emph{partial
+strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ ($Strat: \textsc{nodes}(T)
+\rightarrow C$) labels each node of the tree with the opponent's action to be
+played in that node.   Figure~\ref{fig:strategy} shows an example partial
+strategy.  The environment starts by choosing action $\alpha$.  If the
+controller plays $a$, the environment responds with $\beta$ in the next round,
+and so on. Given a partial strategy $Strat$, we can map each leaf $l$ of the
+abstract game tree to $\langle s',i'\rangle=\textsc{outcome}(\langle s,
+i\rangle, Strat, l)$ obtained by playing all controllable and uncontrollable
+actions on the path from the root to the leaf.
 
@@
             \IIf{$cex = \False$} \Return $cand$ \EndIIf \Comment{No counterexample: return candidate}
-            \State $T' \gets $ \Call{append}{$T', l, u$} \Comment{Refine AGT with counterexample}
-            \State $cand \gets $ \Call{solveAbstract}{$p, s, k, T'$} \Comment{Solve refined AGT}
+            \State $T' \gets $ \Call{append}{$T', l, u$} \Comment{Refine $T'$ with counterexample}
+            \State $cand \gets $ \Call{solveAbstract}{$p, s, k, T'$} \Comment{Solve refined game tree}
         \EndLoop
@@
             \For{$l \in leaves(gt)$}
-            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$s, k, cand, l$} \Comment{Get rank and state at leaf}
+            \State $\langle k', s'\rangle \gets $ \Call{outcome}{$s, k, cand, l$} \Comment{Get bound and state at leaf}
                 \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $k'$, $s'$, $\emptyset$} \Comment{Solve for the opponent}
@@
 empty abstract game tree $\emptyset$.  Initially the solver is playing on
-behalf of the environment since that player takes the first move in every step.
-The empty game tree does not constrain opponent moves, hence solving such an
-abstraction is equivalent to solving the original concrete game.
+behalf of the environment since that player takes the first move in every game
+round.  The empty game tree does not constrain opponent moves, hence solving
+such an abstraction is equivalent to solving the original concrete game.
 
@@
 with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
-in Figure~\ref{fig:extended}. Then copies of state and action variables are introduced for
-each node in tree and opponent action variables for each edge.  Opponent
-actions are set to $\textsc{action}(e)$, the action that labels the edge of the
-tree $e$. The new state and action variables are connected to the successor
-nodes (\textsc{succ}) by an encoding transition relation and constrained to the
-winning condition of the player.
+in Figure~\ref{fig:extended}. Then copies of state and action variables are
+introduced for each node in tree and opponent action variables for each edge.
+Opponent actions are set to $\textsc{action}(e)$, the action that labels the
+edge of the tree $e$. The state and action variables of node $n$ are connected
+to successor nodes $\textsc{succ}(n)$ by an encoding of the transition relation
+and constrained to the winning condition of the player.
 
@@
 We instead present an approach that iteratively solves games of increasing
-bound while learning states. We show that reaching a fixed point in learned
-states is sufficient for completeness.
+bound while learning states with interpolation over abstract games. We show
+that reaching a fixed point in learned states is sufficient for completeness.
 
@@
             \State $B^M \gets E$
-            \State $B^m[0] \gets E$
+            \State $B^m[0] \gets \top$
             \For{$k = 1 \dots$}
@@
                 \State $B^m[k] \gets \top$
-                \State \Call{checkRank}{$I, k$}
+                \State \Call{checkBound}{$I, k$}
             \EndFor
@@
 %%%            $\forall u_{k..1} \exists c_{k..1} \  s(x_k) \land \lnot E_k \land \delta(x_k, u_k, c_k) \land \lnot E(x_{k-1}) \land ... \land \delta(x_1, u_1, c_1) \land \lnot E(x_0)$
-        \Function{checkRank}{$s, k$}
+        \Function{checkBound}{$s, k$}
             \State \Return \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
@@
 \end{algorithm}
-
-%%%The unbounded synthesis algorithm enhances the bounded method with
-%%%computational learning. States are carefully recorded from unsatisfiable
-%%%abstract games while maintaining an invariant ensuring the monotonicity of
-%%%learning.
 
@@
 We modify the bounded synthesis algorithm to learn a subset of states from
-failed attempts to find candidate strategies.
-
-When the abstract game is unsatisfiable this indicates that the states
-represented by nodes in the abstract game tree are losing for the current
-player. We extract these states from the game tree using interpolation.
+failed attempts to find candidate strategies. When \textsc{findCandidate}
+cannot find a strategy for an abstract game tree beginning at a state $s$, we
+know that $s$ is a losing state. We can also attempt to learn additional losing
+states from the tree via interpolation.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {$e$}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {$f$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, 2 \rangle$};
+        \end{tikzpicture}
+        \caption{A losing AGT $T$}
+        \label{fig:interpolatetree}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {$e$}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {$f$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, 2 \rangle$};
+        \end{tikzpicture}
+        \caption{Tree slice $T_1$}
+        \label{fig:treef1}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] {}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$b$}
+                };
+%%%                node [above=4pt] {$\langle s', 1 \rangle$};
+        \end{tikzpicture}
+        \caption{Tree slice $T_2$}
+        \label{fig:treef2}
+    \end{subfigure}
+\end{figure}
+
+Consider the controller-losing game tree, $T$, in
+Figure~\ref{fig:interpolatetree}. We choose any leaf node $l$ of $T$ and its
+parent $n$. We then split the tree at $n$ such that both slices $T_1$ and $T_2$
+contain $n$. Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$
+except $l$, and $T_2$ (Figure~\ref{fig:treef2}) contains only $n$ and $l$ - a
+single round of the game.  There is no candidate strategy for $T$ so $s \land
+\textsc{treeFormula}(T)$ is unsatisfiable, and by extension $s \land
+\textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$ must also be
+unsatisfiable. This enables the construction of an \emph{interpolant} that
+captures losing states at node $n$.
 
 Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
-it is possible to construct a Craig interpolant\cite{craig1957} $\mathcal{I}$
+it is possible to construct a Craig interpolant~\cite{craig1957} $\mathcal{I}$
 such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
@@
 
-\begin{figure}
-    \centering
-    \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-        \node [circle,draw] (root){}
-            child {node [circle,draw] {}
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [left] {$b$}
-                }
-                edge from parent [fixed] node [left] {$a$}
-            }
-            child {node [circle,draw] {}
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [right] {$e$}
-                }
-                edge from parent [fixed] node [right] {$d$}
-            }
-            node [above=4pt] {$\langle s, 2 \rangle$};
-    \end{tikzpicture}
-    \caption{A controller-losing game tree}
-    \label{fig:interpolatetree}
-\end{figure}
-
-Consider the game tree in Figure \ref{fig:interpolatetree}. The
-tree is losing for the controller, the topmost node is at rank 1, and the $a$ and
-$b$ nodes are at rank 0. Since the tree is controller-losing we know that at
-least one state in the tree will uncontrollably lead to the error state.  As a
-result, $\textsc{treeFormula}(T)$ is unsatisfiable. If we take a step of the
-tree, such as the step where the environment plays $a$, and cut it from the
-rest of the tree then $\textsc{treeFormula}(step) \land
-\textsc{treeFormula}(parent)$ must too be unsatisfiable.
-
-We can construct an interpolant with $F_1 = \textsc{treeFormula}(parent)$ and
-$F_2 = \textsc{treeFormula}(step)$. The only variables shared between $F_1$ and
-$F_2$ are the state variables at rank 1. We know that $F_2 \land \mathcal{I}$
-is unsatisfiable, therefore all states in $\mathcal{I}$ must lose to the
-uncontrollable action $a$. We also know that $F_1 \to \mathcal{I}$, thus
-$\mathcal{I}$ contains all states reachable by the parent tree (on runs that
-avoid the error state.)
+We can construct an interpolant with $F_1 = s \land \textsc{treeFormula}(T_1)$
+and $F_2 = \textsc{treeFormula}(T_2)$. The only variables shared between $F_1$
+and $F_2$ are the state variable copies belonging to node $n$. By construction
+$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
+must lose to the action in the edge between $n$ and $l$, the edge labelled by
+action $b$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
+\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
+following $T_1$ and avoiding the error state.
 
 \begin{algorithm}
-    \caption{Amended tree formulas for Controller and Environment respectively}
+    \caption{Amended tree formulas for Controller and Environment}
     \label{alg:unboundedTreeFormula}
     \begin{algorithmic}[1]
-        \Function{treeFormula}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}}$ }
+        \Function{treeFormula}{$T$}
+        \If{$\Call{bound}{T} = 0$}
+        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
         \Else
-        \State \Return{ $\lnot \Call{$B^M$}{x^{gt}} \land \bigwedge_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{treeFormula}{n})$ }
+        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
+        }
         \EndIf
         \EndFunction
-        \algstore{tf2}
-    \end{algorithmic}
-
-    \begin{algorithmic}
-        \algrestore{tf2}
-        \Function{\textoverline{treeFormula}}{gt}
-        \If{$\Call{rank}{gt} = 0$}
-        \State \Return{\Call{E}{$x^{gt}$}}
+        \algstore{tf1}
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \algrestore{tf1}
+        \Function{\textoverline{treeFormula}}{$T$}
+        \If{$\Call{bound}{T} = 0$}
+        \State \Return{\Call{E}{$X_T$}}
         \Else
-        \State \Return{ $\Call{$B^m[rank(gt)]$}{x^{gt}} \land$ \\
-            $(\Call{E}{x^{gt}} \lor \bigvee_{n \in \Call{succ}{gt}}(\Call{$\delta$}{n} \land \Call{label}{n} \land \Call{$\overline{treeFormula}$}{n}))$ }
+        \State \Return{ \Call{$B^m[\Call{bound}{T}]$}{$X_{T}$} $\land$ \\
+            $$\Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
         \EndIf
@@
 
-Now we can consider the step where the environment plays $b$, and the parent
-now without either step in the snippet. The formula
-$(\textsc{treeFormula}(parent) \land \mathcal{I}) \land
-\textsc{treeFormula}(step)$ must be unsatisfiable.  $\mathcal{I}$ contains all
-states that lose to $a$ so any other state reachable at rank 1 in this subtree
-must lose to $b$.  Therefore we can compute another interpolant that contains
-states that lose to $b$.
-
-This is the foundation for a recursive algorithm that consumes an entire tree
-by removing a single step on each iteration (Algorithm \ref{alg:learn}). All
-learned states from which the controller must lose are recorded in a set of bad
-states $B^M$. Environment-losing are learned in the same way but are recorded
-by removing the state from a set of sets of controller may-losing states $B^m$.
-An environment-losing state of rank $n$ is removed from every set at rank $n$
-and lower.
+When we discover losing states for the controller we record them in a set of
+states that \emph{must} be bad, $B^M$. This set is initialised to the error
+state at the beginning of the algorithm and grows as losing states are
+discovered.  The \textsc{treeFormula} function is modified for the unbounded
+solver (see Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with
+$B^M$. This enables further interpolants to be constructed by recursively
+splitting more nodes from $T_1$ since the states that lost to $T_2$ are now
+blocked by $B^M$.
+
+It is more complicated to learn states from environment-losing abstract game
+trees. States are only losing for the environment for a particular bound. For a
+state learned at a node $n$ it is losing at $\textsc{bound}(n)$. Therefore we
+maintain an array of state sets indexed by the bound at which that set of
+states is losing for the environment. Instead of accumulating states, as we do
+in $B^M$, we instead remove states from the array $B^m$. This array is
+initialised with $\top$ and contains only states that \emph{may} lose for the
+controller. The modified \textsc{\textoverline{treeFormula}} function
+constrains the environment to may-losing states, which also allows for
+recursive learning over the entire tree.
 
@@
 smaller representation is necessary for maintaining the size of formulas passed
-to the SAT solver. However, it can occaisionally lead to state explosion.
+to the SAT solver. However, it can occasionally lead to state explosion.
+
+The recursive learning algorithms, \textsc{learn} (for the controller) and
+\textsc{\textoverline{learn}} (for the environment), are detailed in
+Algorithm~\ref{alg:learn}. When executing \textsc{findCandidate} in the
+unbounded solver, lines 19 and 20 become active and call into the learning
+procedures. The unbounded algorithm performs bounded synthesis for iteratively
+increasing maximum bounds until the initial state is in $B^M$ (environment
+wins) or $B^m$ reaches a fixed point (controller wins). We prove correctness in
+the next section.
 
@@
             \EndIIf
-            \State $n \gets $ non-leaf node with max rank
+            \State $n \gets $ non-leaf node with min bound
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
@@
         \Ensure \emph{May-invariant} holds
-        \Ensure $\sigma(X_T) \land B^m[$\Call{rank}{$T$}$] \equiv \bot$
+        \Ensure $\sigma(X_T) \land B^m[$\Call{bound}{$T$}$] \equiv \bot$
         \Function{\textoverline{learn}}{$\sigma, T$}
@@
             \EndIIf
-            \State $n \gets $ non-leaf node with max rank
+            \State $n \gets $ non-leaf node with min bound
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
             \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
-            \For{$i = 1$ to \Call{rank}{$n$}}
+            \For{$i = 1$ to \Call{bound}{$n$}}
                 \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
@@
 
-The correctness of the unbounded synthesis algorithm can be established
-independently from that of the bounded algorithm. Fortunately, the correctness
-of the bounded solver has been established in \cite{narodytska2014}.
-
 We define two global invariants of the algorithm.  The \emph{may-invariant}
 states that sets $B^m[i]$ grows monotonically with $i$ and that each $B^m[i+1]$
-overapproximates the uncontrollable predecessor of $B^m[i]$: $$\forall
-i<k.~B^m[i] \subseteq B^m[i+1], Upre(B^m[i]) \subseteq B^m[i+1].$$
+overapproximates the states from which the environment can force the game into
+$B^m[i]$. We call this operation $Upre$, the uncontrollable predecessor. So the
+\emph{may-invariant} is: $$\forall i<k.~B^m[i] \subseteq B^m[i+1], Upre(B^m[i])
+\subseteq B^m[i+1].$$
 
@@
 
-The sets $B^m$ and $B^M$ are only modified by the inductive solver, implemented
-by \textsc{learn} and \textoverline{\textsc{learn}} functions.  Below we prove that these
-functions indeed maintain the invariants.
-
-\subsection{Proof of \textsc{\textoverline{learn}}}
+The sets $B^m$ and $B^M$ are only modified by the functions \textsc{learn} and
+\textoverline{\textsc{learn}}.  Below we prove that these functions indeed
+maintain the invariants.
+
+\subsection{Proof of learning functions}
 
@@
 of depth 1, this means that the environment cannot force the game into
-$B^m[\textsc{rank}(n)]$ playing against the counterexample moves in $T_2$.
-Hence, $\II \cap Upre(B^m[\textsc{rank}(n)]) = \emptyset$.  Furthermore, since
+$B^m[\textsc{bound}(n)]$ playing against the counterexample moves in $T_2$.
+Hence, $\II \cap Upre(B^m[\textsc{bound}(n)]) = \emptyset$.  Furthermore, since
 the may-invariant holds, $\II \cap Upre(i) = \emptyset$, for all $i <
-\textsc{rank}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
-\textsc{rank}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
+\textsc{bound}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
+\textsc{bound}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
 first post-condition.
@@
 \textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
-removing $\II$ from $B^m[\textsc{rank}(n)]$ makes $\sigma(X_T) \land
+removing $\II$ from $B^m[\textsc{bound}(n)]$ makes $\sigma(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
@@
 is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
-\equiv B^m[\textsc{rank}(T)](X^T)$; hence $\sigma(X_T) \land
+\equiv B^m[\textsc{bound}(T)](X^T)$; hence $\sigma(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1) \equiv \sigma(X_T) \land
-B^m[\textsc{rank}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
+B^m[\textsc{bound}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
 \textsc{\textoverline{learn}} holds.
 
+The proof of \textsc{learn} is similar to the above proof of \textsc{learn}. An
+interpolant constructed from $F_1=\sigma(X_T) \land \textsc{treeFormula}(T_1)$
+and $F_2 = \textsc{treeFormula}(T_2)$ has the property $\II \land F_2 \equiv
+\bot$ and the precondition ensures that the controller is unable to force the
+game into $B^M$ playing against the counterexample moves in $T_2$. Thus adding
+$\II$ to $B^M$ maintains the must-invariant satisfying the first postcondition.
+
+Likewise, in the second last recursive call of \textsc{learn} with the empty
+tree $T_1$ and root node $n$: $\textsc{treeFormula}(T_1) \equiv \lnot
+B^M(X_T)$.  Hence $\sigma(X_T) \land \textsc{treeFormula}(T_1) \equiv
+\sigma(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $\sigma(X_T) \land
+B^M(X_T)$, the second postcondition, is true.
+
 \subsection{Proof of Termination}
 
-We must prove that \textsc{checkRank} terminates and that upon termination its
-postcondition holds.
-
-We must prove that $\textsc{checkRank}$ terminates and that upon termination
-its postcondition holds, i.e., state $s$ is removed from $B^m[k]$ if there is a
-winning controller strategy on the bounded safety game of rank $k$ or it is
-added to $B^M$ otherwise. Termination follows from completeness of
-counterexample-guided search, which terminates after enumerating all possible
-opponent moves in the worst case.
-
-Assume that there is a winning strategy for the controller at rank $k$. This
-means that at some point the algorithm discovers a counterexample tree of rank
-$k$ for which the environment cannot force into $E$. The algorithm then invokes
-the \textsc{\textoverline{learn}} method, adds $s$ to $B^M$.  Alternatively, if
-there is a winning strategy for the environment at rank $k$ then a
-counterexample losing for the controller will be found. Subsequently
-\textsc{learn} will be called and $s$ eliminated from $B^m[k]$.
-
-\section{Optimisations}
-
-%%%In this section two optimisations to the unbounded synthesis algorithm are
-%%%described.
-
-\subsection{Generalising the initial state}
-
-If more states are removed from $B^m$ on each iteration of \textsc{checkRank}
-the algorithm will converge and terminate sooner. Additionally, any states
-removed from $B^M$ will reduce the number of states to be considered by the
-controller in future iterations.
-
-As it is written, the algorithm only considers an overapproximations of states
-reachable from $I$ for learning. Assuming that the algorithm does not
-terminate, then the addition of an extra step into the game on each iteration
-has the potential to greatly increase the reachable states. Considering some of
-these states earlier than they become reachable may lead to earlier
-termination.
+We must prove that $\textsc{checkBound}$ terminates and that upon termination
+its postcondition holds, i.e., state $s$ is removed from $B^m[\kappa]$ if there
+is a winning controller strategy on the bounded safety game of maximum bound
+$\kappa$ or it is added to $B^M$ otherwise. Termination follows from
+completeness of counterexample-guided search, which terminates after
+enumerating all possible opponent moves in the worst case.
+
+Assume that there is a winning strategy for the controller at bound $\kappa$.
+This means that at some point the algorithm discovers a counterexample tree of
+bound $\kappa$ for which the environment cannot force into $E$. The algorithm
+then invokes the \textsc{\textoverline{learn}} method, which removes $s$ from
+$B^m[\kappa]$.  Alternatively, if there is a winning strategy for the
+environment at bound $\kappa$ then a counterexample losing for the controller
+will be found.  Subsequently \textsc{learn} will be called and $s$ added to
+$B^M$.
+
+\subsection{Optimisation: Generalising the initial state}
+
+This optimisation allows us to learn may and must losing states faster.
+Starting with a larger set of initial states we increase the reachable set and
+hence increase the number of states learned by interpolation. This optimisation
+requires a modification to $\textsc{solveAbstract}$ to handle sets of states,
+which is not shown.
+
+The optimisation is relatively simple and is inspired by a common greedy
+heuristic for minimising $\texttt{unsat}$ cores. $I$ is a value assignment to
+each variable in $X$. If the environment does not win for $\langle I, k
+\rangle$ then we attempt to solve for a generalised version of $I$ by removing
+one assignment at a time. If the controller can win from the larger set of
+states then we continue generalising without it. In this way we learn more
+states by increasing the reachable set. In our benchmarks we have observed that
+this optimisation is beneficial on the first few iterations of
+\textsc{checkBound}.
 
@@
     \begin{algorithmic}
-        \Function{checkRank}{$s, k$}
-            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
+        \Function{checkBound}{$s, k$}
+            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, s, k, \emptyset$}
             \IIf{$r \neq \emptyset$} \Return $r$ \EndIIf
             \State $s' \gets s$
-            \For{$x \in s$}
-                \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, \langle s' \setminus \{x\}, k \rangle, \emptyset$}
-                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf
+            \For{$x \in X$}
+            \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, s' \setminus \{x\}, k, \emptyset$} 
+                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf \Comment{Removes the assignment to $x$ from $s'$}
             \EndFor
@@
 \end{algorithm}
-
-The optimisation that allows this is relatively simple and is inspired by a
-common greedy heuristic for minimising $\texttt{unsat}$ cores. $I$ is a value
-assignment to each variable in $X$. If the environment does not win for
-$\langle I, k \rangle$ then we attempt to solve for a generalised version of
-$I$ by removing one assignment at a time. If the controller can win from the
-larger set of states then we continue generalising without it. In this way we
-learn more states by increasing the reachable set. In our benchmarks we have
-observed that this optimisation is beneficial on the first few iterations of
-\textsc{checkRank}.
-
-%%%\subsection{Choosing opponent moves}
-
-%%%One aspect of the algorithm proposed by Narodytska et al.~\cite{narodytska2014}
-%%%is that it mimics real-world uses of synthesis by allowing cooperation between
-%%%system and environment. Unfortunately, allowing the opponents to pick moves for
-%%%one another can frequently backfire. It is common when modelling real-world
-%%%systems to abstract over out of scope failures with transitions that
-%%%immediately determine the outcome of the game. For example, in a network driver
-%%%a request to send a packet may fail due to a disconnected wire, which might be
-%%%modelled as an environment controlled transition to a system-winning state.
-
-%%%This slows down the synthesis process as all erroneous transition are
-%%%accumulated into the abstract game tree. To mitigate this effect we have a
-%%%heuristic to guess non-erroneous labels as \emph{default moves} that are used
-%%%instead of allowing the opponent to choose. This does not effect the
-%%%correctness of the algorithm. If line~(28) previously returned \texttt{sat} and
-%%%is \texttt{unsat} with default moves then those moves would eventually have
-%%%been added to the abstraction in a refinement step. If the formula is
-%%%\texttt{unsat} even with the opponent choosing moves then it must be
-%%%\texttt{unsat} with default moves.
 
@@
 
-\begin{figure}[h]
-    \centering
-    \begin{tikzpicture}[y=0.04cm, x=0.15cm]
-        %axis
-        \draw (0,0) -- coordinate (x axis mid) (60,0);
-            \draw (0,0) -- coordinate (y axis mid) (0,150);
-            %ticks
-            \foreach \x in {0,10,...,60}
-                \draw (\x,1pt) -- (\x,-3pt)
-                node[anchor=north] {\x};
-            \foreach \y in {0,25,...,150}
-                \draw (1pt,\y) -- (-3pt,\y) 
-                    node[anchor=east] {\y}; 
-        %labels      
-        \node[below=0.8cm] at (x axis mid) {Time (minutes)};
-        \node[rotate=90, above left=1cm and 0.7cm] at (y axis mid) {Instances solved};
-
-        %Plots
-        \draw plot[mark=x, mark options={fill=white}] file {bench_termite.dat};
-        \draw plot[mark=+, mark options={fill=white}] file {bench_demiurge.dat};
-
-        \begin{scope}[shift={(30,20)}] 
-            \draw (0,0) -- 
-                plot[mark=x, mark options={fill=white}] (0.25,0) -- (0.5,0) 
-                node[right]{Unbounded synthesis};
-            \draw[yshift=\baselineskip] (0,0) -- 
-                plot[mark=+, mark options={fill=white}] (0.25,0) -- (0.5,0)
-                node[right]{Demiurge};
-        \end{scope}
-
-    \end{tikzpicture}
-\end{figure}
-
 The implementation of our algorithm uses \textsc{Glucose}~\cite{audemard2014}
 for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
-generation. We compared our implementation to
-\textsc{Demiurge}~\cite{bloem2014}, the only SAT-based solver to compete in the
-sequential realisability track of SYNTCOMP'15. The benchmarks were run on a
-cluster of Intel Quad Core Xeon E5405 CPUs with 16GB of memory.  Each solver
-was allowed exclusive access to a node for one hour to solve an instance.
-
-Our implementation was able to solver 103 out of the 250 specification in the
-alloted time, including 12 instances that were unable to be solved by any other
-solver in SYNTCOMP'15. \textsc{Demiurge} was able to solve 136 specifications.
-77 of those instances were also solved by our implementation, so
-\textsc{Demiurge} solved 59 instances that our solver did not. However, our
-method was able to solve 26 specifications not solved by \textsc{Demiurge}.
-
-For reference, the results of the synthesis competition are shown in
-Figure~\ref{fig:syntcomp}. The competition ran on better hardware with a longer
-time period, which explains why \textsc{Demiurge} performed better.
+generation. The benchmarks were run on a cluster of Intel Quad Core Xeon E5405
+2GHz CPUs with 16GB of memory.  The solver was allowed exclusive access to a node
+for one hour to solve an instance.
+
+The results of our benchmarking are shown, along with the synthesis competition
+results~\cite{syntcompedacc}, in Figure~\ref{fig:syntcomp}. The competition was run on Intel Quad
+Core 32Ghz CPUs with 32GB of memory, also on isolated nodes for one hour per
+instance.
 
@@
 
+Our implementation was able to solver 103 out of the 250 specification in the
+alloted time, including 12 instances that were unable to be solved by any other
+solver in SYNTCOMP'15. The unique instances we solved are listed in
+Figure~\ref{fig:unique}.
+
+\begin{figure}[h]
+    \centering
+    \setlength{\tabcolsep}{16pt}
+    \begin{tabular}{l l}
+        \texttt{6s216rb0\_c0to31} & \texttt{driver\_c10n} \\
+        \texttt{cnt30y} &  \texttt{stay18y} \\
+        \texttt{driver\_a10n} & \texttt{stay20n} \\
+        \texttt{driver\_a8n} & \texttt{stay20y} \\
+        \texttt{driver\_b10y} & \texttt{stay22n} \\
+        \texttt{driver\_b8y} & \texttt{stay22y} \\
+    \end{tabular}
+    \caption{Instances uniquely solved by our approach}
+    \label{fig:unique}
+\end{figure}
+
 Five of the instances unique to our solver are device driver instances and
-another five are from the \texttt{stay} family, which are classified as toy
-instances.  Out of the 26 instances that \textsc{Demiurge} could not solve,
-there were only 6 different benchmark classes. This supports the hypothesis
+another five are from the \texttt{stay} family. This supports the hypothesis
 that different game solving methodologies perform better on certain classes of
@@
 
-Whilst our solver is unable to solve as many instances as other tools yet, it
-was able to solve more unique instances than any solver in the competition.
-This shows that the method has potential. In the parallel track,
-\textsc{Demiurge} competes with a suite of 3 separate but communicating
-solvers. The solvers relay unsafe states to one another, which is compatible
-with the set $B^M$ in our solver. It remains future work to explore the
-possibility of combining these solvers.
+We also present a cactus plot of the number of instances solved over time. For
+reference we plotted our results along with \textsc{Demiurge}~\cite{bloem2014},
+the only SAT-based tool in the competition, and the winner of the sequential
+realisability track \textsc{Simple BDD Solver 2}~\cite{walker2014}. We were
+able to run \textsc{Demiurge} on our hardware but show competition results for
+\textsc{Simple BDD Solver 2}.
+
+\begin{figure}[h]
+    \centering
+    \begin{tikzpicture}[y=0.03cm, x=0.15cm]
+        %axis
+        \draw (0,0) -- coordinate (x axis mid) (60,0);
+            \draw (0,0) -- coordinate (y axis mid) (0,200);
+            %ticks
+            \foreach \x in {0,10,...,60}
+                \draw (\x,1pt) -- (\x,-3pt)
+                node[anchor=north] {\x};
+            \foreach \y in {0,25,...,200}
+                \draw (1pt,\y) -- (-3pt,\y) 
+                    node[anchor=east] {\y}; 
+        %labels      
+        \node[below=0.8cm] at (x axis mid) {Time (minutes)};
+        \node[rotate=90, above left=1cm and 0.7cm] at (y axis mid) {Instances solved};
+
+        %Plots
+        \draw plot[mark=x, mark options={fill=white}] file {bench_termite.dat};
+        \draw plot[mark=+, mark options={fill=white}] file {bench_demiurge.dat};
+        \draw plot[mark=triangle, mark options={fill=white}] file {bench_simple.dat};
+
+        \begin{scope}[shift={(30,20)}] 
+            \draw (0,0) -- 
+                plot[mark=x, mark options={fill=white}] (0.25,0) -- (0.5,0) 
+                node[right]{Unbounded synthesis};
+            \draw[yshift=\baselineskip] (0,0) -- 
+                plot[mark=+, mark options={fill=white}] (0.25,0) -- (0.5,0)
+                node[right]{Demiurge};
+            \draw[yshift=2\baselineskip] (0,0) -- 
+                plot[mark=triangle, mark options={fill=white}] (0.25,0) -- (0.5,0)
+                node[right]{Simple BDD Solver 2};
+        \end{scope}
+
+    \end{tikzpicture}
+\end{figure}
+
+Whilst our solver is unable to solve as many instances as other tools, it was
+able to solve more unique instances than any solver in the competition. This
+confirms that our methodology is able to fill gaps in a state of the art
+synthesis toolbox by more efficiently solving instances that are hard for other
+techniques. For this reason our solver would be a worthwhile addition to a
+portfolio solver. In the parallel track of the competition, \textsc{Demiurge}
+competes with a suite of 3 separate but communicating solvers. The solvers
+relay unsafe states to one another, which is compatible with the set $B^M$ in
+our solver. It remains future work to explore the possibility of combining
+these solvers.
 
@@
 
-The algorithm presented in this paper solves realisability for safety games.
-There is a method for extracting strategies from abstract game trees that is
-compatible with our method \cite{een2015}. It involves a similar interpolation
-approach for discovering states in game tree nodes.
+The original bounded synthesis algorithm of Narodytska et
+al.~\cite{narodytska2014} solves realisability without constructing a strategy.
+In \cite{een2015} the realisability algorithm is extended with strategy
+extraction. The technique relies on interpolation of abstract game trees.  In
+the present work we use interpolation in a similar way for computational
+learning.
 
@@
 
-We also aim to experiment with optimisations by focusing on heuristics for
-avoiding unhelpful actions as well as constructing succinct representations of
-interpolants without resorting to BDDs.
-
 \bibliographystyle{splncs03}

### 2016-01-29 paper.tex
Improvements to sections 3-5

@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
     \begin{algorithmic}[1]
@@
                 \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
-                    \State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
-                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$\sigma, \hat{T}$}}{\Call{\textoverline{learn}}{$\sigma, \hat{T}$}} \EndIfElse
+                    %\State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
+                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$s, \hat{T}$}}{\Call{\textoverline{learn}}{$s, \hat{T}$}} \EndIfElse
                 \EndIf
@@
 
-    \caption{Bounded Synthesis}
+    \caption{Bounded synthesis}
     \label{alg:bounded}
@@
 We instead present an approach that iteratively solves games of increasing
-bound while learning states with interpolation over abstract games. We show
+bound while learning winning states from abstract games using 
+interpolation. We show
 that reaching a fixed point in learned states is sufficient for completeness.
@@
 %%%\subsection{Extending Bounded Synthesis to Unbounded Games}
+
+%%%During the CEGAR loop of the bounded algorithm, each of the competing solvers
+%%%searches for satisfying assignments of labels to abstract game trees. When the
+%%%abstract game is unsatisfiable this indicates that the states represented by
+%%%nodes in the abstract game tree are losing for the current player. We extract
+%%%these states from the game tree using interpolation.
+
+%%%When executing the unbounded solver, lines 19 and 20 become active in the
+%%%bounded solver. These lines call the learning procedures when the solver fails
+%%%to find a candidate for an abstract game tree. The states symbolically
+%%%represented by nodes in the tree are losing for whichever player could not find
+%%%a winning candidate and can be extracted from the tree using interpolation.
+
+%%%The states in an abstract game with no controller candidate are
+%%%\emph{must-losing}. The environment can always force the game into the error
+%%%states from these states.
+
+%%%From abstractions with no environment candidate we record the complement of
+%%%states in the tree as \emph{may-losing}. The environment cannot reach the error
+%%%state in a number of steps equal to the distance to the bottom of the tree.
+
+%%%We maintain a set of states for each rank up to the current bound. We maintain
+%%%an invariant over these sets via careful construction so that they are
+%%%monotonically increasing by rank. We also ensure that the environment is unable
+%%%to force play from one set to the next. Due to these invariants, when two
+%%%adjacent sets become equivalent we know that the algorithm has reached a fixed
+%%%point and the controller is winning in the unbounded game (line 6).
+
+\subsection{Learning States with Interpolants}
+
+We extend the bounded synthesis algorithm to learn states losing for one of the players from
+failed attempts to find candidate strategies.  The learning procedure kicks in whenever
+\textsc{findCandidate} cannot find a candidate strategy for an abstract game tree with initial
+state $s$.  We know that $s$ is a losing state.  We can learn additional losing
+states from the tree via interpolation.  This is achieved in lines~18--20 in Algorithm~\ref{alg:bounded},
+enabled in the unbounded version of the algorithm, which invoke \textsc{learn} or
+\textsc{\textoverline{learn}} to learn controller or environment losing states respectively
+(Algorithm~\ref{alg:learn}).
+
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {$e$}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {$f$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, 2 \rangle$};
+        \end{tikzpicture}
+        \caption{A losing AGT $T$}
+        \label{fig:interpolatetree}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$a$}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {$e$}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {$f$}
+                    }
+                    edge from parent [fixed] node [right] {$d$}
+                }
+                node [above=4pt] {$\langle s, 2 \rangle$};
+        \end{tikzpicture}
+        \caption{Tree slice $T_1$}
+        \label{fig:treef1}
+    \end{subfigure}%
+    \begin{subfigure}[t]{.32\textwidth}
+        \centering
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+            \node [circle,draw] {}
+                child {node [circle,draw] {}
+                    edge from parent [fixed] node [left] {$b$}
+                };
+%%%                node [above=4pt] {$\langle s', 1 \rangle$};
+        \end{tikzpicture}
+        \caption{Tree slice $T_2$}
+        \label{fig:treef2}
+    \end{subfigure}
+    \caption{Splitting of an abstract game tree by the learning procedure.}
+\end{figure}
+
+\begin{algorithm}[t]
+    \begin{algorithmic}[1]
+        \Require $s(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
+        \Require \emph{Must-invariant} holds
+        \Ensure \emph{Must-invariant} holds
+        \Ensure $s(X_T) \land B^M$
+        \Function{learn}{$s, T$}
+            \IIf{$T = \emptyset$}
+                \Return
+            \EndIIf
+            \State $n \gets $ non-leaf node with min bound
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
+            \State $B^M \gets B^M \lor \mathcal{I}$
+            \State \Call{learn}{$s, T_1$}
+        \EndFunction
+        \algstore{learn}
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \algrestore{learn}
+        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
+        \Require \emph{May-invariant} holds
+        \Ensure \emph{May-invariant} holds
+        \Ensure $s(X_T) \land B^m[$\Call{bound}{$T$}$] \equiv \bot$
+        \Function{\textoverline{learn}}{$s, T$}
+            \IIf{$T = \emptyset$}
+                \Return
+            \EndIIf
+            \State $n \gets $ non-leaf node with min bound
+            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
+            \For{$i = 1$ to \Call{bound}{$n$}}
+                \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
+            \EndFor
+            \State \Call{\textoverline{learn}}{$s, T_1$}
+        \EndFunction
+    \end{algorithmic}
+    \caption{Learning algorithms}
+    \label{alg:learn}
+\end{algorithm}
+
+Consider the game tree $T$ in
+Figure~\ref{fig:interpolatetree}, whose edges are labelled with environment
+actions.  Assume that there does not exist a partial strategy for the controller
+winning against this tree and that the \textsc{findCandidate} function could not
+find a candidate strategy.
+We choose a non-leaf node $n$ of $T$ with maximal depth, i.e., a node whose children
+are leafs (Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that both slices $T_1$ and $T_2$
+contain a copy of $n$ (line~4). Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$
+except $n$'s children, and $T_2$ (Figure~\ref{fig:treef2}), which contains only $n$ and its children.  
+There is no candidate strategy for $T$ so $s \land
+\textsc{treeFormula}(T)$ is unsatisfiable.  By construction, 
+$\textsc{treeFormula}(T) = \textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$
+and hence $s \land
+\textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$ is also 
+unsatisfiable. This enables the construction of an \emph{interpolant} that
+captures losing states at node $n$.
+
+Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
+it is possible to construct a Craig interpolant~\cite{craig1957} $\mathcal{I}$
+such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
+$\mathcal{I}$ refers only to the intersection of variables in $F_1$ and $F_2$.
+An interpolant can be constructed efficiently from a resolution proof of the
+unsatisfiability of $F_1 \land F_2$~\cite{pudlak1997}.
+
+We can construct an interpolant with $F_1 = s \land \textsc{treeFormula}(T_1)$
+and $F_2 = \textsc{treeFormula}(T_2)$ (line~5). The only variables shared between $F_1$
+and $F_2$ are the state variable copies belonging to node $n$. By construction
+$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
+are losing against abstract game tree $T_2$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
+\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
+following $T_1$ and avoiding error states.
+
+We have discovered a set $\mathcal{I}$ of states losing for the controller.  
+We record these newly discovered states in a set $B^M$ of
+``bad'' states that \emph{must} be losing for the controller (Algorithm~\ref{alg:learn}, line~6). This set is initialised to the error
+set $E$ and grows as new losing states are
+discovered.  The \textsc{treeFormula} function is modified for the unbounded
+solver (Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with
+$B^M$. This enables further interpolants to be constructed by the learning procedure 
+recursively splitting more nodes from $T_1$ (Algorithm~\ref{alg:learn}, line~7) since the states that lost to $T_2$ are now
+blocked by $B^M$.
+
+\begin{algorithm}[t]
+    \caption{Amended tree formulas for Controller and Environment}
+    \label{alg:unboundedTreeFormula}
+    \begin{algorithmic}[1]
+        \Function{treeFormula}{$T$}
+        \If{$\Call{bound}{T} = 0$}
+        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
+        \Else
+        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
+        }
+        \EndIf
+        \EndFunction
+        \algstore{tf1}
+    \end{algorithmic}
+
+    \begin{algorithmic}[1]
+        \algrestore{tf1}
+        \Function{\textoverline{treeFormula}}{$T$}
+        \If{$\Call{bound}{T} = 0$}
+        \State \Return{\Call{E}{$X_T$}}
+        \Else
+        \State \Return{ \Call{$B^m[\Call{bound}{T}]$}{$X_{T}$} $\land$ \\
+            $$\Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
+        \EndIf
+        \EndFunction
+    \end{algorithmic}
+\end{algorithm}
+
+Learning of states losing for the environment is similar (\textsc{\textoverline{learn}} in Algorithm~\ref{alg:learn}). 
+The main difference is that environment-losing states are only losing for a particular
+bound: given that there does not exist an environment strategy that
+forces the game into an error state in $k$ rounds or less; there may still exist a longer
+environment-winning strategy.  We therefore record learned environment-losing states
+along with associated bounds.  To this end, we maintain a conceptually infinite array of
+sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.  $B^m[k]$ 
+are initialized to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
+is discovered for a node $n$ with bound $\textsc{bound}(n)$ in line~13, Algorithm~\ref{alg:learn}, this set is subtracted from 
+$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
+
+%It is more complicated to learn states from environment-losing abstract game
+%trees. 
+%States are only losing for the environment for a particular bound. For a
+%state learned at a node $n$ it is losing at $\textsc{bound}(n)$. Therefore we
+%maintain an array of state sets indexed by the bound at which that set of
+%states is losing for the environment. Instead of accumulating states, as we do
+%in $B^M$, we instead remove states from the array $B^m$. This array is
+%initialised with $\top$ and contains only states that \emph{may} lose for the
+%controller. 
+The modified \textsc{\textoverline{treeFormula}} function (\ref{alg:unboundedTreeFormula})
+constrains the environment to may-losing states, which also allows for
+recursive learning over the entire tree.
+
+%We compile interpolants into DNF cubes by constructing a temporary BDD. This
+%smaller representation is necessary for maintaining the size of formulas passed
+%to the SAT solver. However, it can occasionally lead to state explosion.
+
+%The recursive learning algorithms, \textsc{learn} (for the controller) and
+%\textsc{\textoverline{learn}} (for the environment), are detailed in
+%Algorithm~\ref{alg:learn}. When executing \textsc{findCandidate} in the
+%unbounded solver, lines 19 and 20 become active and call into the learning
+%procedures.
+
+Figure~\ref{alg:unbounded} shows the main loop of the unbounded synthesis algorithm.
+The algorithm invokes the modified bounded synthesis procedure with increasing bound $k$
+until the initial state is in $B^M$ (environment wins) or $B^m$ reaches a fixed point 
+(controller wins). We prove correctness in the next section.
 
@@
                 \State $B^m[k] \gets \top$
-                \State \Call{checkBound}{$I, k$}
+                \State \Call{checkBound}{$k$}
             \EndFor
@@
         \Ensure \emph{May} and \emph{must} invariants hold
-        \Ensure $\lnot s \in B^m[k]$ if there is a winning controller strategy of length $k$ starting at $s$
-        \Ensure $s \in B^M$ if there is a winning environment strategy of length $k$ starting at $s$
+        \Ensure $I \not\in B^m[k]$ if there exists a winning controller strategy with bound $k$
+        \Ensure $I \in B^M$ if there exists a winning environment strategy with bound $k$
 %%%        $\exists u_{k..1} \forall c_{k..1} \  s(x_k) \land (E_k \lor (\delta(x_k, u_k, c_k) \land E(x_{k-1}) \lor ... \land E(x_1) \lor (\delta(x_1, u_1, c_1) \land E(x_0))...)$
 %%%            $\forall u_{k..1} \exists c_{k..1} \  s(x_k) \land \lnot E_k \land \delta(x_k, u_k, c_k) \land \lnot E(x_{k-1}) \land ... \land \delta(x_1, u_1, c_1) \land \lnot E(x_0)$
-        \Function{checkBound}{$s, k$}
-            \State \Return \Call{solveAbstract}{$\texttt{env}, \langle s, k \rangle, \emptyset$}
+        \Function{checkBound}{$k$}
+            \State \Return \Call{solveAbstract}{$\texttt{env}, I, \emptyset$}
         \EndFunction
@@
 
-%%%During the CEGAR loop of the bounded algorithm, each of the competing solvers
-%%%searches for satisfying assignments of labels to abstract game trees. When the
-%%%abstract game is unsatisfiable this indicates that the states represented by
-%%%nodes in the abstract game tree are losing for the current player. We extract
-%%%these states from the game tree using interpolation.
-
-%%%When executing the unbounded solver, lines 19 and 20 become active in the
-%%%bounded solver. These lines call the learning procedures when the solver fails
-%%%to find a candidate for an abstract game tree. The states symbolically
-%%%represented by nodes in the tree are losing for whichever player could not find
-%%%a winning candidate and can be extracted from the tree using interpolation.
-
-%%%The states in an abstract game with no controller candidate are
-%%%\emph{must-losing}. The environment can always force the game into the error
-%%%states from these states.
-
-%%%From abstractions with no environment candidate we record the complement of
-%%%states in the tree as \emph{may-losing}. The environment cannot reach the error
-%%%state in a number of steps equal to the distance to the bottom of the tree.
-
-%%%We maintain a set of states for each rank up to the current bound. We maintain
-%%%an invariant over these sets via careful construction so that they are
-%%%monotonically increasing by rank. We also ensure that the environment is unable
-%%%to force play from one set to the next. Due to these invariants, when two
-%%%adjacent sets become equivalent we know that the algorithm has reached a fixed
-%%%point and the controller is winning in the unbounded game (line 6).
-
-\subsection{Learning States with Interpolants}
-
-We modify the bounded synthesis algorithm to learn a subset of states from
-failed attempts to find candidate strategies. When \textsc{findCandidate}
-cannot find a strategy for an abstract game tree beginning at a state $s$, we
-know that $s$ is a losing state. We can also attempt to learn additional losing
-states from the tree via interpolation.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{.32\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {$e$}
-                    }
-                    child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {$f$}
-                    }
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-        \end{tikzpicture}
-        \caption{A losing AGT $T$}
-        \label{fig:interpolatetree}
-    \end{subfigure}%
-    \begin{subfigure}[t]{.32\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {$e$}
-                    }
-                    child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {$f$}
-                    }
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-        \end{tikzpicture}
-        \caption{Tree slice $T_1$}
-        \label{fig:treef1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{.32\textwidth}
-        \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] {}
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [left] {$b$}
-                };
-%%%                node [above=4pt] {$\langle s', 1 \rangle$};
-        \end{tikzpicture}
-        \caption{Tree slice $T_2$}
-        \label{fig:treef2}
-    \end{subfigure}
-\end{figure}
-
-Consider the controller-losing game tree, $T$, in
-Figure~\ref{fig:interpolatetree}. We choose any leaf node $l$ of $T$ and its
-parent $n$. We then split the tree at $n$ such that both slices $T_1$ and $T_2$
-contain $n$. Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$
-except $l$, and $T_2$ (Figure~\ref{fig:treef2}) contains only $n$ and $l$ - a
-single round of the game.  There is no candidate strategy for $T$ so $s \land
-\textsc{treeFormula}(T)$ is unsatisfiable, and by extension $s \land
-\textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$ must also be
-unsatisfiable. This enables the construction of an \emph{interpolant} that
-captures losing states at node $n$.
-
-Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
-it is possible to construct a Craig interpolant~\cite{craig1957} $\mathcal{I}$
-such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
-$\mathcal{I}$ refers only to the intersection of variables in $F_1$ and $F_2$.
-An interpolant can be constructed efficiently from a resolution proof of the
-unsatisfiability of $F_1 \land F_2$ \cite{pudlak1997}.
-
-We can construct an interpolant with $F_1 = s \land \textsc{treeFormula}(T_1)$
-and $F_2 = \textsc{treeFormula}(T_2)$. The only variables shared between $F_1$
-and $F_2$ are the state variable copies belonging to node $n$. By construction
-$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
-must lose to the action in the edge between $n$ and $l$, the edge labelled by
-action $b$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
-\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
-following $T_1$ and avoiding the error state.
-
-\begin{algorithm}
-    \caption{Amended tree formulas for Controller and Environment}
-    \label{alg:unboundedTreeFormula}
-    \begin{algorithmic}[1]
-        \Function{treeFormula}{$T$}
-        \If{$\Call{bound}{T} = 0$}
-        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
-        \Else
-        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
-            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
-        }
-        \EndIf
-        \EndFunction
-        \algstore{tf1}
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \algrestore{tf1}
-        \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{bound}{T} = 0$}
-        \State \Return{\Call{E}{$X_T$}}
-        \Else
-        \State \Return{ \Call{$B^m[\Call{bound}{T}]$}{$X_{T}$} $\land$ \\
-            $$\Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
-        \EndIf
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-When we discover losing states for the controller we record them in a set of
-states that \emph{must} be bad, $B^M$. This set is initialised to the error
-state at the beginning of the algorithm and grows as losing states are
-discovered.  The \textsc{treeFormula} function is modified for the unbounded
-solver (see Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with
-$B^M$. This enables further interpolants to be constructed by recursively
-splitting more nodes from $T_1$ since the states that lost to $T_2$ are now
-blocked by $B^M$.
-
-It is more complicated to learn states from environment-losing abstract game
-trees. States are only losing for the environment for a particular bound. For a
-state learned at a node $n$ it is losing at $\textsc{bound}(n)$. Therefore we
-maintain an array of state sets indexed by the bound at which that set of
-states is losing for the environment. Instead of accumulating states, as we do
-in $B^M$, we instead remove states from the array $B^m$. This array is
-initialised with $\top$ and contains only states that \emph{may} lose for the
-controller. The modified \textsc{\textoverline{treeFormula}} function
-constrains the environment to may-losing states, which also allows for
-recursive learning over the entire tree.
-
-We compile interpolants into DNF cubes by constructing a temporary BDD. This
-smaller representation is necessary for maintaining the size of formulas passed
-to the SAT solver. However, it can occasionally lead to state explosion.
-
-The recursive learning algorithms, \textsc{learn} (for the controller) and
-\textsc{\textoverline{learn}} (for the environment), are detailed in
-Algorithm~\ref{alg:learn}. When executing \textsc{findCandidate} in the
-unbounded solver, lines 19 and 20 become active and call into the learning
-procedures. The unbounded algorithm performs bounded synthesis for iteratively
-increasing maximum bounds until the initial state is in $B^M$ (environment
-wins) or $B^m$ reaches a fixed point (controller wins). We prove correctness in
-the next section.
-
 \subsection{Correctness}
-
-\begin{algorithm}
-    \begin{algorithmic}[1]
-        \Require $\sigma(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
-        \Require \emph{Must-invariant} holds
-        \Ensure \emph{Must-invariant} holds
-        \Ensure $\sigma(X_T) \land B^M$
-        \Function{learn}{$\sigma, T$}
-            \IIf{$T = \emptyset$}
-                \Return
-            \EndIIf
-            \State $n \gets $ non-leaf node with min bound
-            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
-            \State $B^M \gets B^M \lor \mathcal{I}$
-            \State \Call{learn}{$\sigma, T_1$}
-        \EndFunction
-        \algstore{learn}
-    \end{algorithmic}
-
-    \begin{algorithmic}[1]
-        \algrestore{learn}
-        \Require $\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
-        \Require \emph{May-invariant} holds
-        \Ensure \emph{May-invariant} holds
-        \Ensure $\sigma(X_T) \land B^m[$\Call{bound}{$T$}$] \equiv \bot$
-        \Function{\textoverline{learn}}{$\sigma, T$}
-            \IIf{$T = \emptyset$}
-                \Return
-            \EndIIf
-            \State $n \gets $ non-leaf node with min bound
-            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$\sigma(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
-            \For{$i = 1$ to \Call{bound}{$n$}}
-                \State $B^m[i] \gets B^m[i] \land \lnot \mathcal{I}$
-            \EndFor
-            \State \Call{\textoverline{learn}}{$\sigma, T_1$}
-        \EndFunction
-    \end{algorithmic}
-    \caption{Learning algorithms}
-    \label{alg:learn}
-\end{algorithm}
 
@@
 Line~(11) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
-1.  Consider formulas $F_1=\sigma(X_T) \land
+1.  Consider formulas $F_1=s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ and $F_2 =
@@
 $X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
-any solution of $F_1 \land F_2$ also satisfies $\sigma(X_T) \land
+any solution of $F_1 \land F_2$ also satisfies $s(X_T) \land
 \textsc{treeFormula}(T)$, which is unsatisfiable (precondition (b)).  Hence the
@@
 Intuitively, the interpolant computed in line~(13) overapproximates the set of
-states reachable from $\sigma$ by following the tree from the root node to $n$,
+states reachable from $s$ by following the tree from the root node to $n$,
 and underapproximates the set of states from which the environment loses
@@
 Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any
-assignment to $X_n$ that satisfies $\sigma(X_T) \land
+assignment to $X_n$ that satisfies $s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
-removing $\II$ from $B^m[\textsc{bound}(n)]$ makes $\sigma(X_T) \land
+removing $\II$ from $B^m[\textsc{bound}(n)]$ makes $s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
@@
 is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
-\equiv B^m[\textsc{bound}(T)](X^T)$; hence $\sigma(X_T) \land
-\textsc{\textoverline{treeFormula}}(T_1) \equiv \sigma(X_T) \land
+\equiv B^m[\textsc{bound}(T)](X^T)$; hence $s(X_T) \land
+\textsc{\textoverline{treeFormula}}(T_1) \equiv s(X_T) \land
 B^m[\textsc{bound}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
@@
 The proof of \textsc{learn} is similar to the above proof of \textsc{learn}. An
-interpolant constructed from $F_1=\sigma(X_T) \land \textsc{treeFormula}(T_1)$
+interpolant constructed from $F_1=s(X_T) \land \textsc{treeFormula}(T_1)$
 and $F_2 = \textsc{treeFormula}(T_2)$ has the property $\II \land F_2 \equiv
@@
 tree $T_1$ and root node $n$: $\textsc{treeFormula}(T_1) \equiv \lnot
-B^M(X_T)$.  Hence $\sigma(X_T) \land \textsc{treeFormula}(T_1) \equiv
-\sigma(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $\sigma(X_T) \land
+B^M(X_T)$.  Hence $s(X_T) \land \textsc{treeFormula}(T_1) \equiv
+s(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $s(X_T) \land
 B^M(X_T)$, the second postcondition, is true.
@@
 The optimisation is relatively simple and is inspired by a common greedy
-heuristic for minimising $\texttt{unsat}$ cores. $I$ is a value assignment to
-each variable in $X$. If the environment does not win for $\langle I, k
+heuristic for minimising $\texttt{unsat}$ cores. Initial state $I$ assigns a value to
+each variable in $X$. If the environment loses $\langle I, k
 \rangle$ then we attempt to solve for a generalised version of $I$ by removing
-one assignment at a time. If the controller can win from the larger set of
-states then we continue generalising without it. In this way we learn more
+one variable assignment at a time. If the environment loses from the larger set of
+states then we continue generalising. In this way we learn more
 states by increasing the reachable set. In our benchmarks we have observed that
@@
     \begin{algorithmic}
-        \Function{checkBound}{$s, k$}
-            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, s, k, \emptyset$}
+        \Function{checkBound}{$k$}
+            \State $r \gets $ \Call{solveAbstract}{$\texttt{env}, I, k, \emptyset$}
             \IIf{$r \neq \emptyset$} \Return $r$ \EndIIf
-            \State $s' \gets s$
+            \State $s' \gets I$
             \For{$x \in X$}
@@
 The results of our benchmarking are shown, along with the synthesis competition
-results~\cite{syntcompedacc}, in Figure~\ref{fig:syntcomp}. The competition was run on Intel Quad
-Core 32Ghz CPUs with 32GB of memory, also on isolated nodes for one hour per
+results~\cite{syntcompedacc}, in Table~\ref{tab:syntcomp}. The competition was run on Intel Quad
+Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes for one hour per
 instance.
 
-\begin{figure}
+\begin{table}
     \centering
@@
     \caption{Synthesis Competition 2015 Results}
-    \label{fig:syntcomp}
-\end{figure}
-
-Our implementation was able to solver 103 out of the 250 specification in the
-alloted time, including 12 instances that were unable to be solved by any other
+    \label{tab:syntcomp}
+\end{table}
+
+Our implementation was able to solve 103 out of the 250 specification in the
+alloted time, including 12 instances that were not solved by any other
 solver in SYNTCOMP'15. The unique instances we solved are listed in
-Figure~\ref{fig:unique}.
-
-\begin{figure}[h]
+Table~\ref{tab:unique}.
+
+\begin{table}[h]
     \centering
@@
     \caption{Instances uniquely solved by our approach}
-    \label{fig:unique}
-\end{figure}
+    \label{tab:unique}
+\end{table}
 
@@
 
-We also present a cactus plot of the number of instances solved over time. For
+We also present a cactus plot of the number of instances solved over time (Figure~\ref{fig:cactus}). For
 reference we plotted our results along with \textsc{Demiurge}~\cite{bloem2014},
@@
     \end{tikzpicture}
+    \caption{Number of instances solver over time by the unbounded solver, Demiurge,
+    and Simple BDD Solver.}\label{fig:cactus}
 \end{figure}
@@
 portfolio solver. In the parallel track of the competition, \textsc{Demiurge}
-competes with a suite of 3 separate but communicating solvers. The solvers
+uses a suite of 3 separate but communicating solvers. The solvers
 relay unsafe states to one another, which is compatible with the set $B^M$ in
@@
 In \cite{een2015} the realisability algorithm is extended with strategy
-extraction. The technique relies on interpolation of abstract game trees.  In
-the present work we use interpolation in a similar way for computational
-learning.
+extraction. The technique relies on interpolation over abstract game trees to
+compute the winning strategy.  In
+the present work we use interpolation in a different way in order to learn
+losing states of the game.
 

### 2016-01-29 paper.tex
minor

@@
 \textsc{findCandidate} cannot find a candidate strategy for an abstract game tree with initial
-state $s$.  We know that $s$ is a losing state.  We can learn additional losing
+state $s$.  At this point we have established that $s$ is a losing state.  We can learn additional losing
 states from the tree via interpolation.  This is achieved in lines~18--20 in Algorithm~\ref{alg:bounded},
@@
 \textsc{treeFormula}(T)$ is unsatisfiable.  By construction, 
-$\textsc{treeFormula}(T) = \textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$
+$\textsc{treeFormula}(T) \equiv \textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$
 and hence $s \land
@@
 
-We can construct an interpolant with $F_1 = s \land \textsc{treeFormula}(T_1)$
+We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(T_1)$
 and $F_2 = \textsc{treeFormula}(T_2)$ (line~5). The only variables shared between $F_1$
-and $F_2$ are the state variable copies belonging to node $n$. By construction
+and $F_2$ are the state variable copies belonging to node $n$. By the properties of the interpolant,
 $F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
@@
 are initialized to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
-is discovered for a node $n$ with bound $\textsc{bound}(n)$ in line~13, Algorithm~\ref{alg:learn}, this set is subtracted from 
+is discovered for a node $n$ with bound $\textsc{bound}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
 $B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
@@
 %procedures.
+
+\subsection{Main synthesis loop}
 
@@
 
+Both invariants trivially hold after $B^m$ and $B^M$ have been initialized in the
+beginning of the algorithm.
 The sets $B^m$ and $B^M$ are only modified by the functions \textsc{learn} and
-\textoverline{\textsc{learn}}.  Below we prove that these functions indeed
-maintain the invariants.
-
-\subsection{Proof of learning functions}
+\textoverline{\textsc{learn}}.  Below we prove that \textoverline{\textsc{learn}} 
+maintains the invariants.  The proof of \textsc{learn} is similar.
+
+\subsection{Proof of \textoverline{\textsc{learn}}}
 

### 2016-01-29 paper.tex
Edits to diagrams & tables

@@
                 }
-                child {node [circle,draw] {}
+                child {node [circle,draw] (n) {}
                     child {node [circle,draw,right=5pt] {}
@@
                     }
-                    edge from parent [fixed] node [right] {$d$}
+                    edge from parent [fixed] {node [right] {$d$}}
                 }
                 node [above=4pt] {$\langle s, 2 \rangle$};
+            \node [right=1pt of n] {$n$};
         \end{tikzpicture}
@@
                 node [above=4pt] {$\langle s, 2 \rangle$};
+            \node [right=1pt of n] {$n$};
         \end{tikzpicture}
@@
             \node [circle,draw] {}
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [left] {$b$}
-                };
-%%%                node [above=4pt] {$\langle s', 1 \rangle$};
+                child {node [circle,draw,right=5pt] {}
+                    edge from parent [fixed] node [left] {$e$}
+                }
+                child {node [circle,draw,left=5pt] {}
+                    edge from parent [fixed] node [right] {$f$}
+                }
+                node [right=5pt] {$n$};
         \end{tikzpicture}
@@
         Aisy & 98 & \\
-        \hline
-        Unbounded Synthesis & 103 & 12 \\
+        \textit{Unbounded Synthesis} & \textit{103} & \textit{12} \\
     \end{tabular}
@@
     \begin{tabular}{l l}
-        \texttt{6s216rb0\_c0to31} & \texttt{driver\_c10n} \\
-        \texttt{cnt30y} &  \texttt{stay18y} \\
-        \texttt{driver\_a10n} & \texttt{stay20n} \\
-        \texttt{driver\_a8n} & \texttt{stay20y} \\
-        \texttt{driver\_b10y} & \texttt{stay22n} \\
-        \texttt{driver\_b8y} & \texttt{stay22y} \\
+        1. \texttt{6s216rb0\_c0to31} & 7. \texttt{driver\_c10n} \\
+        2. \texttt{cnt30y} &  8. \texttt{stay18y} \\
+        3. \texttt{driver\_a10n} & 9. \texttt{stay20n} \\
+        4. \texttt{driver\_a8n} & 10. \texttt{stay20y} \\
+        5. \texttt{driver\_b10y} & 11. \texttt{stay22n} \\
+        6. \texttt{driver\_b8y} & 12. \texttt{stay22y} \\
     \end{tabular}

### 2016-01-29 paper.tex
minor
Definitions

@@
 the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
-of all nodes of a tree $T$ and $\textsc{leaves}(T)$ the subset of leaf nodes.
+of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes, for
+edge $e$ $\textsc{action}(e)$ is the action that labels the edge, and for node
+$n$ $\textsc{height}(n)$ as the maximal distance from n to a leaf. For a node
+$n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
+where $n'$ is a child node of $n$ and $e$ is the edge connecting $n$ and $n'$.
 
@@
 in Figure~\ref{fig:extended}. Then copies of state and action variables are
-introduced for each node in tree and opponent action variables for each edge.
-Opponent actions are set to $\textsc{action}(e)$, the action that labels the
-edge of the tree $e$. The state and action variables of node $n$ are connected
-to successor nodes $\textsc{succ}(n)$ by an encoding of the transition relation
-and constrained to the winning condition of the player.
+introduced for each node in tree and opponent action variables for each edge
+$e$ set to $\textsc{action}(e)$. The state and action variables of node $n$ are
+connected to successor nodes $\textsc{succ}(n)$ by an encoding of the
+transition relation and constrained to the winning condition of the player.
 
@@
         \Function{treeFormula}{$T$}
-        \If{$\Call{bound}{T} = 0$}
+        \If{$\Call{height}{T} = 0$}
         \State \Return{ $\lnot \Call{E}{X_{T}}$ }
@@
         \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{bound}{T} = 0$}
+        \If{$\Call{height}{T} = 0$}
         \State \Return{\Call{E}{$X_{T}$}}
@@
             \EndIIf
-            \State $n \gets $ non-leaf node with min bound
+            \State $n \gets $ non-leaf node with min height 
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
@@
         \Ensure \emph{May-invariant} holds
-        \Ensure $s(X_T) \land B^m[$\Call{bound}{$T$}$] \equiv \bot$
+        \Ensure $s(X_T) \land B^m[$\Call{height}{$T$}$] \equiv \bot$
         \Function{\textoverline{learn}}{$s, T$}
@@
             \EndIIf
-            \State $n \gets $ non-leaf node with min bound
+            \State $n \gets $ non-leaf node with min height 
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
             \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
-            \For{$i = 1$ to \Call{bound}{$n$}}
+            \For{$i = 1$ to \Call{height}{$n$}}
                 \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
@@
         \Function{treeFormula}{$T$}
-        \If{$\Call{bound}{T} = 0$}
+        \If{$\Call{height}{T} = 0$}
         \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
@@
         \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{bound}{T} = 0$}
+        \If{$\Call{height}{T} = 0$}
         \State \Return{\Call{E}{$X_T$}}
         \Else
-        \State \Return{ \Call{$B^m[\Call{bound}{T}]$}{$X_{T}$} $\land$ \\
+        \State \Return{ \Call{$B^m[\Call{height}{T}]$}{$X_{T}$} $\land$ \\
             $$\Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
@@
 are initialized to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
-is discovered for a node $n$ with bound $\textsc{bound}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
+is discovered for a node $n$ with bound $\textsc{height}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
 $B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
@@
 of depth 1, this means that the environment cannot force the game into
-$B^m[\textsc{bound}(n)]$ playing against the counterexample moves in $T_2$.
-Hence, $\II \cap Upre(B^m[\textsc{bound}(n)]) = \emptyset$.  Furthermore, since
+$B^m[\textsc{height}(n)]$ playing against the counterexample moves in $T_2$.
+Hence, $\II \cap Upre(B^m[\textsc{height}(n)]) = \emptyset$.  Furthermore, since
 the may-invariant holds, $\II \cap Upre(i) = \emptyset$, for all $i <
-\textsc{bound}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
-\textsc{bound}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
+\textsc{height}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
+\textsc{height}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
 first post-condition.
@@
 \textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
-removing $\II$ from $B^m[\textsc{bound}(n)]$ makes $s(X_T) \land
+removing $\II$ from $B^m[\textsc{height}(n)]$ makes $s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
@@
 is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
-\equiv B^m[\textsc{bound}(T)](X^T)$; hence $s(X_T) \land
+\equiv B^m[\textsc{height}(T)](X^T)$; hence $s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1) \equiv s(X_T) \land
-B^m[\textsc{bound}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
+B^m[\textsc{height}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
 \textsc{\textoverline{learn}} holds.

### 2016-01-29 paper.tex
Redid change
minor

@@
 the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
-of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes, for
-edge $e$ $\textsc{action}(e)$ is the action that labels the edge, and for node
-$n$ $\textsc{height}(n)$ as the maximal distance from n to a leaf. For a node
+of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
+edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
+$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. For node
 $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
@@
         \captionsetup{width=.8\textwidth}
-        \caption{Extended AGT}
+        \caption{AGT extended to bound 3 with arbitrary controller actions}
         \label{fig:extended}
     \end{subfigure}
-    \caption{Abstract game trees}
+    \caption{Abstract game trees.}
     \label{fig:alltrees}

### 2016-02-25 paper.tex
minor
Loop optimisation

@@
 edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
-$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. For node
+$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. $\textsc{height}(T)$ is the height 
+of the root node of the tree.  For node
 $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
@@
     \end{algorithmic}
+
     \caption{Generalise $I$ optimisation}
     \label{alg:opt1}
+\end{algorithm}
+
+\subsection{Optimisation: Encouraging loops}
+
+This optimisation is designed to encourage the system player to choose a
+candidate strategy that contains loops. If the controller's strategy forces the
+game into an error free loop, then the system wins the game. This optimisation
+does not attempt to find such a strategy but instead optimistically searches
+for candidate strategies that may loop. The modifications add formulas to the
+SAT call when the controller is playing. These formulas represent loops of a
+particular length in a branch of the tree. The SAT solver is called multiple
+times for each branch in the game tree. The search begins by looking for a loop
+in the first two game rounds of the branch and successively relaxing this
+condition to longer loop lengths. If a loop is found or we reach the end of the
+branch then the algorithm moves on to greedily find more loops in other
+branches.
+
+\begin{algorithm}
+    \begin{algorithmic}
+        \Function{loopFmls}{$T$}
+            \State $i \gets 0$
+            \For{$l \in leaves(gt)$}
+                \State $j \gets 0$
+                \For{$n \in$ \Call{parents}{$l$}}
+                    \For{$p \in$ \Call{parents}{$n$}}
+                        \State $loopFmls[i][j] \gets X_n \equiv X_p$
+                    \EndFor
+                    \State $j \gets j + 1$
+                \EndFor
+                \State $i \gets i + 1$
+            \EndFor
+            \State \Return $loopFmls$
+        \EndFunction
+    \end{algorithmic}
+
+    \begin{algorithmic}
+        \Function{findCandidate}{$p, s, k, T$}
+        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
+            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
+            \If{$p = \texttt{cont}$}
+                \State $loopFmls \gets $ \Call{loopFmls}{$\hat{T}$}
+                \State $currentLoops \gets \top$
+                \For{$i \in loopFmls$}
+                    \For{$j \in loopFmls[i]$}
+                        \State $sol \gets $ \Call{SAT}{$loopFmls[i][j] \land currentLoops \land s(X_{\hat{T}}) \land f$}
+                        \If{$sol = \texttt{sat}$}
+                            \State $currentLoops \gets currentLoops \land loopFmls[i][j]$
+                            \State \textbf{break}
+                        \EndIf
+                    \EndFor
+                \EndFor
+            \Else
+                \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
+            \EndIf
+            \If{$sol = \texttt{unsat}$} 
+                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
+                    %\State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
+                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$s, \hat{T}$}}{\Call{\textoverline{learn}}{$s, \hat{T}$}} \EndIfElse
+                \EndIf
+                \State \Return $\emptyset$ \Comment{No candidate exists}
+            \Else
+                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
+            \EndIf
+        \EndFunction
+    \end{algorithmic}
+
+    \caption{Loop Encouragement}
+    \label{alg:opt2}
 \end{algorithm}

### 2016-03-24 paper.tex
Some minor fixes for reviewer comments

--- 2016-01-29 paper.tex
@@
     this technique to unbounded synthesis.  Our method applies Craig
-    interpolation to abstract game trees produced by counterexample-guided
+    interpolation to abstract game trees produced by counterexample guided
     search in order to construct a monotonic sequence of may-losing regions.
@@
 
-The reactive synthesis problem is 2EXPTIME-complete so
+The reactive synthesis problem is EXPTIME-complete so
 naive algorithms are infeasible on anything but simple systems.
@@
 in either symbolic or explicit form, the set of winning states.
-Instead, it uses counterexample-guided backtracking search to identify a small subset
+Instead, it uses counterexample guided backtracking search to identify a small subset
 of runs that are sufficient to solve the game.
@@
 can keep the game within the safe region for a bounded number of steps.  
-This is insufficient in most practical situations that require unbounded realizability.
+This is insufficient in most practical situations that require unbounded realisability.
 
 %Previous work has applied this idea to
-%realizability of bounded games \cite{narodytska2014} by forming abstractions of
+%realisability of bounded games \cite{narodytska2014} by forming abstractions of
 %the game and refining in a counterexample-guided framework.
@@
 The learning procedure ensures that reaching a fixed point in these sets is sufficient to establish 
-unbounded realizability.
-Our method can be seen as a hybrid between the counterexample-guided algorithm
+unbounded realisability.
+Our method can be seen as a hybrid between the counterexample guided algorithm
 by Narodytska et al. and methods based on losing set computation: we use the
@@
 
-A \emph{controller strategy} $\pi^c : (X, U) \to C$ is a mapping of states and
+A \emph{controller strategy} $\pi^c : (2^X, 2^U) \to 2^C$ is a mapping of states and
 uncontrollable inputs to controllable actions. A controller strategy is winning
@@
 u_0)), (x_1, u_1, \pi^c(x_1, u_1)) \dots$ $(x_n, u_n, \pi^c(x_n, u_n))$ are
-winning.  Bounded \emph{realizability} is the problem of determining the
+winning.  Bounded \emph{realisability} is the problem of determining the
 existence of such a strategy for a bounded game.
 
-An \emph{environment strategy} $\pi^e : X \to U$ is a mapping of states to
+An \emph{environment strategy} $\pi^e : 2^X \to 2^U$ is a mapping of states to
 uncontrollable actions. A bounded run is winning for the environment if $x_0 =
 I \land \exists i \in \{1..n\} (E(x_i))$ and an environment strategy is winning
-for a bounded game if all runs $(x_0, \pi^e(x_1), c_1), (x_1, \pi^e(x_1), c_1)
+for a bounded game if all runs $(x_0, \pi^e(x_0), c_0), (x_1, \pi^e(x_1), c_1)
 \dots (x_n, \pi^e(x_n), c_n)$ are winning for the environment.  Safety games
@@
 
-\subsection{Counterexample-guided bounded synthesis}
+\subsection{Counterexample guided bounded synthesis}
 
@@
 Specifically, we consider abstractions represented as trees of actions,
-referred to as \emph{abstract game trees}.  Figure~\ref{fig:agt} shows an
+referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows an
 example abstract game tree for the controller (abstract game trees for the
@@
 solver to search for candidate partial strategies. Candidate strategies are
-checked and by a call to the opponent solver, which also refines the abstract
+checked by a call to the opponent solver, which also refines the abstract
 game tree. The refinements are then solved recursively.  The full procedure is
@@
         \Loop
-            \IIf{$cand = \emptyset $} \Return $\emptyset $ \EndIIf \Comment{No candidate: return with no solution}
+            \IIf{$cand = \texttt{NULL}$} \Return $\texttt{NULL}$ \EndIIf \Comment{No candidate: return with no solution}
             \State $\langle cex, l, u \rangle \gets $ \Call{verify}{$p, s, k, T, cand$} \Comment{Verify candidate}
@@
                 \EndIf
-                \State \Return $\emptyset$ \Comment{No candidate exists}
+                \State \Return $\texttt{NULL}$ \Comment{No candidate exists}
             \Else
@@
             \State $\langle k', s'\rangle \gets $ \Call{outcome}{$s, k, cand, l$} \Comment{Get bound and state at leaf}
-                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $k'$, $s'$, $\emptyset$} \Comment{Solve for the opponent}
-                \IIf{$u \neq \emptyset$} \Return $\langle \False, l, u \rangle$ \EndIIf \Comment{Return counterexample}
+                \State $u \gets $ \Call{solveAbstract}{\Call{opponent}{$p$}, $s'$, $k'$, $\emptyset$} \Comment{Solve for the opponent}
+                \IIf{$u \neq \texttt{NULL}$} \Return $\langle \True, l, u \rangle$ \EndIIf \Comment{Return counterexample}
             \EndFor
-            \State \Return $\langle \True, \emptyset, \emptyset \rangle$
+            \State \Return $\langle \False, \emptyset, \emptyset \rangle$
         \EndFunction
@@
 with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
-in Figure~\ref{fig:extended}. Then copies of state and action variables are
-introduced for each node in tree and opponent action variables for each edge
-$e$ set to $\textsc{action}(e)$. The state and action variables of node $n$ are
-connected to successor nodes $\textsc{succ}(n)$ by an encoding of the
-transition relation and constrained to the winning condition of the player.
+in Figure~\ref{fig:extended}.  For each node in the tree, new SAT variables are
+introduced corresponding to the state ($X_T$) and action ($U_T$ or $C_T$)
+variables of that node. Additional variables for the opponent actions in the
+edges of $T$ are introduced ($U_e$ or $C_e$) and set to $\textsc{action}(e)$.
+The state and action variables of node $n$ are connected to successor nodes
+$\textsc{succ}(n)$ by an encoding of the transition relation and constrained to
+the winning condition of the player.
+
+%%%Then copies of state and action variables are introduced for each node in the
+%%%tree and opponent action variables for each edge $e$ are set to
+%%%$\textsc{action}(e)$.
 
@@
         \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
-            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_T, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
         }
@@
         \State \Return{ $\Call{E}{X_{T}} \lor$ \\
-        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
+        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_T, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
         \EndIf
@@
 We instead present an approach that iteratively solves games of increasing
-bound while learning winning states from abstract games using 
-interpolation. We show
-that reaching a fixed point in learned states is sufficient for completeness.
+bound while learning winning states from abstract games using interpolation. We
+show that reaching a fixed point in learned states is sufficient for
+completeness.
 
@@
 
-We extend the bounded synthesis algorithm to learn states losing for one of the players from
-failed attempts to find candidate strategies.  The learning procedure kicks in whenever
-\textsc{findCandidate} cannot find a candidate strategy for an abstract game tree with initial
-state $s$.  At this point we have established that $s$ is a losing state.  We can learn additional losing
-states from the tree via interpolation.  This is achieved in lines~18--20 in Algorithm~\ref{alg:bounded},
-enabled in the unbounded version of the algorithm, which invoke \textsc{learn} or
-\textsc{\textoverline{learn}} to learn controller or environment losing states respectively
+We extend the bounded synthesis algorithm to learn states losing for one of the
+players from failed attempts to find candidate strategies.  The learning
+procedure kicks in whenever \textsc{findCandidate} cannot find a candidate
+strategy for an abstract game tree $T$ with initial state $s$ and bound $k$.
+For the system, this means the environment partial strategy represented by $T$
+will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
+bound is irrelevant).  For the environment, the system can avoid $E$ for $k$
+rounds no matter the assignments to $U$ variables. Hence, we have established
+that $s$ is a losing state either always (for the system) or for bound $k$ (for
+the environment. We can learn additional losing states from the tree via
+interpolation.  This is achieved in lines~18--20 in
+Algorithm~\ref{alg:bounded}, enabled in the unbounded version of the algorithm,
+which invoke \textsc{learn} or \textsc{\textoverline{learn}} to learn
+controller or environment losing states respectively
 (Algorithm~\ref{alg:learn}).
@@
                 child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {$b$}
+                    }
                     edge from parent [fixed] node [left] {$a$}
@@
                 child {node [circle,draw] (n) {}
-                    child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {$e$}
-                    }
-                    child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {$f$}
-                    }
                     edge from parent [fixed] node [right] {$d$}
@@
         \Ensure \emph{Must-invariant} holds
-        \Ensure $s(X_T) \land B^M$
+        \Ensure $s(X_T) \land B^M \not\equiv \bot$
+        \Comment $s$ will be added to $B^M$
         \Function{learn}{$s, T$}
@@
         \Ensure $s(X_T) \land B^m[$\Call{height}{$T$}$] \equiv \bot$
+        \Comment{$s$ will be removed from $B^m[$\Call{height}{$T$}$]$}
         \Function{\textoverline{learn}}{$s, T$}
@@
 actions.  Assume that there does not exist a partial strategy for the controller
-winning against this tree and that the \textsc{findCandidate} function could not
-find a candidate strategy.
-We choose a non-leaf node $n$ of $T$ with maximal depth, i.e., a node whose children
-are leafs (Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that both slices $T_1$ and $T_2$
-contain a copy of $n$ (line~4). Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$
-except $n$'s children, and $T_2$ (Figure~\ref{fig:treef2}), which contains only $n$ and its children.  
-There is no candidate strategy for $T$ so $s \land
+winning against this tree and that the \textsc{findCandidate} function could
+not find a candidate strategy. We choose a non-leaf node $n$ of $T$ with
+maximal depth, i.e., a node whose children are leafs
+(Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that
+both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
+Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$ except $n$'s
+children, and $T_2$ (Figure~\ref{fig:treef2}), which contains only $n$ and its
+children.  There is no candidate strategy for $T$ so $s \land
 \textsc{treeFormula}(T)$ is unsatisfiable.  By construction,
-$\textsc{treeFormula}(T) \equiv \textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$
-and hence $s \land
-\textsc{treeFormula}(T_1) \land \textsc{treeFormula}(T_2)$ is also 
-unsatisfiable. This enables the construction of an \emph{interpolant} that
-captures losing states at node $n$.
+$\textsc{treeFormula}(T) \equiv \textsc{treeFormula}(T_1) \land
+\textsc{treeFormula}(T_2)$ and hence $s \land \textsc{treeFormula}(T_1) \land
+\textsc{treeFormula}(T_2)$ is also unsatisfiable. This enables the construction
+of an \emph{interpolant} that captures losing states at node $n$.
 
@@
 
-We have discovered a set $\mathcal{I}$ of states losing for the controller.  
-We record these newly discovered states in a set $B^M$ of
-``bad'' states that \emph{must} be losing for the controller (Algorithm~\ref{alg:learn}, line~6). This set is initialised to the error
-set $E$ and grows as new losing states are
+We have discovered a set $\mathcal{I}$ of states losing for the controller. We
+record these newly discovered states in a set $B^M$ of ``bad'' states that
+\emph{must} be losing for the controller (Algorithm~\ref{alg:learn}, line~6).
+This set is initialised to the error set $E$ and grows as new losing states are
 discovered.  The \textsc{treeFormula} function is modified for the unbounded
-solver (Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with
-$B^M$. This enables further interpolants to be constructed by the learning procedure 
-recursively splitting more nodes from $T_1$ (Algorithm~\ref{alg:learn}, line~7) since the states that lost to $T_2$ are now
-blocked by $B^M$.
+solver (Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with $B^M$.
+This enables further interpolants to be constructed by the learning procedure
+recursively splitting more nodes from $T_1$ (Algorithm~\ref{alg:learn}, line~7)
+since the states that are losing to $T_2$ are now blocked by $B^M$.
 
@@
         \State \Return{ \Call{$B^m[\Call{height}{T}]$}{$X_{T}$} $\land$ \\
-            $$\Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
+            $$\bigg( \Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})\bigg)$$ }
         \EndIf
@@
 sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.  $B^m[k]$ 
-are initialized to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
+are initialised to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
 is discovered for a node $n$ with bound $\textsc{height}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
@@
 %controller. 
-The modified \textsc{\textoverline{treeFormula}} function (\ref{alg:unboundedTreeFormula})
+The modified \textsc{\textoverline{treeFormula}} function (Algorithm~\ref{alg:unboundedTreeFormula})
 constrains the environment to may-losing states, which also allows for
@@
                 \EndIIf
-                \IIf{$\exists i \  B^m[i] \equiv B^m[i+1]$}
-                    \Return \texttt{realisable} \Comment{Reached fixed point}
+                \IIf{$\exists i < k . \  B^m[i] \equiv B^m[i+1]$} \Comment{Reached fixed point}
+                    \State \hspace{\algorithmicindent} \Return \texttt{realisable} 
                 \EndIIf
@@
         \Function{checkBound}{$k$}
-            \State \Return \Call{solveAbstract}{$\texttt{env}, I, \emptyset$}
+            \State \Return \Call{solveAbstract}{$\texttt{env}, I, k, \emptyset$}
         \EndFunction
@@
 We define two global invariants of the algorithm.  The \emph{may-invariant}
-states that sets $B^m[i]$ grows monotonically with $i$ and that each $B^m[i+1]$
+states that sets $B^m[i]$ grow monotonically with $i$ and that each $B^m[i+1]$
 overapproximates the states from which the environment can force the game into
@@
 
-Both invariants trivially hold after $B^m$ and $B^M$ have been initialized in the
+Both invariants trivially hold after $B^m$ and $B^M$ have been initialised in the
 beginning of the algorithm.
@@
 
-Line~(11) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
+Line~(11--12) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
 1.  Consider formulas $F_1=s(X_T) \land
@@
 any solution of $F_1 \land F_2$ also satisfies $s(X_T) \land
-\textsc{treeFormula}(T)$, which is unsatisfiable (precondition (b)).  Hence the
+\textsc{\textoverline{treeFormula}}(T)$, which is unsatisfiable (precondition (b)).  Hence the
 interpolation operation is defined for $F_1$ and $F_2$.
@@
 of depth 1, this means that the environment cannot force the game into
-$B^m[\textsc{height}(n)]$ playing against the counterexample moves in $T_2$.
-Hence, $\II \cap Upre(B^m[\textsc{height}(n)]) = \emptyset$.  Furthermore, since
-the may-invariant holds, $\II \cap Upre(i) = \emptyset$, for all $i <
-\textsc{height}(n)$.  Hence, removing $\II$ from all $B^m[i], i\leq
-\textsc{height}(n)$ in line~(15) preserves the may-invariant, thus satisfying the
-first post-condition.
+$B^m[\textsc{height}(n)-1]$ playing against the counterexample moves in $T_2$.
+Hence, $\II \cap Upre(B^m[\textsc{height}(n)-1]) = \emptyset$.  Furthermore,
+since the may-invariant holds, $\II \cap Upre(B^m[\textsc{height}(i)]) =
+\emptyset$, for all $i < \textsc{height}(n)$.  Hence, removing $\II$ from all
+$B^m[i], i\leq \textsc{height}(n)$ in line~(15) preserves the may-invariant,
+thus satisfying the first post-condition.
 
@@
 is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
-\equiv B^m[\textsc{height}(T)](X^T)$; hence $s(X_T) \land
+\equiv B^m[\textsc{height}(T_1)](X^T)$; hence $s(X_T) \land
 \textsc{\textoverline{treeFormula}}(T_1) \equiv s(X_T) \land
-B^m[\textsc{height}(T)](X^T) \equiv \bot$.  Thus the second postcondition of
+B^m[\textsc{height}(T_1)](X^T) \equiv \bot$.  Thus the second postcondition of
 \textsc{\textoverline{learn}} holds.
@@
 s(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $s(X_T) \land
-B^M(X_T)$, the second postcondition, is true.
+B^M(X_T) \not\equiv \bot$, the second postcondition, is true.
 
@@
 We must prove that $\textsc{checkBound}$ terminates and that upon termination
-its postcondition holds, i.e., state $s$ is removed from $B^m[\kappa]$ if there
+its postcondition holds, i.e., state $I$ is removed from $B^m[\kappa]$ if there
 is a winning controller strategy on the bounded safety game of maximum bound
 $\kappa$ or it is added to $B^M$ otherwise. Termination follows from
-completeness of counterexample-guided search, which terminates after
+completeness of counterexample guided search, which terminates after
 enumerating all possible opponent moves in the worst case.
@@
 bound $\kappa$ for which the environment cannot force into $E$. The algorithm
-then invokes the \textsc{\textoverline{learn}} method, which removes $s$ from
+then invokes the \textsc{\textoverline{learn}} method, which removes $I$ from
 $B^m[\kappa]$.  Alternatively, if there is a winning strategy for the
 environment at bound $\kappa$ then a counterexample losing for the controller
-will be found.  Subsequently \textsc{learn} will be called and $s$ added to
+will be found.  Subsequently \textsc{learn} will be called and $I$ added to
 $B^M$.
@@
             \State $r \gets$ \Call{solveAbstract}{$\texttt{env}, s' \setminus \{x\}, k, \emptyset$} 
-                \IIf{$r = \emptyset$} $s' \gets s' \setminus \{x\}$ \EndIIf \Comment{Removes the assignment to $x$ from $s'$}
+                \IIf{$r = \texttt{NULL}$} $s' \gets s' \setminus \{x\}$ \EndIIf \Comment{Removes the assignment to $x$ from $s'$}
             \EndFor
-            \State \Return $\emptyset$
+            \State \Return $\texttt{NULL}$
         \EndFunction
@@
 interpretation \cite{walker2014,brenguier2014}. Satisfiability solving has been used
-previously for synthesis in a suite of methods proposed by Bloem et al
-\cite{bloem2014}. The authors propose a propose employing competing SAT solvers
+previously for synthesis in a suite of methods proposed by Bloem et al.
+\cite{bloem2014}. The authors propose employing competing SAT solvers
 to learn clauses, which is similar to our approach but does not unroll the
 game. They also suggest QBF solver, template-based, and Effectively
-Propositional Logic (EPL) approaches.
+Propositional Logic (EPR) approaches.
 
@@
 synthesis \cite{ehlers2010}. Lazy synthesis \cite{finkbeiner2012} is a
-counterexample-guided approach to bounded synthesis that refines an
+counterexample guided approach to bounded synthesis that refines an
 implementation for the game instead of an abstraction of it.

### 2016-03-24 paper.tex
Added k as an input to height

@@
 edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
-$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. $\textsc{height}(T)$ is the height 
-of the root node of the tree.  For node
-$n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
-where $n'$ is a child node of $n$ and $e$ is the edge connecting $n$ and $n'$.
+$n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a game
+bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root node
+of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs
+$\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the edge
+connecting $n$ and $n'$.
 
@@
         \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
-            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
+            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$k, \hat{T}$}}{\Call{\textoverline{treeFormula}}{$k, \hat{T}$}} \EndIfElse
             \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
@@
     \begin{algorithmic}[1]
-        \Function{treeFormula}{$T$}
-        \If{$\Call{height}{T} = 0$}
+        \Function{treeFormula}{$k, T$}
+        \If{$\Call{height}{k, T} = 0$}
         \State \Return{ $\lnot \Call{E}{X_{T}}$ }
@@
         \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
-            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_T, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_T, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{k, n})$$
         }
@@
         \algrestore{tf1}
-        \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{height}{T} = 0$}
+        \Function{\textoverline{treeFormula}}{$k, T$}
+        \If{$\Call{height}{k, T} = 0$}
         \State \Return{\Call{E}{$X_{T}$}}
@@
         \State \Return{ $\Call{E}{X_{T}} \lor$ \\
-        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_T, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
+        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_T, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{k, n})$$ }
         \EndIf
@@
     \begin{algorithmic}[1]
-        \Require $s(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
+        \Require $s(X_T) \land $ \Call{treeFormula}{$k, T$} $\equiv \bot$
         \Require \emph{Must-invariant} holds
@@
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$k, T_1$}, \Call{treeFormula}{$k, T_2$}}
             \State $B^M \gets B^M \lor \mathcal{I}$
@@
         \algrestore{learn}
-        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
+        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T$} $\equiv \bot$
         \Require \emph{May-invariant} holds
         \Ensure \emph{May-invariant} holds
-        \Ensure $s(X_T) \land B^m[$\Call{height}{$T$}$] \equiv \bot$
-        \Comment{$s$ will be removed from $B^m[$\Call{height}{$T$}$]$}
+        \Ensure $s(X_T) \land B^m[$\Call{height}{$k, T$}$] \equiv \bot$
+        \Comment{$s$ will be removed from $B^m[$\Call{height}{$k, T$}$]$}
         \Function{\textoverline{learn}}{$s, T$}
@@
             \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
-            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
-            \For{$i = 1$ to \Call{height}{$n$}}
+            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T_1$}, \Call{\textoverline{treeFormula}}{$k, T_2$}}
+            \For{$i = 1$ to \Call{height}{$k, n$}}
                 \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
@@
 children.  There is no candidate strategy for $T$ so $s \land
-\textsc{treeFormula}(T)$ is unsatisfiable.  By construction,
-$\textsc{treeFormula}(T) \equiv \textsc{treeFormula}(T_1) \land
-\textsc{treeFormula}(T_2)$ and hence $s \land \textsc{treeFormula}(T_1) \land
-\textsc{treeFormula}(T_2)$ is also unsatisfiable. This enables the construction
+\textsc{treeFormula}(k, T)$ is unsatisfiable.  By construction,
+$\textsc{treeFormula}(k, T) \equiv \textsc{treeFormula}(k, T_1) \land
+\textsc{treeFormula}(k, T_2)$ and hence $s \land \textsc{treeFormula}(k, T_1) \land
+\textsc{treeFormula}(k, T_2)$ is also unsatisfiable. This enables the construction
 of an \emph{interpolant} that captures losing states at node $n$.
@@
 
-We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(T_1)$
-and $F_2 = \textsc{treeFormula}(T_2)$ (line~5). The only variables shared between $F_1$
+We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(k, T_1)$
+and $F_2 = \textsc{treeFormula}(k, T_2)$ (line~5). The only variables shared between $F_1$
 and $F_2$ are the state variable copies belonging to node $n$. By the properties of the interpolant,
@@
     \begin{algorithmic}[1]
-        \Function{treeFormula}{$T$}
-        \If{$\Call{height}{T} = 0$}
+        \Function{treeFormula}{$k, T$}
+        \If{$\Call{height}{k, T} = 0$}
         \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
@@
         \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
-            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
+            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{k, n})$$
         }
@@
         \algrestore{tf1}
-        \Function{\textoverline{treeFormula}}{$T$}
-        \If{$\Call{height}{T} = 0$}
+        \Function{\textoverline{treeFormula}}{$k, T$}
+        \If{$\Call{height}{k, T} = 0$}
         \State \Return{\Call{E}{$X_T$}}
         \Else
-        \State \Return{ \Call{$B^m[\Call{height}{T}]$}{$X_{T}$} $\land$ \\
-            $$\bigg( \Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})\bigg)$$ }
+        \State \Return{ \Call{$B^m[\Call{height}{k, T}]$}{$X_{T}$} $\land$ \\
+            $$\bigg( \Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{k, n})\bigg)$$ }
         \EndIf
@@
 are initialised to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
-is discovered for a node $n$ with bound $\textsc{height}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
+is discovered for a node $n$ with bound $\textsc{height}(k, n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
 $B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
@@
 1.  Consider formulas $F_1=s(X_T) \land
-\textsc{\textoverline{treeFormula}}(T_1)$ and $F_2 =
-\textsc{\textoverline{treeFormula}}(T_2)$.  These formulas only share variables
+\textsc{\textoverline{treeFormula}}(k, T_1)$ and $F_2 =
+\textsc{\textoverline{treeFormula}}(k, T_2)$.  These formulas only share variables
 $X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
 any solution of $F_1 \land F_2$ also satisfies $s(X_T) \land
-\textsc{\textoverline{treeFormula}}(T)$, which is unsatisfiable (precondition (b)).  Hence the
+\textsc{\textoverline{treeFormula}}(k, T)$, which is unsatisfiable (precondition (b)).  Hence the
 interpolation operation is defined for $F_1$ and $F_2$.
@@
 of depth 1, this means that the environment cannot force the game into
-$B^m[\textsc{height}(n)-1]$ playing against the counterexample moves in $T_2$.
-Hence, $\II \cap Upre(B^m[\textsc{height}(n)-1]) = \emptyset$.  Furthermore,
-since the may-invariant holds, $\II \cap Upre(B^m[\textsc{height}(i)]) =
-\emptyset$, for all $i < \textsc{height}(n)$.  Hence, removing $\II$ from all
-$B^m[i], i\leq \textsc{height}(n)$ in line~(15) preserves the may-invariant,
+$B^m[\textsc{height}(k, n)-1]$ playing against the counterexample moves in $T_2$.
+Hence, $\II \cap Upre(B^m[\textsc{height}(k, n)-1]) = \emptyset$.  Furthermore,
+since the may-invariant holds, $\II \cap Upre(B^m[i]) =
+\emptyset$, for all $i < \textsc{height}(k, n)$.  Hence, removing $\II$ from all
+$B^m[i], i\leq \textsc{height}(k, n)$ in line~(15) preserves the may-invariant,
 thus satisfying the first post-condition.
@@
 assignment to $X_n$ that satisfies $s(X_T) \land
-\textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
-removing $\II$ from $B^m[\textsc{height}(n)]$ makes $s(X_T) \land
-\textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
+\textsc{\textoverline{treeFormula}}(k, T_1)$ also satisfies $\II$.  Hence,
+removing $\II$ from $B^m[\textsc{height}(k, n)]$ makes $s(X_T) \land
+\textsc{\textoverline{treeFormula}}(k, T_1)$ unsatisfiable, and hence all
 preconditions of the recursive invocation of \textsc{\textoverline{learn}} in
@@
 At the second last recursive call to \textsc{\textoverline{learn}}, tree $T_1$
-is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
-\equiv B^m[\textsc{height}(T_1)](X^T)$; hence $s(X_T) \land
-\textsc{\textoverline{treeFormula}}(T_1) \equiv s(X_T) \land
-B^m[\textsc{height}(T_1)](X^T) \equiv \bot$.  Thus the second postcondition of
+is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(k, T_1)
+\equiv B^m[\textsc{height}(k, T_1)](X^T)$; hence $s(X_T) \land
+\textsc{\textoverline{treeFormula}}(k, T_1) \equiv s(X_T) \land
+B^m[\textsc{height}(k, T_1)](X^T) \equiv \bot$.  Thus the second postcondition of
 \textsc{\textoverline{learn}} holds.
@@
 The proof of \textsc{learn} is similar to the above proof of \textsc{learn}. An
-interpolant constructed from $F_1=s(X_T) \land \textsc{treeFormula}(T_1)$
-and $F_2 = \textsc{treeFormula}(T_2)$ has the property $\II \land F_2 \equiv
+interpolant constructed from $F_1=s(X_T) \land \textsc{treeFormula}(k, T_1)$
+and $F_2 = \textsc{treeFormula}(k, T_2)$ has the property $\II \land F_2 \equiv
 \bot$ and the precondition ensures that the controller is unable to force the
@@
 Likewise, in the second last recursive call of \textsc{learn} with the empty
-tree $T_1$ and root node $n$: $\textsc{treeFormula}(T_1) \equiv \lnot
-B^M(X_T)$.  Hence $s(X_T) \land \textsc{treeFormula}(T_1) \equiv
+tree $T_1$ and root node $n$: $\textsc{treeFormula}(k, T_1) \equiv \lnot
+B^M(X_T)$.  Hence $s(X_T) \land \textsc{treeFormula}(k, T_1) \equiv
 s(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $s(X_T) \land

### 2016-03-25 paper.tex
Better overview of bounded synth by removing discussion of quantifiers

@@
 
-Realisability for a bounded safety game can be solved by checking a quantified
-formula consisting of the game's transition relation \emph{unrolled} to the
-length of the bounded game. The quantifiers represent the two players of the
-game: universal for the controller and existential for the environment.  Just
-as the players alternate in choosing actions every round of the game, the
-formula is prefixed by a quantifier alternation over $U$ and $C$ for each
-round.
-
-Instead of solving both quantifiers at once, two competing solvers are used to
-solve abstractions of the game. Each solver builds a candidate strategy for its
-corresponding player and uses its opponent's candidate to restrict its own
-search. The solvers communicate to refine one another's candidate strategies by
-finding counterexamples. Eventually one solver will be unable to find a
-counterexample, indicating that its opponent's candidate strategy is a winning
-strategy for the game.
+The algorithm consists of two competing solvers, one for the system and one for
+the environment, that solve abstractions of the safety game. Each solver builds
+a candidate strategy for its corresponding player consisting of assignments to
+$2^C$ and $2^U$ respectively.
+
+The game abstraction is initially constructed from an opponent candidate
+strategy. It is refined by counterexamples to the solver's own candidate
+strategy that are formed when the opponent solves the abstract game
+corresponding to that strategy. By abstracting the game in this way we
+conceptually direct the solvers' search to consider opponent actions that may
+lead to winning strategies.
 
@@
 Specifically, we consider abstractions represented as trees of actions,
-referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows an
-example abstract game tree for the controller (abstract game trees for the
+referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
+an example abstract game tree for the controller (abstract game trees for the
 environment are similar).  In the abstract game, the controller is required to
@@
         \Function{learn}{$s, T$}
-            \IIf{$T = \emptyset$}
+            \IIf{\Call{succ}{$T$} $= \emptyset$}
                 \Return
@@
         \Ensure $s(X_T) \land B^m[$\Call{height}{$k, T$}$] \equiv \bot$
-        \Comment{$s$ will be removed from $B^m[$\Call{height}{$k, T$}$]$}
+        \Comment{$s$ will be removed from $B^m$}
         \Function{\textoverline{learn}}{$s, T$}
-            \IIf{$T = \emptyset$}
+            \IIf{\Call{succ}{$T$} $= \emptyset$}
                 \Return
@@
             \State $B^M \gets E$
-            \State $B^m[0] \gets \top$
+            \State $B^m[0] \gets E$
             \For{$k = 1 \dots$}
@@
                 \EndIIf
-                \State $B^m[k] \gets \top$
+                \State $B^m[k] \gets E$
                 \State \Call{checkBound}{$k$}

### 2016-03-25 paper.tex
Hopefully cleared up some confusing wording in sect 2

@@
 
-The algorithm consists of two competing solvers, one for the system and one for
+The algorithm consists of two competing solvers, one for the controller and one for
 the environment, that solve abstractions of the safety game. Each solver builds
@@
 referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
-an example abstract game tree for the controller (abstract game trees for the
-environment are similar).  In the abstract game, the controller is required to
-pick actions from the tree, starting from the root node.  After reaching a
-leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
-restricts the initial controller action to the set $\{a, d\}$.  After choosing
-action $a$, the controller is required to play action $b$ in the next round.
-It then reaches a leaf of the tree and continues playing unrestricted.  The
-root of the tree is annotated by the initial state $s$ of the abstract game and
-the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
-of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
-edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
-$n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a game
-bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root node
-of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs
-$\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the edge
-connecting $n$ and $n'$.
+an example abstract game tree restricting the controller (abstract game trees
+restricting the environment are similar).  In the abstract game, the controller
+is required to pick actions from the tree, starting from the root node.  After
+reaching a leaf, it continues playing unrestricted.  The tree in
+Figure~\ref{fig:agt} restricts the initial controller action to the set $\{a,
+d\}$.  After choosing action $a$, the controller is required to play action $b$
+in the next round.  It then reaches a leaf of the tree and continues playing
+unrestricted.  The root of the tree is annotated by the initial state $s$ of
+the abstract game and the bound $k$ on the number of rounds.  We denote
+$\textsc{nodes}(T)$ the set of all nodes of a tree $T$, $\textsc{leaves}(T)$
+the subset of leaf nodes.  For edge $e$, $\textsc{action}(e)$ is the action
+that labels the edge, and for node $n$, $\textsc{height}(k, n)$ is the distance
+from n to the last round of a game bounded to $k$ rounds.  $\textsc{height}(k,
+T)$ is the height of the root node of the tree.  For node $n$ of the tree,
+$\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$ where $n'$ is a
+child node of $n$ and $e$ is the edge connecting $n$ and $n'$.
 
@@
 state $s$, bound $k$ and an abstract game tree $T$ and returns a winning
-partial strategy for $\textsc{opponent}(p)$, if one exists.  The initial
-invocation of the algorithm takes the initial state $I$, bound $\kappa$ and an
-empty abstract game tree $\emptyset$.  Initially the solver is playing on
-behalf of the environment since that player takes the first move in every game
-round.  The empty game tree does not constrain opponent moves, hence solving
-such an abstraction is equivalent to solving the original concrete game.
+partial strategy for $p$, if one exists.  The initial invocation of the
+algorithm takes the initial state $I$, bound $\kappa$ and an empty abstract
+game tree $\emptyset$.  Initially the solver is playing on behalf of the
+environment since that player takes the first move in every game round.  The
+empty game tree does not constrain opponent moves, hence solving such an
+abstraction is equivalent to solving the original concrete game.
 
@@
 (CEGAR) loop.  The first step of the algorithm uses the \textsc{findCandidate}
-function, described below, to come up with a candidate partial strategy for
-$T$. If it fails to find a strategy, this means that no winning partial
-strategy exists for $T$.  If, on the other hand, a candidate partial strategy
-is found, we need to verify if it is indeed winning for $T$.
+function, described below, to come up with a candidate partial strategy that is
+winning when the opponent is restricted to $T$.  If it fails to find a
+strategy, this means that no winning strategy exists against the opponent
+playing its partial strategy represented by $T$.  If, on the other hand, a
+candidate partial strategy is found, we need to verify if it is indeed winning
+for the abstract game $T$.
 
@@
 strategy for an abstract game tree $T$ with initial state $s$ and bound $k$.
-For the system, this means the environment partial strategy represented by $T$
+For the controller, this means the environment partial strategy represented by $T$
 will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
-bound is irrelevant).  For the environment, the system can avoid $E$ for $k$
+bound is irrelevant).  For the environment, the controller can avoid $E$ for $k$
 rounds no matter the assignments to $U$ variables. Hence, we have established
-that $s$ is a losing state either always (for the system) or for bound $k$ (for
+that $s$ is a losing state either always (for the controller) or for bound $k$ (for
 the environment. We can learn additional losing states from the tree via

### 2016-03-25 paper.tex
Added some high level intuition to the conclusion

@@
 \cite{mcmillan2002}, or by interpolation \cite{mcmillan2003}. However, there
-are no corresponding adaptations to synthesis. Incremental induction
-\cite{bradley2011} is another technique for unbounded model checking with an
-equivalent synthesis method \cite{morgenstern2013}, which computes the set of
-safe states incrementally.
+are no corresponding adaptations to synthesis.
+
+Incremental induction \cite{bradley2011} is another technique for unbounded
+model checking with an equivalent synthesis method \cite{morgenstern2013},
+which computes sets of states that overapproximate the losing states (similar
+to our $B^m$) and another set of winning states (similar to the negation of
+$B^M$). Their algorithm maintains a similar invariant over the sets of losing
+states as our approach and has the same termination condition. It differs in
+how the sets are computed, which it does by inductively proving the number of
+game rounds required by the environment to win from the initial state.
 
@@
 other synthesis techniques by combining counterexample guided game solving with
-winning set computation. The results show that our approach is able to solve
-more unique instances than other solvers by performing well on certain classes
-of games that are hard for other methodologies.
+winning set computation. Intuitively, the abstraction refinement framework of
+the bounded synthesis algorithm restricts the search to consider only moves
+that may lead to winning strategies. By constructing sets of bad states during
+this search we aim to consider only states that are relevant to a winning
+strategy while solving the unbounded game.
+
+The results show that our approach is able to solve more unique instances than
+other solvers by performing well on certain classes of games that are hard for
+other methodologies.
 

### 2016-03-25 leonid_rebuttal.txt
Leonid's version of rebuttal

We thank the reviewers for detailed feedback!

We will take your comments into account to correct and improve the 
presentation.  We will also address concerns raised in the reviews 
as outlined below.

Reviewer A
~~~~~~~~~~

Relation to algorithms that compute the winning region:

Our technique can indeed be seen as a hybrid between 
counterexample-guided synthesis and traditional winning region 
computation.  By letting the former guide the exploration, we are 
able to solve the game while learning fewer states.  This is the 
aim of the abstraction refinement framework, to restrict the 
dueling solvers to consider only opponent moves that may lead
to spoiling strategies.  As the reviewer points out, this allows 
computing deeper stategies where other methods hit the state 
explosion barrier.

Deriving losing states from a specific Abstract Game Tree (AGT):

Assume that player $P$ loses from state $s$ when the opponent $O$ 
plays according to AGT $t$.  This means that there exists a 
winning opponent stategy from $s$, hence the state is losing for 
$P$ by definition and can be added to the set of losing states for 
the entire game.

Note that the opposite is not true: assuming that $P$ can win from 
$s$ when $O$ plays according to $t$, $s$ is not necessarily a 
winning state for $P$, as there may exist a different AGT $t'$ 
winning for the opponent.  This is why we can learn losing, but 
not winning, states from an AGT.

Reviewer B
~~~~~~~~~~

Thank you for your in-depth review!

(06) Experiments

We observe that Incremental Induction and Learning-Based backends 
of Demiurge individually can not solve the 12 instances; however 
they can solve them when combined in parallel with the 
template-based backend.  We conjecture that it is the 
template-based algorithm that solves these instances within the 
parallel portfolio.  This algorithm is very different from our 
method as well as other Demiurge algorithms.  We will validate 
this conjecture and more thoroughly evaluate our solver against 
the template-based method.

We will also add a theoretical and experimental comparison with 
the inductive synthesis method of [16].  Note that, as shown in 
[4], the algorithm of [16] is less efficient than Demiurge, which 
we compare against.

(07), (42) Strategy computation.

The method described in (42) is correct.  In addition, [9] 
describes a strategy extraction algorithm that operates on 
abstract game trees and does not require CPre computation.  This 
algorithm can be easily adapted to work with our method.

(08) Open sourcing

We are preparing an open source version of the tool for SYNTCOMP 
2016.

Reviewer C
~~~~~~~~~~

We will characterize the unique instances solved by our method.  
Intuitively, in these examples the winning set cannot be compactly 
represented using BDD or CNF and there exists a winning strategy, 
where at every round the player picks one of a small number of 
winning moves (i.e., there exists a compact winning abstract game 
tree for the player).

Regarding the number of unique solutions: while not a radical 
improvement, we push the virtual best in the sequential 
realizability track from 203 to 215 out of 250 instances.  As 
fewer unsolved instances remain, even such moderate improvements 
are hard to achieve.

It is also interesting that we are able to solve at least 12 
instances that are not solved by the winning solver in the 
competition, which is more than any other solver, thus emphasizing 
that our method has complementary strengthes to other techniques.

Reviewer D
~~~~~~~~~~

Indeed, BDD-based methods are still more scalable and versatile 
than SAT-based synthesis, including our method.  In this context, 
our contribution is two-fold.  First, we propose a SAT-based 
method that works on some classes of problems where BDDs blow up.  
Hence, it may be productive to combine this method with BDD or 
other SAT methods in a portfolio.  Second, by developing a new 
SAT-based synthesis technique, we contribute to the ongoing push 
to fully exploit the power of the SAT technology in synthesis, 
inspired by the success of SAT-based verification.

### 2016-04-22 paper.tex
Some of the example diagrams

@@
 
-\author{}
-\institute{}
-
-%\author{Alexander Legg\inst{1} 
-%    \and Nina Narodytska\inst{2}
-%    \and Leonid Ryzhyk\inst{2}}
-%
-%\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
-%    Communications and the Digital Economy and the Australian Research Council through the ICT
-%    Centre of Excellence program.} and UNSW \\
-%    \email{alexander.legg@nicta.com.au}
-%    \and Samsung Research America}
+\author{Alexander Legg\inst{1} 
+    \and Nina Narodytska\inst{2}
+    \and Leonid Ryzhyk\inst{2}}
+
+\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
+    Communications and the Digital Economy and the Australian Research Council through the ICT
+    Centre of Excellence program.} and UNSW \\
+    \email{alexander.legg@nicta.com.au}
+    \and Samsung Research America}
 
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    edge from parent [fixed] node [right] {$d$}
-                }
+                        edge from parent [fixed] node [left] {\texttt{req=0}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{req=0}}
+                }
+%%%                child {node [circle,draw] {}
+%%%                    edge from parent [fixed] node [right] {\texttt{gr=1}}
+%%%                }
                 node [above=4pt] {$\langle s, k \rangle$};
@@
                     child {node [circle,draw] {}
-                        node [left=4pt] {$\epsilon$}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    node [left=4pt] {$\beta$}
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    node [left=4pt] {$\gamma$}
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [right=4pt] {$\alpha$}
+                        node [left=4pt] {\texttt{gr0=0, gr1=0}}
+                        edge from parent [fixed] node [left] {\texttt{req=0}}
+                    }
+                    node [left=4pt] {\texttt{gr0=0, gr1=0}}
+                    edge from parent [fixed] node [left] {\texttt{req=0}}
+                }
+%%%                child {node [circle,draw] {}
+%%%                    node [left=4pt] {\texttt{req=1}}
+%%%                    edge from parent [fixed] node [right] {\texttt{gr=1}}
+%%%                }
+                node [left=4pt] {}
                 node [above=4pt] {$\langle s, k \rangle$};
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [right] {$e$}
-                    }
-                    edge from parent [fixed] node [right] {$d$}
+                        edge from parent [fixed] node [left] {\texttt{req=0}}
+                    }
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{req=1}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{req=0}}
                 }

### 2016-04-22 paper.tex
Added explanation of the example

@@
 algorithm.
+
+First, we introduce a running example to assist the explanation. We consider a
+simple arbiter system in which the environment requests for either 1 or 0
+resources, and the controller may grant access to two resources. The controller
+must ensure that the number of unhandled requests is not $\geq 2$.
+Figure~\ref{fig:example} shows the variables (\ref{fig:examplevars}) and the
+formulas for computing next-state variables (\ref{fig:exampletrans}) for this
+example.
+
+\begin{figure}
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \begin{tabular}{l | l | l}
+            \textbf{Controllable} & \textbf{Uncontrollable} & \textbf{State} \\
+            \hline
+            \texttt{request = \{0, 1\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
+            & \texttt{grant1 = \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
+            & & \texttt{nrequests = \{0, 1, 2, 3\}} \\
+        \end{tabular}
+        \caption{Variables}
+        \label{fig:examplevars}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \begin {align*}
+            \texttt{resource0'} & \texttt{ = grant0;} \\
+            \texttt{resource1'} & \texttt{ = grant1;} \\
+            \texttt{nrequests'} & \texttt{ = (nrequests + request >= resource0 + resource1)} \\ 
+                                & \texttt{ ? (nrequests + request - resource0 - resource1) : 0;}
+        \end{align*}
+        \caption{Transition Relation}
+        \label{fig:exampletrans}
+    \end{subfigure}
+    \caption{Example}
+    \label{fig:example}
+\end{figure}
 

### 2016-04-22 paper.tex
Refer to the example in section 2

@@
 First, we introduce a running example to assist the explanation. We consider a
-simple arbiter system in which the environment requests for either 1 or 0
+simple arbiter system in which the environment requests for either 1 or 2
 resources, and the controller may grant access to two resources. The controller
-must ensure that the number of unhandled requests is not $\geq 2$.
+must ensure that the number of unhandled requests is not $\geq 3$.
 Figure~\ref{fig:example} shows the variables (\ref{fig:examplevars}) and the
@@
             \hline
-            \texttt{request = \{0, 1\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
+            \texttt{request = \{1, 2\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
             & \texttt{grant1 = \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
@@
 
-The algorithm consists of two competing solvers, one for the controller and one for
-the environment, that solve abstractions of the safety game. Each solver builds
-a candidate strategy for its corresponding player consisting of assignments to
-$2^C$ and $2^U$ respectively.
+The bounded synthesis algorithm consists of two competing solvers, one for the
+controller and one for the environment, that solve abstractions of the safety
+game. Each solver builds a candidate strategy for its corresponding player
+consisting of assignments to $2^C$ and $2^U$ respectively.
 
@@
 referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows
-an example abstract game tree restricting the controller (abstract game trees
-restricting the environment are similar).  In the abstract game, the controller
-is required to pick actions from the tree, starting from the root node.  After
-reaching a leaf, it continues playing unrestricted.  The tree in
-Figure~\ref{fig:agt} restricts the initial controller action to the set $\{a,
-d\}$.  After choosing action $a$, the controller is required to play action $b$
-in the next round.  It then reaches a leaf of the tree and continues playing
-unrestricted.  The root of the tree is annotated by the initial state $s$ of
-the abstract game and the bound $k$ on the number of rounds.  We denote
-$\textsc{nodes}(T)$ the set of all nodes of a tree $T$, $\textsc{leaves}(T)$
-the subset of leaf nodes.  For edge $e$, $\textsc{action}(e)$ is the action
-that labels the edge, and for node $n$, $\textsc{height}(k, n)$ is the distance
-from n to the last round of a game bounded to $k$ rounds.  $\textsc{height}(k,
-T)$ is the height of the root node of the tree.  For node $n$ of the tree,
-$\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$ where $n'$ is a
-child node of $n$ and $e$ is the edge connecting $n$ and $n'$.
+an example abstract game tree restricting the environment (abstract game trees
+restricting the controller are similar).  In the abstract game, the controller
+can freely choose actions whilst the environment is required to pick actions
+from the tree.  After reaching a leaf, the enivironment continues playing
+unrestricted.  The tree in Figure~\ref{fig:agt} restricts the first and second
+environment actions to \texttt{request=1}. It then reaches a leaf of the tree
+and continues playing unrestricted.
+
+The root of the tree is annotated by the initial state $s$ of the abstract game
+and the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the
+set of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.
+For edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for
+node $n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a
+game bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root
+node of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of
+pairs $\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the
+edge connecting $n$ and $n'$.
 
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {\texttt{req=0}}
-                    }
-                    edge from parent [fixed] node [left] {\texttt{req=0}}
+                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
@@
                     child {node [circle,draw] {}
-                        node [left=4pt] {\texttt{gr0=0, gr1=0}}
-                        edge from parent [fixed] node [left] {\texttt{req=0}}
-                    }
-                    node [left=4pt] {\texttt{gr0=0, gr1=0}}
-                    edge from parent [fixed] node [left] {\texttt{req=0}}
+                        node [left=4pt] {\texttt{gr0=1, gr1=0}}
+                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                    }
+                    node [left=4pt] {\texttt{gr0=0, gr1=1}}
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {\texttt{req=0}}
-                    }
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [right] {\texttt{req=1}}
-                    }
-                    edge from parent [fixed] node [left] {\texttt{req=0}}
+                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                    }
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
@@
 
-Given a controller (environment) abstract game tree $T$ a \emph{partial
-strategy} $Strat: \textsc{nodes}(T) \rightarrow U$ ($Strat: \textsc{nodes}(T)
-\rightarrow C$) labels each node of the tree with the opponent's action to be
-played in that node.   Figure~\ref{fig:strategy} shows an example partial
-strategy.  The environment starts by choosing action $\alpha$.  If the
-controller plays $a$, the environment responds with $\beta$ in the next round,
-and so on. Given a partial strategy $Strat$, we can map each leaf $l$ of the
-abstract game tree to $\langle s',i'\rangle=\textsc{outcome}(\langle s,
-i\rangle, Strat, l)$ obtained by playing all controllable and uncontrollable
-actions on the path from the root to the leaf.
+Given an environment (controller) abstract game tree $T$ a \emph{partial
+strategy} $Strat: \textsc{nodes}(T) \rightarrow C$ ($Strat: \textsc{nodes}(T)
+\rightarrow U$) labels each node of the tree with the controller's
+(environment's) action to be played in that node.   Given a partial strategy
+$Strat$, we can map each leaf $l$ of the abstract game tree to $\langle
+s',i'\rangle=\textsc{outcome}(\langle s, i\rangle, Strat, l)$ obtained by
+playing all controllable and uncontrollable actions on the path from the root
+to the leaf.  Figure~\ref{fig:strategy} shows an example partial strategy for
+the controller.  The controller responds to the requests by first granting
+access to \texttt{resource1}, then to \texttt{resource0} in the second round.
 
@@
 the move used by the opponent to defeat a leaf of the partial strategy, which
-is appended to the corresponding node in $T$ in order to refine it in line~(9)
-as shown in Figure~(\ref{fig:refined}).
+is appended to the corresponding node in $T$ in order to refine it in line~(9).
+In our example, the controller's partial strategy in Figure~\ref{fig:strategy}
+is defeated by the counterexample action \texttt{req=2}.
+Figure~\ref{fig:refined} shows a refined AGT that includes this action.
 

### 2016-04-22 paper.tex
Change interpolation diagrams to running example

@@
         \centering
+
         \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    node [left=5pt] {$n$}
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
                 child {node [circle,draw] (n) {}
-                    child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {$e$}
-                    }
-                    child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {$f$}
-                    }
-                    edge from parent [fixed] {node [right] {$d$}}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-            \node [right=1pt of n] {$n$};
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
+                    node [left=5pt] {$n$}
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
                 child {node [circle,draw] (n) {}
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-            \node [right=1pt of n] {$n$};
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
@@
                 child {node [circle,draw,right=5pt] {}
-                    edge from parent [fixed] node [left] {$e$}
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
                 child {node [circle,draw,left=5pt] {}
-                    edge from parent [fixed] node [right] {$f$}
-                }
-                node [right=5pt] {$n$};
+                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                }
+                node [left=5pt] {$n$};
         \end{tikzpicture}

### 2016-04-24 paper.tex
Changed interpolate example to controller moves

--- 2016-04-22 paper.tex
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
+                    child {node [circle,draw,right=5pt] {}
+                        edge from parent [fixed] node [left] {\texttt{gr0=1, gr1=0}}
+                    }
+                    child {node [circle,draw,left=5pt] {}
+                        edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                    }
+                    node [left=5pt] {$n$}
+                    edge from parent [fixed] node [left] {\texttt{gr0=0, gr1=1}}
                 }
                 child {node [circle,draw] (n) {}
-                    child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {$e$}
-                    }
-                    child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {$f$}
-                    }
-                    edge from parent [fixed] {node [right] {$d$}}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-            \node [right=1pt of n] {$n$};
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
+                    node [left=5pt] {$n$}
+                    edge from parent [fixed] node [left] {\texttt{gr0=1, gr1=0}}
                 }
                 child {node [circle,draw] (n) {}
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [above=4pt] {$\langle s, 2 \rangle$};
-            \node [right=1pt of n] {$n$};
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                }
+                node [above=4pt] {$\langle s, k \rangle$};
         \end{tikzpicture}
@@
                 child {node [circle,draw,right=5pt] {}
-                    edge from parent [fixed] node [left] {$e$}
+                    edge from parent [fixed] node [left] {\texttt{gr0=1, gr1=0}}
                 }
                 child {node [circle,draw,left=5pt] {}
-                    edge from parent [fixed] node [right] {$f$}
-                }
-                node [right=5pt] {$n$};
+                    edge from parent [fixed] node [right] {\texttt{gr0=1, gr1=1}}
+                }
+                node [left=5pt] {$n$};
         \end{tikzpicture}

### 2016-04-24 paper.tex
Started on text for section 3 example - needs more work

@@
 Consider the game tree $T$ in
-Figure~\ref{fig:interpolatetree}, whose edges are labelled with environment
-actions.  Assume that there does not exist a partial strategy for the controller
-winning against this tree and that the \textsc{findCandidate} function could
-not find a candidate strategy. We choose a non-leaf node $n$ of $T$ with
-maximal depth, i.e., a node whose children are leafs
-(Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that
-both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
+Figure~\ref{fig:interpolatetree}, whose edges are labelled with controller
+actions.  There does not exist a partial strategy for the environment that is
+winning against the controller actions in this tree since any requests will
+always be granted. Thus the \textsc{findCandidate} function can not find a
+candidate strategy. We now learn the states for which these controller actions
+can force the game to stay in the controller winning set.
+
+We choose a non-leaf node $n$ of $T$ with maximal depth, i.e., a node whose
+children are leafs (Algorithm~\ref{alg:learn}, line~3). We then split the tree
+at $n$ such that both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
 Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$ except $n$'s
@@
 $\textsc{treeFormula}(k, T) \equiv \textsc{treeFormula}(k, T_1) \land
-\textsc{treeFormula}(k, T_2)$ and hence $s \land \textsc{treeFormula}(k, T_1) \land
-\textsc{treeFormula}(k, T_2)$ is also unsatisfiable. This enables the construction
-of an \emph{interpolant} that captures losing states at node $n$.
+\textsc{treeFormula}(k, T_2)$ and hence $s \land \textsc{treeFormula}(k, T_1)
+\land \textsc{treeFormula}(k, T_2)$ is also unsatisfiable. This enables the
+construction of an \emph{interpolant} that captures losing states at node $n$.
 

### 2016-04-26 paper.tex
Edits to text of section 3 regarding example

--- 2016-04-22 paper.tex
@@
                     child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                        edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
                     }
                     child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                        edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
                     }
                     node [left=5pt] {$n$}
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
+                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=0, gr1=1}}
                 }
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [right] {\texttt{req=2}}
-                    }
-                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                        edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
+                    }
+                    edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
                 }
@@
                     node [left=5pt] {$n$}
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
+                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
                 }
@@
                 child {node [circle,draw,right=5pt] {}
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
+                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
                 }
                 child {node [circle,draw,left=5pt] {}
-                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                    edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
                 }
@@
     \caption{Splitting of an abstract game tree by the learning procedure.}
+    \label{fig:interpolanttrees}
 \end{figure}
@@
 
-Consider the game tree $T$ in
-Figure~\ref{fig:interpolatetree}, whose edges are labelled with environment
-actions.  Assume that there does not exist a partial strategy for the controller
-winning against this tree and that the \textsc{findCandidate} function could
-not find a candidate strategy. We choose a non-leaf node $n$ of $T$ with
-maximal depth, i.e., a node whose children are leafs
-(Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that
-both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
+Consider the game tree $T$ in Figure~\ref{fig:interpolatetree} from our running
+example.  This tree is labelled with controller actions that prevent the
+environment from forcing the game into the error state. No matter how many
+requests the environment makes, the controller will grant a surplus. So the
+\textsc{findCandidate} function can not find a candidate strategy for the
+environment.  We now learn the states for which these controller actions can
+force the game to stay in the controller winning set.
+
+We choose a non-leaf node $n$ of $T$ with maximal depth, i.e., a node whose
+children are leafs (Algorithm~\ref{alg:learn}, line~3). We then split the tree
+at $n$ such that both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
 Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$ except $n$'s
@@
 children.  There is no candidate strategy for $T$ so $s \land
-\textsc{treeFormula}(k, T)$ is unsatisfiable.  By construction,
-$\textsc{treeFormula}(k, T) \equiv \textsc{treeFormula}(k, T_1) \land
-\textsc{treeFormula}(k, T_2)$ and hence $s \land \textsc{treeFormula}(k, T_1) \land
-\textsc{treeFormula}(k, T_2)$ is also unsatisfiable. This enables the construction
-of an \emph{interpolant} that captures losing states at node $n$.
+\textsc{\textoverline{treeFormula}}(k, T)$ is unsatisfiable.  By construction,
+$\textsc{\textoverline{treeFormula}}(k, T) \equiv
+\textsc{\textoverline{treeFormula}}(k, T_1) \land
+\textsc{\textoverline{treeFormula}}(k, T_2)$ and hence $s \land
+\textsc{\textoverline{treeFormula}}(k, T_1) \land
+\textsc{\textoverline{treeFormula}}(k, T_2)$ is also unsatisfiable. This enables
+the construction of an \emph{interpolant} that captures losing states at node
+$n$.
 
@@
 \mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
-following $T_1$ and avoiding error states.
-
-We have discovered a set $\mathcal{I}$ of states losing for the controller. We
-record these newly discovered states in a set $B^M$ of ``bad'' states that
-\emph{must} be losing for the controller (Algorithm~\ref{alg:learn}, line~6).
-This set is initialised to the error set $E$ and grows as new losing states are
-discovered.  The \textsc{treeFormula} function is modified for the unbounded
-solver (Algorithm~\ref{alg:unboundedTreeFormula}) to replace $E$ with $B^M$.
-This enables further interpolants to be constructed by the learning procedure
-recursively splitting more nodes from $T_1$ (Algorithm~\ref{alg:learn}, line~7)
-since the states that are losing to $T_2$ are now blocked by $B^M$.
-
-\begin{algorithm}[t]
-    \caption{Amended tree formulas for Controller and Environment}
-    \label{alg:unboundedTreeFormula}
-    \begin{algorithmic}[1]
-        \Function{treeFormula}{$k, T$}
-        \If{$\Call{height}{k, T} = 0$}
-        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
-        \Else
-        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
-            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{k, n})$$
-        }
-        \EndIf
-        \EndFunction
-        \algstore{tf1}
-    \end{algorithmic}
+following $T_1$ and avoiding error states. For example, in the pictured trees
+the set $nrequests = 0$ is reachable at $n$ and no move the environment chooses
+will reach the error state in one round from this set.
+
+We have discovered a set $\mathcal{I}$ of states losing for the environment.
+Environment-losing states are only losing for a particular bound: given that
+there does not exist an environment strategy that forces the game into an error
+state in $k$ rounds or less; there may still exist a longer environment-winning
+strategy.  We therefore record learned environment-losing states along with
+associated bounds.  To this end, we maintain a conceptually infinite array of
+sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
+$B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an environment-losing
+set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
+n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from
+$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
+
+The \textsc{\textoverline{treeFormula}} function is modified for the unbounded
+solver (Algorithm~\ref{alg:unboundedTreeFormula}) to constrain the environment
+to the appropriate $B^m$. This enables further interpolants to be constructed
+by the learning procedure recursively splitting more nodes from $T_1$
+(Algorithm~\ref{alg:learn}, line~7) since the states that are losing to $T_2$
+are no longer contained in $B^m$.
+
+\begin{algorithm}[t] \caption{Amended tree formulas for Controller and
+    Environment} \label{alg:unboundedTreeFormula} \begin{algorithmic}[1]
+        \Function{treeFormula}{$k, T$} \If{$\Call{height}{k, T} = 0$} \State
+        \Return{ $\lnot \Call{$B^M$}{X_{T}}$ } \Else \State \Return{$\lnot
+            \Call{$B^M$}{X_{T}} \land$ \\ $$\bigwedge_{\langle e, n \rangle \in
+            \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e =
+            \Call{action}{e} \land \Call{treeFormula}{k, n})$$ } \EndIf
+        \EndFunction \algstore{tf1} \end{algorithmic}
 
@@
 
-Learning of states losing for the environment is similar (\textsc{\textoverline{learn}} in Algorithm~\ref{alg:learn}). 
-The main difference is that environment-losing states are only losing for a particular
-bound: given that there does not exist an environment strategy that
-forces the game into an error state in $k$ rounds or less; there may still exist a longer
-environment-winning strategy.  We therefore record learned environment-losing states
-along with associated bounds.  To this end, we maintain a conceptually infinite array of
-sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.  $B^m[k]$ 
-are initialised to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
-is discovered for a node $n$ with bound $\textsc{height}(k, n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
-$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
-
-%It is more complicated to learn states from environment-losing abstract game
-%trees. 
-%States are only losing for the environment for a particular bound. For a
-%state learned at a node $n$ it is losing at $\textsc{bound}(n)$. Therefore we
-%maintain an array of state sets indexed by the bound at which that set of
-%states is losing for the environment. Instead of accumulating states, as we do
-%in $B^M$, we instead remove states from the array $B^m$. This array is
-%initialised with $\top$ and contains only states that \emph{may} lose for the
-%controller. 
-The modified \textsc{\textoverline{treeFormula}} function (Algorithm~\ref{alg:unboundedTreeFormula})
-constrains the environment to may-losing states, which also allows for
-recursive learning over the entire tree.
-
-%We compile interpolants into DNF cubes by constructing a temporary BDD. This
-%smaller representation is necessary for maintaining the size of formulas passed
-%to the SAT solver. However, it can occasionally lead to state explosion.
-
-%The recursive learning algorithms, \textsc{learn} (for the controller) and
-%\textsc{\textoverline{learn}} (for the environment), are detailed in
-%Algorithm~\ref{alg:learn}. When executing \textsc{findCandidate} in the
-%unbounded solver, lines 19 and 20 become active and call into the learning
-%procedures.
+Learning of states losing from the controller is similar (\textsc{learn} in
+Algorithm~\ref{alg:learn}). The main difference is that environment-losing
+states are losing for all bounds. Therefore we record these states in a single
+set $B^M$ of must-losing states (Algorithm~\ref{alg:learn}, line~6).  This set
+is initialised to the error set $E$ and grows as new losing states are
+discovered.  The modified \textsc{\textoverline{treeFormula}} function
+(Algorithm~\ref{alg:unboundedTreeFormula}) blocks must-losing states, which
+also allows for recursive learning over the entire tree.
 

### 2016-04-27 paper.tex
Leonid's affiliation
Fixes to the example introduction

@@
 
+\let\proof\relax
+\let\endproof\relax
+\usepackage{amsthm}
+
 \algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
@@
 
+\newtheorem*{exmp}{Example}
+
 \pagestyle{headings}
@@
     \email{alexander.legg@nicta.com.au}
-    \and Samsung Research America}
+    \and Samsung Research America\thanks{Work completed at 
+    Carnegie Mellon University}}
 
@@
 
-First, we introduce a running example to assist the explanation. We consider a
+\begin{exmp}
+We introduce a running example to assist the explanation. We consider a
 simple arbiter system in which the environment requests for either 1 or 2
-resources, and the controller may grant access to two resources. The controller
-must ensure that the number of unhandled requests is not $\geq 3$.
-Figure~\ref{fig:example} shows the variables (\ref{fig:examplevars}) and the
-formulas for computing next-state variables (\ref{fig:exampletrans}) for this
-example.
+resources, and the controller may grant access to two resources. The total
+number of requests grows each round by the number of environment requests and
+shrinks by the number of resources granted by the controller in the previous
+round.  The controller must ensure that the number of unhandled requests does
+not accumulate to more than 2.  Figure~\ref{fig:example} shows the variables
+(\ref{fig:examplevars}) and the formulas for computing next-state variables
+(\ref{fig:exampletrans}) for this example. We use primed identifiers to denote
+next-state variables and curly braces to define the domain of a variable.
+
+This example is the $n=2$ instance of the more general problem of an arbiter of
+$n$ resources. The set of winning states grows exponentially with the number of
+resources the controller could use to grant requests, which makes the problem
+hard for BDD solvers. This example illustrates the strength of the approach
+presented here, which does not need to enumerate all winning states as will be
+shown in Section~3.
+
+\end{exmp}
 
@@
             \hline
-            \texttt{request = \{1, 2\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
-            & \texttt{grant1 = \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
-            & & \texttt{nrequests = \{0, 1, 2, 3\}} \\
+            \texttt{request : \{1, 2\}} & \texttt{grant0 = \{0, 1\}} & \texttt{resource0 = \{0, 1\}} \\
+            & \texttt{grant1 : \{0, 1\}} & \texttt{resource1 = \{0, 1\}} \\
+            & & \texttt{nrequests : \{0, 1, 2, 3\}} \\
         \end{tabular}
@@
         \label{fig:examplevars}
+    \end{subfigure}
+
+    \begin{subfigure}[t]{\textwidth}
+        \centering
+        \texttt{resource0 = 0; resource1 = 0; nrequests = 0;}
+        \caption{Initial State}
+        \label{fig:exampleinit}
     \end{subfigure}
@@
 is appended to the corresponding node in $T$ in order to refine it in line~(9).
-In our example, the controller's partial strategy in Figure~\ref{fig:strategy}
-is defeated by the counterexample action \texttt{req=2}.
-Figure~\ref{fig:refined} shows a refined AGT that includes this action.
+
+\begin{exmp}
+    The controller's partial strategy in
+    Figure~\ref{fig:strategy} is defeated by the counterexample action
+    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
+    this action.
+\end{exmp}
 

### 2016-04-28 paper.tex
Example dump, needs lots of editing

@@
 resources the controller could use to grant requests, which makes the problem
-hard for BDD solvers. This example illustrates the strength of the approach
-presented here, which does not need to enumerate all winning states as will be
-shown in Section~3.
+hard for BDD solvers. In Section~3 we will outline how this example can be
+solved without enumerating all winning states.
 
@@
 \tikzstyle{fixed}=[solid]
-\tikzstyle{unfixed}=[dash pattern = on 2pt off 2pt]
 \begin{figure}
@@
     \begin{subfigure}[t]{.25\textwidth}
+        \centering
+        \begin{tikzpicture}[level distance = 5mm]
+            \node [circle,draw,inner sep=1pt] (root){}
+                child {node [circle,draw,inner sep=1pt] {}
+                    child {node [circle,draw,inner sep=1pt] {}
+                        child {node [circle,draw,inner sep=1pt] {}
+                            child {node [circle,draw,inner sep=1pt] {}
+                                child {node [circle,draw,inner sep=1pt] {}
+                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                                }
+                                node [left=4pt] {\texttt{req=1}}
+                            }
+                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                        }
+                        node [left=4pt] {\texttt{req=1}}
+                    }
+                node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                }
+                node [left=4pt] {\texttt{req=1}}
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \captionsetup{width=.8\textwidth}
+        \caption{Environment-winning trace}
+        \label{fig:trace}
+    \end{subfigure}
+    \begin{subfigure}[t]{.1\textwidth}
         \centering
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
                     edge from parent [fixed] node [left] {\texttt{req=1}}
                 }
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
-                }
-%%%                child {node [circle,draw] {}
-%%%                    edge from parent [fixed] node [right] {\texttt{gr=1}}
-%%%                }
-                node [above=4pt] {$\langle s, k \rangle$};
+                node [left=4pt] {}
+                node [above=4pt] {$\langle s, 3 \rangle$};
         \end{tikzpicture}
@@
         \label{fig:agt}
-    \end{subfigure}%
+    \end{subfigure}
     \begin{subfigure}[t]{.25\textwidth}
+        \centering
+        \begin{tikzpicture}[level distance = 5mm]
+            \node [circle,draw,inner sep=1pt] (root){}
+                child {node [circle,draw,inner sep=1pt] {}
+                    child {node [circle,draw,inner sep=1pt] {}
+                        child {node [circle,draw,inner sep=1pt] {}
+                            child {node [circle,draw,inner sep=1pt] {}
+                                child {node [circle,draw,inner sep=1pt] {}
+                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                                }
+                                node [left=4pt] {\texttt{req=1}}
+                            }
+                            node [left=4pt] {\texttt{gr0=0 gr1=1}}
+                        }
+                        node [left=4pt] {\texttt{req=1}}
+                    }
+                node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                }
+                node [left=4pt] {\texttt{req=1}}
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \captionsetup{width=.8\textwidth}
+        \caption{Controller-winning trace}
+        \label{fig:trace2}
+    \end{subfigure}
+    \begin{subfigure}[t]{.1\textwidth}
         \centering
@@
                     child {node [circle,draw] {}
-                        node [left=4pt] {\texttt{gr0=1, gr1=0}}
+                        % node [left=4pt] {\texttt{gr0=1 gr1=0}}
                         edge from parent [fixed] node [left] {\texttt{req=1}}
                     }
-                    node [left=4pt] {\texttt{gr0=0, gr1=1}}
+                    % node [left=4pt] {\texttt{gr0=0 gr1=1}}
                     edge from parent [fixed] node [left] {\texttt{req=1}}
@@
                 node [left=4pt] {}
-                node [above=4pt] {$\langle s, k \rangle$};
+                node [above=4pt] {$\langle s, 3 \rangle$};
         \end{tikzpicture}
         \captionsetup{width=.8\textwidth}
-        \caption{AGT with partial strategy}
-        \label{fig:strategy}
-    \end{subfigure}%
+        \caption{AGT}
+        \label{fig:agt2}
+    \end{subfigure}
     \begin{subfigure}[t]{.25\textwidth}
@@
                     child {node [circle,draw] {}
+                        child {node [circle,draw] {}
                             edge from parent [fixed] node [left] {\texttt{req=1}}
@@
                     }
-                node [above=4pt] {$\langle s, k \rangle$};
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [right] {\texttt{req=2}}
+                        }
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{req=1}}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [right] {\texttt{req=2}}
+                        }
+                        edge from parent [fixed] node [right] {\texttt{req=2}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{req=2}}
+                }
+                node [above=4pt] {$\langle s, 2 \rangle$};
         \end{tikzpicture}
@@
     \end{subfigure}%
-    \begin{subfigure}[t]{.25\textwidth}
-        \centering
-        \begin{tikzpicture}[scale=0.75,dash pattern = on 2pt off 2pt, level distance = 10mm]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [left] {$c$}
-                        }
-                        edge from parent [fixed] node [left] {$b$}
-                    }
-                    edge from parent [fixed] node [left] {$a$}
-                }
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [left] {$f$}
-                        }
-                        edge from parent [fixed] node [right] {$e$}
-                    }
-                    edge from parent [fixed] node [right] {$d$}
-                }
-                node [above=4pt] {$\langle s, 3 \rangle$};
-        \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
-        \caption{AGT extended to bound 3 with arbitrary controller actions}
-        \label{fig:extended}
-    \end{subfigure}
     \caption{Abstract game trees.}
@@
 playing all controllable and uncontrollable actions on the path from the root
-to the leaf.  Figure~\ref{fig:strategy} shows an example partial strategy for
-the controller.  The controller responds to the requests by first granting
-access to \texttt{resource1}, then to \texttt{resource0} in the second round.
+to the leaf.
+
+%%%Figure~\ref{fig:strategy} shows an example partial strategy for
+%%%the controller.  The controller responds to the requests by first granting
+%%%access to \texttt{resource1}, then to \texttt{resource0} in the second round.
+
+\begin{exmp}
+
+    To solve our example, the algorithm begins by invoking a SAT solver to find
+    a trace of length $k$ from the initial state to the error state (see
+    Figure~\ref{fig:trace}). This trace is used to create an abstract game tree
+    by taking the first action in the trace and fixing it as the first action
+    in the game tree (Figure~\ref{fig:agt}).  This abstract game tree
+    represents the conjecture that \texttt{req=1} may be a good first action.
+    Next the solver attempts to find a counterexample to this conjecture by
+    finding a trace that avoids the error state with the first environment
+    action fixed to \texttt{req=1} (Figure~\ref{fig:trace2}), i.e.  by playing
+    the abstract game on behalf of the controller.
+
+    Observe that the controller chooses to grant only one resource, just enough
+    to ensure that an error state will not be reached. This example will show
+    the worst case when the SAT solver returns actions that will lead to large
+    abstract game trees. However, in practice we employ heuristics to guide the
+    SAT solver away from the worst case.
+
+    The environment's candidate is refined using the controller's
+    counterexample traces. In Figure~\ref{fig:agt2}, the game tree tree is
+    refined based on the first action of the controller's trace. The
+    \texttt{req=1} action in the second round is still winning under the
+    assumption that the controller will not grant any resources. The
+    environment is free to make that assumption because the controller's
+    counterexample does not yet contain any grants. Eventually, the refinements
+    to the environment's candidate will force the controller to grant more
+    resources and result in a game tree such as in Figure~\ref{fig:refined}.
+
+\end{exmp}
 
@@
 solver to search for candidate partial strategies. Candidate strategies are
-checked by a call to the opponent solver, which also refines the abstract
-game tree. The refinements are then solved recursively.  The full procedure is
+checked by a call to the opponent solver, which also refines the abstract game
+tree. The refinements are then solved recursively.  The full procedure is
 illustrated in Algorithm~\ref{alg:bounded}.
@@
 
-\begin{exmp}
-    The controller's partial strategy in
-    Figure~\ref{fig:strategy} is defeated by the counterexample action
-    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
-    this action.
-\end{exmp}
+%%%\begin{exmp}
+%%%    The controller's partial strategy in
+%%%    Figure~\ref{fig:strategy} is defeated by the counterexample action
+%%%    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
+%%%    this action.
+%%%\end{exmp}
 

### 2016-04-29 paper.tex
Tidied section 2

@@
 from the tree.  After reaching a leaf, the enivironment continues playing
-unrestricted.  The tree in Figure~\ref{fig:agt} restricts the first and second
-environment actions to \texttt{request=1}. It then reaches a leaf of the tree
-and continues playing unrestricted.
+unrestricted.  The tree in Figure~\ref{fig:agt} restricts the first environment
+action to \texttt{request=1}. At the leaf of the tree the game continues
+unrestricted.
 
@@
     \centering
-    \begin{subfigure}[t]{.25\textwidth}
+    \begin{subfigure}[t]{.2\textwidth}
         \centering
@@
     \end{subfigure}
-    \begin{subfigure}[t]{.25\textwidth}
+    \begin{subfigure}[t]{.2\textwidth}
         \centering
@@
     \end{subfigure}
-    \begin{subfigure}[t]{.25\textwidth}
+    \begin{subfigure}[t]{.3\textwidth}
         \centering
@@
                         child {node [circle,draw] {}
-                            edge from parent [fixed] node [right] {\texttt{req=2}}
+                            edge from parent [fixed] node [below left=0pt and -5pt] {\texttt{req=2}}
                         }
@@
                         child {node [circle,draw] {}
-                            edge from parent [fixed] node [right] {\texttt{req=2}}
+                            edge from parent [fixed] node [above left=-3pt and 0pt] {\texttt{req=2}}
                         }
@@
     To solve our example, the algorithm begins by invoking a SAT solver to find
-    a trace of length $k$ from the initial state to the error state (see
-    Figure~\ref{fig:trace}). This trace is used to create an abstract game tree
-    by taking the first action in the trace and fixing it as the first action
-    in the game tree (Figure~\ref{fig:agt}).  This abstract game tree
-    represents the conjecture that \texttt{req=1} may be a good first action.
-    Next the solver attempts to find a counterexample to this conjecture by
-    finding a trace that avoids the error state with the first environment
-    action fixed to \texttt{req=1} (Figure~\ref{fig:trace2}), i.e.  by playing
-    the abstract game on behalf of the controller.
+    a trace of length $k$ (here we consider $k=3$) from the initial state to
+    the error state (see Figure~\ref{fig:trace}). This trace is used to create
+    an abstract game tree by taking the first action in the trace and fixing it
+    as the first action in the game tree (Figure~\ref{fig:agt}).  This abstract
+    game tree represents the conjecture that \texttt{req=1} may be a good first
+    action.  Next the solver attempts to find a counterexample to this
+    conjecture by finding a trace that avoids the error state with the first
+    environment action fixed to \texttt{req=1} (Figure~\ref{fig:trace2}), i.e.
+    by playing the abstract game on behalf of the controller.
 
@@
 solver to search for candidate partial strategies. Candidate strategies are
-checked by a call to the opponent solver, which also refines the abstract game
-tree. The refinements are then solved recursively.  The full procedure is
-illustrated in Algorithm~\ref{alg:bounded}.
+used as game abstractions for the opponent and can be verified by a call to the
+opponent solver. The solver recursively refines the abstraction by finding
+counterexample strategies. The full procedure is illustrated in
+Algorithm~\ref{alg:bounded}.
 
@@
 The tree is first extended to the maximum bound with edges that are labeled
-with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14) as shown
-in Figure~\ref{fig:extended}.  For each node in the tree, new SAT variables are
-introduced corresponding to the state ($X_T$) and action ($U_T$ or $C_T$)
-variables of that node. Additional variables for the opponent actions in the
-edges of $T$ are introduced ($U_e$ or $C_e$) and set to $\textsc{action}(e)$.
-The state and action variables of node $n$ are connected to successor nodes
-$\textsc{succ}(n)$ by an encoding of the transition relation and constrained to
-the winning condition of the player.
+with arbitrary opponent actions (Algorithm~\ref{alg:bounded}, line 14).  For
+each node in the tree, new SAT variables are introduced corresponding to the
+state ($X_T$) and action ($U_T$ or $C_T$) variables of that node. Additional
+variables for the opponent actions in the edges of $T$ are introduced ($U_e$ or
+$C_e$) and set to $\textsc{action}(e)$.  The state and action variables of node
+$n$ are connected to successor nodes $\textsc{succ}(n)$ by an encoding of the
+transition relation and constrained to the winning condition of the player.
 

### 2016-04-29 paper.tex
Added UPre to example in section 3

@@
 
-Consider the game tree $T$ in Figure~\ref{fig:interpolatetree} from our running
-example.  This tree is labelled with controller actions that prevent the
-environment from forcing the game into the error state. No matter how many
-requests the environment makes, the controller will grant a surplus. So the
-\textsc{findCandidate} function can not find a candidate strategy for the
-environment.  We now learn the states for which these controller actions can
-force the game to stay in the controller winning set.
+\begin{exmp}
+
+    Consider the game tree $T$ in Figure~\ref{fig:interpolatetree} from our
+    running example.  This tree is labelled with controller actions that
+    prevent the environment from forcing the game into the error state. No
+    matter how many requests the environment makes, the controller will grant a
+    surplus. So the \textsc{findCandidate} function can not find a candidate
+    strategy for the environment.  We now learn the states for which these
+    controller actions can force the game to stay in the controller winning
+    set.
+
+\end{exmp}
 
@@
 \mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
-following $T_1$ and avoiding error states. For example, in the pictured trees
-the set $nrequests = 0$ is reachable at $n$ and no move the environment chooses
-will reach the error state in one round from this set.
+following $T_1$ and avoiding error states.
+
+\begin{exmp}
+
+    In the pictured trees the interpolant $\texttt{nrequests} = 0$ is reachable
+    at $n$ and no move the environment chooses will reach an error state in one
+    round from this set. A BDD solver using an uncontrollable predecessor
+    operation on the error set would be forced to find a much more complicated
+    set: $\texttt{nrequests} = 3 \lor (\texttt{nrequests} = 2 \land (
+    \texttt{resource0} = 0 \lor \texttt{resource1} = 0)) \lor
+    (\texttt{nrequests} = 1 \land (\texttt{resource0} = 0 \land
+    \texttt{resource1} = 0))$. This set gets more complicate with more
+    resources in the system. The advantage of interpolation is that we may
+    overapproximate the losing states.
+
+\end{exmp}
 
@@
 associated bounds.  To this end, we maintain a conceptually infinite array of
-sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
-$B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an environment-losing
-set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
-n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from
-$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
+sets $B^m[k]$ that are may-losing (\textit{bad}) for the controller, indexed by
+bound $k$.  $B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an
+environment-losing set $\mathcal{I}$ is discovered for a node $n$ with bound
+$\textsc{height}(k, n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is
+subtracted from $B^m[i]$, for all $i$ less than or equal to the bound
+(lines~14--16).
 

### 2016-04-29 paper.tex
Added a missing reference to the related work

@@
 previously for synthesis in a suite of methods proposed by Bloem et al.
-\cite{bloem2014}. The authors propose employing competing SAT solvers
-to learn clauses, which is similar to our approach but does not unroll the
-game. They also suggest QBF solver, template-based, and Effectively
+\cite{bloem2014}. The authors propose employing competing SAT solvers to learn
+clauses representing bad states, which is similar to our approach but does not
+unroll the game.  They also suggest QBF solver, template-based, and Effectively
 Propositional Logic (EPR) approaches.
@@
 Incremental induction \cite{bradley2011} is another technique for unbounded
-model checking with an equivalent synthesis method \cite{morgenstern2013},
-which computes sets of states that overapproximate the losing states (similar
-to our $B^m$) and another set of winning states (similar to the negation of
-$B^M$). Their algorithm maintains a similar invariant over the sets of losing
-states as our approach and has the same termination condition. It differs in
-how the sets are computed, which it does by inductively proving the number of
-game rounds required by the environment to win from the initial state.
+model checking that inspired several approaches to synthesis including the work
+presented here.  Morgenstern et al.~\cite{morgenstern2013} proposed an
+technique that computes sets of states that overapproximate the losing states
+(similar to our $B^m$) and another set of winning states (similar to the
+negation of $B^M$).  Their algorithm maintains a similar invariant over the
+sets of losing states as our approach and has the same termination condition.
+It differs in how the sets are computed, which it does by inductively proving
+the number of game rounds required by the environment to win from a state.
+Chiang and Jiang~\cite{Chiang2015} recently proposed a similar approach that
+focusses on computing the winning region for the controller forwards from the
+initial state in order to take advantage of reachability information and bad
+transition learning without needing to discard learnt clauses.
 

### 2016-05-01 paper.tex
Fixes to section 2 & 3

@@
 \newtheorem*{exmp}{Example}
+\newtheorem*{exmpI}{Example: Intuition behind the algorithm}
+\newtheorem*{exmpInt}{Example: Why we use interpolants}
 
@@
 \begin{exmp}
-We introduce a running example to assist the explanation. We consider a
-simple arbiter system in which the environment requests for either 1 or 2
-resources, and the controller may grant access to two resources. The total
-number of requests grows each round by the number of environment requests and
-shrinks by the number of resources granted by the controller in the previous
-round.  The controller must ensure that the number of unhandled requests does
-not accumulate to more than 2.  Figure~\ref{fig:example} shows the variables
-(\ref{fig:examplevars}) and the formulas for computing next-state variables
-(\ref{fig:exampletrans}) for this example. We use primed identifiers to denote
-next-state variables and curly braces to define the domain of a variable.
+
+We introduce a running example to assist the explanation. We consider a simple
+arbiter system in which the environment makes a request for a number of
+resources (1 or 2), and the controller may grant access to up to two resources.
+The total number of requests grows each round by the number of environment
+requests and shrinks by the number of resources granted by the controller in
+the previous round.  The controller must ensure that the number of unhandled
+requests does not accumulate to more than 2.  Figure~\ref{fig:example} shows
+the variables (\ref{fig:examplevars}) and the formulas for computing next-state
+variables (\ref{fig:exampletrans}) for this example. We use primed identifiers
+to denote next-state variables and curly braces to define the domain of a
+variable.
 
 This example is the $n=2$ instance of the more general problem of an arbiter of
-$n$ resources. The set of winning states grows exponentially with the number of
-resources the controller could use to grant requests, which makes the problem
-hard for BDD solvers. In Section~3 we will outline how this example can be
-solved without enumerating all winning states.
+$n$ resources. The set of winning states has no compact representation, which
+makes the problem hard for BDD solvers. In Section~3 we will outline how this
+example can be solved without enumerating all winning states.
 
@@
 
-The game abstraction is initially constructed from an opponent candidate
-strategy. It is refined by counterexamples to the solver's own candidate
-strategy that are formed when the opponent solves the abstract game
-corresponding to that strategy. By abstracting the game in this way we
-conceptually direct the solvers' search to consider opponent actions that may
-lead to winning strategies.
-
 An abstraction of the game restricts actions available to one of the players.
@@
     \end{subfigure}
-    \begin{subfigure}[t]{.1\textwidth}
+    \begin{subfigure}[t]{.15\textwidth}
         \centering
@@
                             }
-                            node [left=4pt] {\texttt{gr0=0 gr1=1}}
+                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                         }
@@
                     }
-                node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                node [left=4pt] {\texttt{gr0=1 gr1=1}}
                 }
@@
                         % node [left=4pt] {\texttt{gr0=1 gr1=0}}
-                        edge from parent [fixed] node [left] {\texttt{req=1}}
+                        edge from parent [fixed] node [left] {\texttt{req=2}}
                     }
@@
     \end{subfigure}
-    \begin{subfigure}[t]{.3\textwidth}
+    \begin{subfigure}[t]{.2\textwidth}
         \centering
@@
                         child {node [circle,draw] {}
-                            edge from parent [fixed] node [left] {\texttt{req=1}}
-                        }
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [below left=0pt and -5pt] {\texttt{req=2}}
-                        }
-                        edge from parent [fixed] node [left] {\texttt{req=1}}
-                    }
-                    child {node [circle,draw] {}
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [above left=-3pt and 0pt] {\texttt{req=2}}
+                            edge from parent [fixed] node [right] {\texttt{req=2}}
                         }
@@
                 }
-                node [above=4pt] {$\langle s, 2 \rangle$};
+                node [above=4pt] {$\langle s, 3 \rangle$};
         \end{tikzpicture}
@@
 
-\begin{exmp}
+\begin{exmpI}
+
+    %% Make it clear that this is intuition only
 
     To solve our example, the algorithm begins by invoking a SAT solver to find
-    a trace of length $k$ (here we consider $k=3$) from the initial state to
-    the error state (see Figure~\ref{fig:trace}). This trace is used to create
-    an abstract game tree by taking the first action in the trace and fixing it
-    as the first action in the game tree (Figure~\ref{fig:agt}).  This abstract
-    game tree represents the conjecture that \texttt{req=1} may be a good first
-    action.  Next the solver attempts to find a counterexample to this
-    conjecture by finding a trace that avoids the error state with the first
-    environment action fixed to \texttt{req=1} (Figure~\ref{fig:trace2}), i.e.
-    by playing the abstract game on behalf of the controller.
-
-    Observe that the controller chooses to grant only one resource, just enough
-    to ensure that an error state will not be reached. This example will show
-    the worst case when the SAT solver returns actions that will lead to large
-    abstract game trees. However, in practice we employ heuristics to guide the
-    SAT solver away from the worst case.
-
-    The environment's candidate is refined using the controller's
-    counterexample traces. In Figure~\ref{fig:agt2}, the game tree tree is
-    refined based on the first action of the controller's trace. The
-    \texttt{req=1} action in the second round is still winning under the
-    assumption that the controller will not grant any resources. The
-    environment is free to make that assumption because the controller's
-    counterexample does not yet contain any grants. Eventually, the refinements
-    to the environment's candidate will force the controller to grant more
-    resources and result in a game tree such as in Figure~\ref{fig:refined}.
-
-\end{exmp}
-
-The bounded synthesis algorithm solves abstract game trees by invoking a SAT
-solver to search for candidate partial strategies. Candidate strategies are
-used as game abstractions for the opponent and can be verified by a call to the
-opponent solver. The solver recursively refines the abstraction by finding
-counterexample strategies. The full procedure is illustrated in
-Algorithm~\ref{alg:bounded}.
+    a trace of length $k$ (here we consider $k=3$) from the initial state to an
+    error state (see Figure~\ref{fig:trace}).  Given a trace we make an initial
+    conjecture that the first action, \texttt{req=1}, is winning for the
+    environment. This conjecture is captured in the abstract game tree show in
+    Figure~\ref{fig:agt}.  Next the solver attempts to find a counterexample to
+    this conjecture by finding a trace that avoids the error states with the
+    first environment action fixed to \texttt{req=1}, i.e. by playing the
+    abstract game on behalf of the controller.
+
+    The resulting trace is shown in Figure~\ref{fig:trace2}. The environment's
+    candidate is refined using this trace and Figure~\ref{fig:agt2} shows the
+    refined game tree.  The controller's counterexample was to grant two
+    resources in the first round, to which the environment responds by now
+    making two requests in the second round.
+
+    Eventually, the refinements to the environment's candidate will force the
+    controller to grant more resources and result in a game tree such as in
+    Figure~\ref{fig:refined}.
+
+\end{exmpI}
 
@@
 
-The algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
+The full procedure is illustrated in Algorithm~\ref{alg:bounded}.  The
+algorithm takes a concrete game $G$ with maximum bound $\kappa$ as an implicit
 argument.  In addition, it takes a player $p$ (controller or environment),
@@
 is appended to the corresponding node in $T$ in order to refine it in line~(9).
-
-%%%\begin{exmp}
-%%%    The controller's partial strategy in
-%%%    Figure~\ref{fig:strategy} is defeated by the counterexample action
-%%%    \texttt{req=2}.  Figure~\ref{fig:refined} shows a refined AGT that includes
-%%%    this action.
-%%%\end{exmp}
 
@@
 We instead present an approach that iteratively solves games of increasing
-bound while learning winning states from abstract games using interpolation. We
+bound while learning bad states from abstract games using interpolation. We
 show that reaching a fixed point in learned states is sufficient for
@@
 procedure kicks in whenever \textsc{findCandidate} cannot find a candidate
-strategy for an abstract game tree $T$ with initial state $s$ and bound $k$.
-For the controller, this means the environment partial strategy represented by $T$
-will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
-bound is irrelevant).  For the environment, the controller can avoid $E$ for $k$
-rounds no matter the assignments to $U$ variables. Hence, we have established
-that $s$ is a losing state either always (for the controller) or for bound $k$ (for
-the environment. We can learn additional losing states from the tree via
-interpolation.  This is achieved in lines~18--20 in
+strategy for an abstract game tree. We can learn additional losing states from
+the tree via interpolation.  This is achieved in lines~18--20 in
 Algorithm~\ref{alg:bounded}, enabled in the unbounded version of the algorithm,
@@
 
+%%%For the controller, this means the environment partial strategy represented by $T$
+%%%will always reach $E$ from $s$ no matter the assignments to $C$ variables (the
+%%%bound is irrelevant).  For the environment, the controller can avoid $E$ for $k$
+%%%rounds no matter the assignments to $U$ variables. Hence, we have established
+%%%that $s$ is a losing state either always (for the controller) or for bound $k$
+%%%(for the environment).
+
 \begin{figure}[t]
@@
                     child {node [circle,draw,right=5pt] {}
-                        edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
+                        edge from parent [fixed] node [left=2pt,text width=1cm] {\texttt{gr0=1 gr1=0}}
                     }
                     child {node [circle,draw,left=5pt] {}
-                        edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
+                        edge from parent [fixed] node [right=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
                     }
                     node [left=5pt] {$n$}
-                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=0, gr1=1}}
+                    edge from parent [fixed] node [left=2pt,text width=1cm] {\texttt{gr0=0 gr1=1}}
                 }
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
-                    }
-                    edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
+                        edge from parent [fixed] node [right=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
+                    }
+                    edge from parent [fixed] node [right=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
                 }
@@
                     node [left=5pt] {$n$}
-                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
+                    edge from parent [fixed] node [left=2pt,text width=1cm] {\texttt{gr0=1 gr1=0}}
                 }
@@
                     child {node [circle,draw] {}
-                        edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
-                    }
-                    edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
+                        edge from parent [fixed] node [right=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
+                    }
+                    edge from parent [fixed] node [right=2pt=2pt=2pt=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
                 }
@@
                 child {node [circle,draw,right=5pt] {}
-                    edge from parent [fixed] node [left,text width=1cm] {\texttt{gr0=1, gr1=0}}
+                    edge from parent [fixed] node [left=2pt,text width=1cm] {\texttt{gr0=1 gr1=0}}
                 }
                 child {node [circle,draw,left=5pt] {}
-                    edge from parent [fixed] node [right,text width=1cm] {\texttt{gr0=1, gr1=1}}
+                    edge from parent [fixed] node [right=2pt,text width=1cm] {\texttt{gr0=1 gr1=1}}
                 }
@@
 
-\begin{exmp}
-
-    Consider the game tree $T$ in Figure~\ref{fig:interpolatetree} from our
-    running example.  This tree is labelled with controller actions that
-    prevent the environment from forcing the game into the error state. No
-    matter how many requests the environment makes, the controller will grant a
-    surplus. So the \textsc{findCandidate} function can not find a candidate
-    strategy for the environment.  We now learn the states for which these
-    controller actions can force the game to stay in the controller winning
-    set.
-
-\end{exmp}
+\begin{exmpInt}
+
+    Consider node $n$ in Figure~\ref{fig:interpolatetree}. At this node there
+    are two controller actions that prevent the environment from forcing the
+    game into an error state in one game round. We want to use this tree to
+    learn the states from which the controller can win playing one of these
+    actions.
+
+    One option is using a BDD solver, working backwards from the error set, to
+    find all losing states. One iteration of this operation on our example
+    would give the set: $\texttt{nrequests} = 3 \lor (\texttt{nrequests} = 2
+    \land ( \texttt{resource0} = 0 \lor \texttt{resource1} = 0)) \lor
+    (\texttt{nrequests} = 1 \land (\texttt{resource0} = 0 \land
+    \texttt{resource1} = 0))$.  In the general case there is no compact
+    representation of the losing set, so we try to avoid computing it by
+    employing interpolation instead. The potential benefit of interpolation is
+    that an interpolant can approximate the losing states in such a way that
+    there is just enough information to prove realisability. It is also cheap
+    to compute an interpolant from the proof of unsatisfiability of a SAT
+    solver.
+
+    At node $n$, the interpolant $\texttt{nrequests} = 1 \land
+    \texttt{resource1} = 1$ captures the information we need. Any action by the
+    environment followed by one of the controller actions at $n$ will be
+    winning for the controller.
+
+\end{exmpInt}
 
@@
 following $T_1$ and avoiding error states.
-
-\begin{exmp}
-
-    In the pictured trees the interpolant $\texttt{nrequests} = 0$ is reachable
-    at $n$ and no move the environment chooses will reach an error state in one
-    round from this set. A BDD solver using an uncontrollable predecessor
-    operation on the error set would be forced to find a much more complicated
-    set: $\texttt{nrequests} = 3 \lor (\texttt{nrequests} = 2 \land (
-    \texttt{resource0} = 0 \lor \texttt{resource1} = 0)) \lor
-    (\texttt{nrequests} = 1 \land (\texttt{resource0} = 0 \land
-    \texttt{resource1} = 0))$. This set gets more complicate with more
-    resources in the system. The advantage of interpolation is that we may
-    overapproximate the losing states.
-
-\end{exmp}
 

### 2016-05-01 paper.tex
Added an explicit proof of solveUnbounded to section 3.3

@@
 
-Both invariants trivially hold after $B^m$ and $B^M$ have been initialised in the
-beginning of the algorithm.
-The sets $B^m$ and $B^M$ are only modified by the functions \textsc{learn} and
-\textoverline{\textsc{learn}}.  Below we prove that \textoverline{\textsc{learn}} 
-maintains the invariants.  The proof of \textsc{learn} is similar.
+Correctness of $\textsc{solveUnbounded}$ follows from these invariants. The
+must-invariant guarantees that the environment can force the game into an error
+state from $B^M$, therefore checking whether the initial state is in $B^M$ (as
+in line~5) is sufficient to return \texttt{unrealisable}. The may-invariant
+tells us that if $B^m[i] \equiv B^m[i+1]$ (line~6) then $Upre(B^m[i+1])
+\subseteq B^m[i+1]$, i.e. there is a fixed point in the uncontrollable
+predecessor. This guarantees that the controller can force the game to stay
+within $B^m[i+1]$ and avoid error states indefinitely.
+
+Both invariants trivially hold after $B^m$ and $B^M$ have been initialised in
+the beginning of the algorithm. The sets $B^m$ and $B^M$ are only modified by
+the functions \textsc{learn} and \textoverline{\textsc{learn}}.  Below we prove
+that \textoverline{\textsc{learn}} maintains the invariants.  The proof of
+\textsc{learn} is similar.
 

### 2016-05-01 paper.tex
Added a sentence about the parallel track & unique instances

@@
 Our implementation was able to solve 103 out of the 250 specification in the
-alloted time, including 12 instances that were not solved by any other
-solver in SYNTCOMP'15. The unique instances we solved are listed in
-Table~\ref{tab:unique}.
+alloted time, including 12 instances that were not solved by any other solver
+in the sequential track of the competition. The unique instances we solved are
+listed in Table~\ref{tab:unique}.
 
@@
 relay unsafe states to one another, which is compatible with the set $B^M$ in
-our solver. It remains future work to explore the possibility of combining
-these solvers.
+our solver. This technique can already solve each of the unique instances
+solved by our solver but there may still be value in the addition of this work
+to the portfolio.  It remains future work to explore this possibility.
 

### 2016-05-02 paper.tex
Added AbySynthe & Simple BDD Solver benchmarks from cluster

@@
         %Plots
-        \draw plot[mark=x, mark options={fill=white}] file {bench_termite.dat};
-        \draw plot[mark=+, mark options={fill=white}] file {bench_demiurge.dat};
-        \draw plot[mark=triangle, mark options={fill=white}] file {bench_simple.dat};
+        \draw plot[mark=o, mark size=1pt, mark options={fill=white,color=red}] file {bench_termite.dat};
+        \draw plot[mark=+, mark options={fill=white,color=blue}] file {bench_demiurge.dat};
+        \draw plot[mark=triangle, mark options={fill=white}] file {bench_simple2.dat};
+        \draw plot[mark=x, mark options={fill=white,color=green}] file {bench_abs1.dat};
+%%%        \draw plot[mark=x, mark options={fill=white,color=green}] file {bench_abs2.dat};
+%%%        \draw plot[mark=x, mark options={fill=white,color=green}] file {bench_abs3.dat};
 
@@
             \draw (0,0) -- 
-                plot[mark=x, mark options={fill=white}] (0.25,0) -- (0.5,0) 
+                plot[mark=o, mark size=1pt, mark options={fill=white,color=red}] (0.25,0) -- (0.5,0) 
                 node[right]{Unbounded synthesis};
             \draw[yshift=\baselineskip] (0,0) -- 
-                plot[mark=+, mark options={fill=white}] (0.25,0) -- (0.5,0)
+                plot[mark=x, mark options={fill=white,color=green}] (0.25,0) -- (0.5,0)
+                node[right]{AbsSynthe (seq2)};
+            \draw[yshift=2\baselineskip] (0,0) -- 
+                plot[mark=+, mark options={fill=white,color=blue}] (0.25,0) -- (0.5,0)
                 node[right]{Demiurge};
-            \draw[yshift=2\baselineskip] (0,0) -- 
+            \draw[yshift=3\baselineskip] (0,0) -- 
                 plot[mark=triangle, mark options={fill=white}] (0.25,0) -- (0.5,0)

### 2016-05-04 paper.tex
Align figure 2 properly
Minor edits

@@
 The reactive synthesis problem is EXPTIME-complete so
-naive algorithms are infeasible on anything but simple systems.
+na\"ive algorithms are infeasible on anything but simple systems.
 % In this way the environment is equivalent to an existential search of
@@
 any other solver.  These results confirm that there exist classes of problems that are
-hard for traditional synthesis techniques, but can be efficiently solved by our method.
-While further performance improvements are clearly needed, in its current state the method 
-can be used in a portfolio solver to increase the number of tractable instances.
+hard for traditional synthesis techniques, but can be efficiently solved by our
+method.  While further performance improvements are clearly needs, in its
+current state the method may be a worthwhile addition to a portfolio solver.
+%%%While further performance improvements are clearly needed, in its current state the method 
+%%%can be used in a portfolio solver to increase the number of tractable instances.
 
@@
 $n$ resources. The set of winning states has no compact representation, which
-makes the problem hard for BDD solvers. In Section~3 we will outline how this
-example can be solved without enumerating all winning states.
+makes the problem hard for BDD solvers. In Section~3 we will outline how the
+unbounded game can be solved without enumerating all winning states.
 
@@
     \centering
+    \captionsetup[subfigure]{width=\textwidth,justification=raggedleft}
     \begin{subfigure}[t]{.2\textwidth}
         \centering
-        \begin{tikzpicture}[level distance = 5mm]
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \begin{tikzpicture}[level distance = 5mm,baseline]
             \node [circle,draw,inner sep=1pt] (root){}
@@
         \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
-        \caption{Environment-winning trace}
+    \end{minipage}
+        \caption{Environment winning trace}
         \label{fig:trace}
@@
         \centering
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm]
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
             \node [circle,draw] (root){}
@@
         \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
+        \end{minipage}
         \caption{AGT}
@@
         \centering
-        \begin{tikzpicture}[level distance = 5mm]
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \begin{tikzpicture}[level distance = 5mm,baseline]
             \node [circle,draw,inner sep=1pt] (root){}
@@
         \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
-        \caption{Controller-winning trace}
+        \end{minipage}
+        \caption{Controller winning trace}
         \label{fig:trace2}
@@
         \end{tikzpicture}
-        \captionsetup{width=.8\textwidth}
+        \end{minipage}
         \caption{Refined AGT}
         \label{fig:refined}
-    \end{subfigure}%
+    \end{subfigure}
     \caption{Abstract game trees.}
@@
     conjecture that the first action, \texttt{req=1}, is winning for the
-    environment. This conjecture is captured in the abstract game tree show in
+    environment. This conjecture is captured in the abstract game tree shown in
     Figure~\ref{fig:agt}.  Next the solver attempts to find a counterexample to
@@
     there is just enough information to prove realisability. It is also cheap
-    to compute an interpolant from the proof of unsatisfiability of a SAT
-    solver.
+    to compute an interpolant using a SAT solver.
 
@@
 \end{exmpInt}
+
+Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
+it is possible to construct a Craig interpolant~\cite{craig1957} $\mathcal{I}$
+such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
+$\mathcal{I}$ refers only to the intersection of variables in $F_1$ and $F_2$.
+An interpolant can be constructed efficiently from a resolution proof of the
+unsatisfiability of $F_1 \land F_2$~\cite{pudlak1997}.
 
@@
 \textsc{\textoverline{treeFormula}}(k, T_1) \land
-\textsc{\textoverline{treeFormula}}(k, T_2)$ is also unsatisfiable. This enables
-the construction of an \emph{interpolant} that captures losing states at node
-$n$.
-
-Given two formulas $F_1$ and $F_2$ such that $F_1 \land F_2$ is unsatisfiable,
-it is possible to construct a Craig interpolant~\cite{craig1957} $\mathcal{I}$
-such that $F_1 \to \mathcal{I}$, $F_2 \land \mathcal{I}$ is unsatisfiable, and
-$\mathcal{I}$ refers only to the intersection of variables in $F_1$ and $F_2$.
-An interpolant can be constructed efficiently from a resolution proof of the
-unsatisfiability of $F_1 \land F_2$~\cite{pudlak1997}.
-
-We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(k, T_1)$
-and $F_2 = \textsc{treeFormula}(k, T_2)$ (line~5). The only variables shared between $F_1$
-and $F_2$ are the state variable copies belonging to node $n$. By the properties of the interpolant,
-$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
-are losing against abstract game tree $T_2$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
-\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
+\textsc{\textoverline{treeFormula}}(k, T_2)$ is also unsatisfiable.
+
+%%%This enables the construction of an interpolant that captures losing states at
+%%%node $n$.
+
+We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(k,
+T_1)$ and $F_2 = \textsc{treeFormula}(k, T_2)$ (line~5). The only variables
+shared between $F_1$ and $F_2$ are the state variable copies belonging to node
+$n$. By the properties of the interpolant, $F_2 \land \mathcal{I}$ is
+unsatisfiable, therefore all states in $\mathcal{I}$ are losing against
+abstract game tree $T_2$ in Figure~\ref{fig:treef2}.  We also know that $F_1
+\to \mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
 following $T_1$ and avoiding error states.
@@
 associated bounds.  To this end, we maintain a conceptually infinite array of
-sets $B^m[k]$ that are may-losing (\textit{bad}) for the controller, indexed by
-bound $k$.  $B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an
-environment-losing set $\mathcal{I}$ is discovered for a node $n$ with bound
-$\textsc{height}(k, n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is
-subtracted from $B^m[i]$, for all $i$ less than or equal to the bound
-(lines~14--16).
+sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
+$B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an environment-losing
+set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
+n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from
+$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).
 
@@
 reference we plotted our results along with \textsc{Demiurge}~\cite{bloem2014},
-the only SAT-based tool in the competition, and the winner of the sequential
-realisability track \textsc{Simple BDD Solver 2}~\cite{walker2014}. We were
-able to run \textsc{Demiurge} on our hardware but show competition results for
-\textsc{Simple BDD Solver 2}.
+the only SAT-based tool in the competition, the winner of the sequential
+realisability track \textsc{Simple BDD Solver 2}~\cite{walker2014}, and
+AbsSynthe (seq3) \cite{brenguier2014}. Our results differ from the competition,
+we suspect due to the weaker hardware.
 

### 2016-05-04 paper.tex
Another edit to the proof section

@@
 in line~5) is sufficient to return \texttt{unrealisable}. The may-invariant
-tells us that if $B^m[i] \equiv B^m[i+1]$ (line~6) then $Upre(B^m[i+1])
-\subseteq B^m[i+1]$, i.e. there is a fixed point in the uncontrollable
-predecessor. This guarantees that the controller can force the game to stay
-within $B^m[i+1]$ and avoid error states indefinitely.
+tells us that if $B^m[i] \equiv B^m[i+1]$ (line~6) then $Upre(B^m[i]) \subseteq
+B^m[i]$, i.e. $B^m[i]$ overapproximates the winning states for the environment.
+We know that $I \not\in B^m[k]$ due to the post-condition of
+\textsc{checkBound}, and since the may-invariant tells us that $B^m$ is
+monotonic then $I$ must not be in $B^m[i]$. If $I \not\in B^m[i]$ then $I$ is
+not in the winning states for the environment and the controller can always win
+from $I$.
 

### 2016-05-05 paper.tex
More minor edits

@@
 (SYNTCOMP'15).  While our solver solves fewer total instances than competitors,
-it solves the largest number of unique instances, i.e., instances that could not be solved by 
-any other solver.  These results confirm that there exist classes of problems that are
-hard for traditional synthesis techniques, but can be efficiently solved by our
-method.  While further performance improvements are clearly needs, in its
-current state the method may be a worthwhile addition to a portfolio solver.
+it solves the largest number of unique instances for a sequential solver, i.e.,
+instances that could not be solved by any other solver.  These results confirm
+that there exist classes of problems that are hard for traditional synthesis
+techniques, but can be efficiently solved by our method.  While further
+performance improvements are clearly needed, in its current state the method
+may be a worthwhile addition to a portfolio solver.
 %%%While further performance improvements are clearly needed, in its current state the method 
@@
     to compute an interpolant using a SAT solver.
-
-    At node $n$, the interpolant $\texttt{nrequests} = 1 \land
-    \texttt{resource1} = 1$ captures the information we need. Any action by the
-    environment followed by one of the controller actions at $n$ will be
-    winning for the controller.
-
 \end{exmpInt}
@@
 
+\begin {exmp}
+
+    At node $n$, the interpolant $\texttt{nrequests} = 1 \land
+    \texttt{resource1} = 1$ captures the information we need. Any action by the
+    environment followed by one of the controller actions at $n$ will be
+    winning for the controller.
+
+\end{exmp}
+
 We have discovered a set $\mathcal{I}$ of states losing for the environment.
@@
 sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.
-$B^m[k]$ are initialised to $\top$ for all $k$.  Whenever an environment-losing
+$B^m[k]$ are initialised to $E$ for all $k$.  Whenever an environment-losing
 set $\mathcal{I}$ is discovered for a node $n$ with bound $\textsc{height}(k,
@@
 for SAT solving and \textsc{Periplo}~\cite{rollini2013} for interpolant
-generation. The benchmarks were run on a cluster of Intel Quad Core Xeon E5405
-2GHz CPUs with 16GB of memory.  The solver was allowed exclusive access to a node
-for one hour to solve an instance.
+generation. We intend to open source the tool for SYNTCOMP'16. The benchmarks
+were run on a cluster of Intel Quad Core Xeon E5405 2GHz CPUs with 16GB of
+memory.  The solver was allowed exclusive access to a node for one hour to
+solve an instance.
 
 The results of our benchmarking are shown, along with the synthesis competition
-results~\cite{syntcompedacc}, in Table~\ref{tab:syntcomp}. The competition was run on Intel Quad
-Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes for one hour per
-instance.
+results~\cite{syntcompedacc}, in Table~\ref{tab:syntcomp}. The competition was
+run on Intel Quad Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes
+for one hour per instance. The unique column lists the number of instances that
+only that tool could solve in the competition (excluding our solver). In
+brackets is the number of instances that only that tool could solve, including
+our solver.
 
@@
         \hline
-        Simple BDD Solver (2) & 195 & 10 \\
+        Simple BDD Solver (2) & 195 & 10 \textit{(6)} \\
         AbsSynthe (seq2) & 187 & 2 \\
@@
         AbsSynthe (seq1) & 173 & 1 \\
-        Demiurge (D1real) & 139 & 5 \\
+        Demiurge (D1real) & 139 & 5 \textit{(2)} \\
         Aisy & 98 & \\
@@
 extraction. The technique relies on interpolation over abstract game trees to
-compute the winning strategy.  In
-the present work we use interpolation in a different way in order to learn
-losing states of the game.
+compute the winning strategy.  In the present work we use interpolation in a
+different way in order to learn losing states of the game. In addition, this
+method could be easily adapted to the unbounded realisability algorithm
+presented here to generate unbounded strategies.
 

### 2016-05-05 paper.tex
Add my ridiculous new affiliation

@@
 
-\institute{NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
+    \institute{Data61, CSIRO (formerly NICTA\thanks{NICTA is funded by the Australian Government as represented by the Department of Broadband,
     Communications and the Digital Economy and the Australian Research Council through the ICT
-    Centre of Excellence program.} and UNSW \\
+Centre of Excellence program.}) and UNSW \\
     \email{alexander.legg@nicta.com.au}

### 2016-05-08 paper.tex
Leonid's edits

@@
 
-The reactive synthesis problem is EXPTIME-complete so
+The reactive synthesis problem is EXPTIME-complete (when starting 
+from a symbolic representation of the game graph) so
 na\"ive algorithms are infeasible on anything but simple systems.
@@
 (SYNTCOMP'15).  While our solver solves fewer total instances than competitors,
-it solves the largest number of unique instances for a sequential solver, i.e.,
-instances that could not be solved by any other solver.  These results confirm
+it solves the largest number of unique instances, i.e.,
+instances that could not be solved by any other sequential solver.  
+These results confirm
 that there exist classes of problems that are hard for traditional synthesis
@@
 over boolean state variables $X$, uncontrollable action variables $U$, and
-controllable action variables $C$.  $I$ is the initial state of the game given
+controllable action variables $C$.  
+We use $\mathcal{X}$, $\mathcal{U}$, and $\mathcal{C}$ to denote sets of valuations of 
+variables $X$, $U$, and $C$ respectively.
+$I$ is the initial state of the game given
 as a valuation of state variables.  $E(X)$ is the set of error states
@@
 
-A \emph{controller strategy} $\pi^c : (2^X, 2^U) \to 2^C$ is a mapping of states and
+A \emph{controller strategy} $\pi^c : \mathcal{X} \times \mathcal{U} \to \mathcal{C}$ 
+is a mapping of states and
 uncontrollable inputs to controllable actions. A controller strategy is winning
@@
 
-An \emph{environment strategy} $\pi^e : 2^X \to 2^U$ is a mapping of states to
+An \emph{environment strategy} $\pi^e : \mathcal{X} \to \mathcal{U}$ is a mapping of states to
 uncontrollable actions. A bounded run is winning for the environment if $x_0 =
@@
 requests does not accumulate to more than 2.  Figure~\ref{fig:example} shows
-the variables (\ref{fig:examplevars}) and the formulas for computing next-state
-variables (\ref{fig:exampletrans}) for this example. We use primed identifiers
+the variables (\ref{fig:examplevars}), the initial state of the system (\ref{fig:exampleinit}), 
+and the formulas for computing next-state
+variable assignments (\ref{fig:exampletrans}) for this example. We use primed identifiers
 to denote next-state variables and curly braces to define the domain of a
@@
 This example is the $n=2$ instance of the more general problem of an arbiter of
-$n$ resources. The set of winning states has no compact representation, which
+$n$ resources. For large values of $n$, the set of winning states has no compact representation, which
 makes the problem hard for BDD solvers. In Section~3 we will outline how the
@@
 
-The bounded synthesis algorithm consists of two competing solvers, one for the
-controller and one for the environment, that solve abstractions of the safety
-game. Each solver builds a candidate strategy for its corresponding player
-consisting of assignments to $2^C$ and $2^U$ respectively.
-
-An abstraction of the game restricts actions available to one of the players.
+Our bounded synthesis algorithm constructs abstractions of the game 
+that restrict actions available to one of the players.
 Specifically, we consider abstractions represented as trees of actions,
@@
 can freely choose actions whilst the environment is required to pick actions
-from the tree.  After reaching a leaf, the enivironment continues playing
+from the tree.  After reaching a leaf, the environment continues playing
 unrestricted.  The tree in Figure~\ref{fig:agt} restricts the first environment
@@
 Given an environment (controller) abstract game tree $T$ a \emph{partial
-strategy} $Strat: \textsc{nodes}(T) \rightarrow C$ ($Strat: \textsc{nodes}(T)
-\rightarrow U$) labels each node of the tree with the controller's
+strategy} $Strat: \textsc{nodes}(T) \rightarrow \mathcal{C}$ ($Strat: \textsc{nodes}(T)
+\rightarrow \mathcal{U}$) labels each node of the tree with the controller's
 (environment's) action to be played in that node.   Given a partial strategy
@@
 playing all controllable and uncontrollable actions on the path from the root
-to the leaf.
+to the leaf.  An environment (controller) partial strategy is \emph{winning against $T$} 
+if all its outcomes are states that are winning for the environment (controller)
+in the concrete game.
 
@@
     %% Make it clear that this is intuition only
-
-    To solve our example, the algorithm begins by invoking a SAT solver to find
-    a trace of length $k$ (here we consider $k=3$) from the initial state to an
-    error state (see Figure~\ref{fig:trace}).  Given a trace we make an initial
-    conjecture that the first action, \texttt{req=1}, is winning for the
+    We present the intuition behind our bounded synthesis method by applying
+    its \emph{simplified version} to the running example.
+    We begin by finding 
+    a trace of length $k$ (here we consider $k=3$) that is winning for the environment, 
+    i.e., that starts from the initial state and reaches an
+    error state (see Figure~\ref{fig:trace}).  We use a SAT solver to find such a trace,
+    precisely as one would do in bounded model checking.
+    Given this trace we make an initial
+    conjecture that any trace starting with action \texttt{req=1} is winning for the
     environment. This conjecture is captured in the abstract game tree shown in
-    Figure~\ref{fig:agt}.  Next the solver attempts to find a counterexample to
-    this conjecture by finding a trace that avoids the error states with the
-    first environment action fixed to \texttt{req=1}, i.e. by playing the
-    abstract game on behalf of the controller.
-
-    The resulting trace is shown in Figure~\ref{fig:trace2}. The environment's
-    candidate is refined using this trace and Figure~\ref{fig:agt2} shows the
-    refined game tree.  The controller's counterexample was to grant two
+    Figure~\ref{fig:agt}.  We validate this conjecture by searching for a counterexample
+    trace that avoids the error state with the
+    first environment action fixed to \texttt{req=1}.   
+    Such a trace, that refutes the conjecture, is shown in Figure~\ref{fig:trace2}. 
+    In this trace, the controller wins by responding with \texttt{gr0=1, gr1=1} in the first round.
+    This move represents controller's partial strategy against the 
+    abstract game tree in Figure~\ref{fig:agt}.  This partial strategy is shown in Figure~\ref{}.
+
+    Next we strengthen the abstract game tree taking this partial strategy into account.
+    To this end we again use a SAT solver to find a trace where the environment wins while
+    the controller plays according to the partial strategy.  In the resulting trace (now shown), 
+    the environment plays \texttt{req=2} in the second round.  We refine the abstract game tree
+    using this move as shown in Figure~\ref{fig:agt2}.  The controller's partial strategy was to grant two
     resources in the first round, to which the environment responds by now
@@
 
-    Eventually, the refinements to the environment's candidate will force the
-    controller to grant more resources and result in a game tree such as in
-    Figure~\ref{fig:refined}.
+    When the environment cannot refine the tree by extending existing branches, it backtracks 
+    and creates new branches.
+    Eventually, we obtain the abstract game tree shown in Figure~\ref{fig:refined} for which 
+    there does not exist a winning partial strategy on behalf of the controller.  We conclude that 
+    the game is winning for the environment.
+
+%    Eventually, the refinements to the environment's candidate will force the
+%    controller to grant more resources and result in a game tree such as in
+%    Figure~\ref{fig:refined}.
 
@@
 
+The full bounded synthesis algorithm is more complicated: upon finding a candidate 
+partial strategy on behalf of player $p$ against abstract game tree $T$, it first checks whether the strategy is winning 
+against $T$.  By only considering such strong candidates, we reduce the number of 
+refinements needed to solve the game.  To this end, the algorithm checks whether each 
+outcome of the candidate strategy is a winning state for $\textsc{opponent}(p)$ by recursively 
+invoking the synthesis algorithm on behalf of the opponent.  Thus, our bounded
+synthesis algorithm can be seen as running two competing solvers, for the
+controller and for the environment.
+
+%The solvers build candidate partial strategy 
+%for their corresponding players, which are used to refine the abstractions.
+
 The full procedure is illustrated in Algorithm~\ref{alg:bounded}.  The
@@
 winning when the opponent is restricted to $T$.  If it fails to find a
-strategy, this means that no winning strategy exists against the opponent
-playing its partial strategy represented by $T$.  If, on the other hand, a
+strategy, this means that no winning partial strategy exists against the opponent
+playing according to $T$.  If, on the other hand, a
 candidate partial strategy is found, we need to verify if it is indeed winning
@@
 \textsc{solveAbstract} for the opponent. The dual solver solves games on behalf
-of the controller player.
+of the opponent player.
 
@@
 considered. The na\"ive approach is to use size of the state space as the bound
-($2^{|X|}$) so that all states may be explored by the algorithm. A more nuanced
+(${|\mathcal{X}|}$) so that all states may be explored by the algorithm. A more nuanced
 approach is to use the diameter of the game \cite{biere1999}, which is the
@@
     representation of the losing set, so we try to avoid computing it by
-    employing interpolation instead. The potential benefit of interpolation is
-    that an interpolant can approximate the losing states in such a way that
-    there is just enough information to prove realisability. It is also cheap
-    to compute an interpolant using a SAT solver.
+    employing interpolation instead. The benefit of interpolation is
+    that it allows approximating the losing states efficiently by obtaining 
+    an interpolant from a SAT solver.
 \end{exmpInt}

### 2016-05-08 paper.tex
Changes to section 2 example

@@
                     }
-                node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                node [left=4pt] {\texttt{gr0=1 gr1=0}}
                 }
@@
     \end{minipage}
-        \caption{Environment winning trace}
+        \caption{Controller winning trace}
         \label{fig:trace}
     \end{subfigure}
-    \begin{subfigure}[t]{.15\textwidth}
+    \begin{subfigure}[t]{.2\textwidth}
         \centering
@@
                 child {node [circle,draw] {}
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
+                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
                 }
@@
                     }
-                node [left=4pt] {\texttt{gr0=1 gr1=1}}
-                }
-                node [left=4pt] {\texttt{req=1}}
+                node [left=4pt] {\texttt{gr0=1 gr1=0}}
+                }
+                node [left=4pt] {\texttt{req=2}}
                 node [above=4pt] {$\langle s, 3 \rangle$};
@@
         \end{minipage}
-        \caption{Controller winning trace}
+        \caption{Environment winning trace}
         \label{fig:trace2}
     \end{subfigure}
-    \begin{subfigure}[t]{.15\textwidth}
+    \begin{subfigure}[t]{.2\textwidth}
         \centering
@@
                 child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        % node [left=4pt] {\texttt{gr0=1 gr1=0}}
-                        edge from parent [fixed] node [left] {\texttt{req=2}}
-                    }
-                    % node [left=4pt] {\texttt{gr0=0 gr1=1}}
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
-                }
-%%%                child {node [circle,draw] {}
-%%%                    node [left=4pt] {\texttt{req=1}}
-%%%                    edge from parent [fixed] node [right] {\texttt{gr=1}}
-%%%                }
-                node [left=4pt] {}
+                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
+                }
+                node [left=4pt] {\texttt{req=2}}
                 node [above=4pt] {$\langle s, 3 \rangle$};
@@
         \end{minipage}
-        \caption{AGT}
-        \label{fig:agt2}
-    \end{subfigure}
-    \begin{subfigure}[t]{.2\textwidth}
-        \centering
-        \begin{minipage}[t][3.8cm][t]{\textwidth}
-        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
-            \node [circle,draw] (root){}
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [right] {\texttt{req=2}}
-                        }
-                        edge from parent [fixed] node [right] {\texttt{req=2}}
-                    }
-                    edge from parent [fixed] node [left] {\texttt{req=1}}
-                }
-                child {node [circle,draw] {}
-                    child {node [circle,draw] {}
-                        child {node [circle,draw] {}
-                            edge from parent [fixed] node [right] {\texttt{req=2}}
-                        }
-                        edge from parent [fixed] node [right] {\texttt{req=2}}
-                    }
-                    edge from parent [fixed] node [right] {\texttt{req=2}}
-                }
-                node [above=4pt] {$\langle s, 3 \rangle$};
-        \end{tikzpicture}
-        \end{minipage}
-        \caption{Refined AGT}
-        \label{fig:refined}
+        \caption{Partial Strategy}
+        \label{fig:strategy}
     \end{subfigure}
@@
     We present the intuition behind our bounded synthesis method by applying
-    its \emph{simplified version} to the running example.
-    We begin by finding 
-    a trace of length $k$ (here we consider $k=3$) that is winning for the environment, 
-    i.e., that starts from the initial state and reaches an
-    error state (see Figure~\ref{fig:trace}).  We use a SAT solver to find such a trace,
-    precisely as one would do in bounded model checking.
-    Given this trace we make an initial
-    conjecture that any trace starting with action \texttt{req=1} is winning for the
-    environment. This conjecture is captured in the abstract game tree shown in
-    Figure~\ref{fig:agt}.  We validate this conjecture by searching for a counterexample
-    trace that avoids the error state with the
-    first environment action fixed to \texttt{req=1}.   
-    Such a trace, that refutes the conjecture, is shown in Figure~\ref{fig:trace2}. 
-    In this trace, the controller wins by responding with \texttt{gr0=1, gr1=1} in the first round.
-    This move represents controller's partial strategy against the 
-    abstract game tree in Figure~\ref{fig:agt}.  This partial strategy is shown in Figure~\ref{}.
+    its \emph{simplified version} to the running example.  We begin by finding
+    a trace of length $k$ (here we consider $k=3$) that is winning for the
+    controller, i.e., that starts from the initial state and avoids the error
+    set for three game rounds (see Figure~\ref{fig:trace}).  We use a SAT
+    solver to find such a trace, precisely as one would do in bounded model
+    checking.  Given this trace we make an initial conjecture that any trace
+    starting with action \texttt{gr0=1 gr1=0} is winning for the controller.
+    This conjecture is captured in the abstract game tree shown in
+    Figure~\ref{fig:agt}.  We validate this conjecture by searching for a
+    counterexample trace that reaches an error state with the first controller
+    action fixed to \texttt{gr0=1 gr1=0}.   Such a trace, that refutes the
+    conjecture, is shown in Figure~\ref{fig:trace2}.  In this trace, the
+    environment wins by playing \texttt{req=2} in the first round.  This move
+    represents the environment's partial strategy against the abstract game
+    tree in Figure~\ref{fig:agt}.  This partial strategy is shown in
+    Figure~\ref{fig:strategy}.
 
     Next we strengthen the abstract game tree taking this partial strategy into account.
-    To this end we again use a SAT solver to find a trace where the environment wins while
-    the controller plays according to the partial strategy.  In the resulting trace (now shown), 
-    the environment plays \texttt{req=2} in the second round.  We refine the abstract game tree
-    using this move as shown in Figure~\ref{fig:agt2}.  The controller's partial strategy was to grant two
-    resources in the first round, to which the environment responds by now
-    making two requests in the second round.
-
-    When the environment cannot refine the tree by extending existing branches, it backtracks 
-    and creates new branches.
-    Eventually, we obtain the abstract game tree shown in Figure~\ref{fig:refined} for which 
-    there does not exist a winning partial strategy on behalf of the controller.  We conclude that 
-    the game is winning for the environment.
-
-%    Eventually, the refinements to the environment's candidate will force the
-%    controller to grant more resources and result in a game tree such as in
-%    Figure~\ref{fig:refined}.
+    To this end we again use a SAT solver to find a trace where the contoller
+    wins while the environment plays according to the partial strategy.  In the
+    resulting trace (Figure~\ref{fig:trace3}), the controller plays \texttt{gr0=1 gr1=1} in
+    the second round.  We refine the abstract game tree using this move as
+    shown in Figure~\ref{fig:refined1}.  The environment's partial strategy was
+    to make two requests in the first round, to which the controller responds
+    by now granting an additional two resources in the second round.
+
+    When the controller cannot refine the tree by extending existing branches,
+    it backtracks and creates new branches. Eventually, we obtain the abstract
+    game tree shown in Figure~\ref{fig:refined2} for which there does not exist
+    a winning partial strategy on behalf of the environment.  We conclude that
+    the bounded game is winning for the controller.
 
 \end{exmpI}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.2\textwidth}
+        \centering
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \begin{tikzpicture}[level distance = 5mm,baseline]
+            \node [circle,draw,inner sep=1pt] (root){}
+                child {node [circle,draw,inner sep=1pt] {}
+                    child {node [circle,draw,inner sep=1pt] {}
+                        child {node [circle,draw,inner sep=1pt] {}
+                            child {node [circle,draw,inner sep=1pt] {}
+                                child {node [circle,draw,inner sep=1pt] {}
+                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
+                                }
+                                node [left=4pt] {\texttt{req=1}}
+                            }
+                            node [left=4pt] {\texttt{gr0=1 gr1=1}}
+                        }
+                        node [left=4pt] {\texttt{req=1}}
+                    }
+                node [left=4pt] {\texttt{gr0=1 gr1=0}}
+                }
+                node [left=4pt] {\texttt{req=2}}
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \end{minipage}
+        \caption{Controller winning trace}
+        \label{fig:trace3}
+    \end{subfigure}
+    \begin{subfigure}[t]{.3\textwidth}
+        \centering
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \hspace*{0.8cm}
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
+                }
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \end{minipage}
+        \caption{First refined AGT}
+        \label{fig:refined1}
+    \end{subfigure}
+    \begin{subfigure}[t]{.4\textwidth}
+        \centering
+        \begin{minipage}[t][3.8cm][t]{\textwidth}
+        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
+            \node [circle,draw] (root){}
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
+                        }
+                        edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
+                    }
+                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
+                }
+                child {node [circle,draw] {}
+                    child {node [circle,draw] {}
+                        child {node [circle,draw] {}
+                            edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
+                        }
+                        edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
+                    }
+                    edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
+                }
+                node [above=4pt] {$\langle s, 3 \rangle$};
+        \end{tikzpicture}
+        \end{minipage}
+        \caption{Final Refined AGT}
+        \label{fig:refined2}
+    \end{subfigure}
+    \caption{Refined abstract game trees.}
+    \label{fig:refinedtrees}
+\end{figure}
 

### 2016-05-08 paper.tex
Edits to results table

@@
 were run on a cluster of Intel Quad Core Xeon E5405 2GHz CPUs with 16GB of
-memory.  The solver was allowed exclusive access to a node for one hour to
+memory.  The solvers were allowed exclusive access to a node for one hour to
 solve an instance.
@@
 run on Intel Quad Core 3.2Ghz CPUs with 32GB of memory, also on isolated nodes
-for one hour per instance. The unique column lists the number of instances that
-only that tool could solve in the competition (excluding our solver). In
-brackets is the number of instances that only that tool could solve, including
-our solver.
+for one hour per instance. The competition results differ significantly from
+our own benchmarks due to this more powerful hardware.  For our benchmarks we
+report only the results for solvers we were able to run on our cluster. The
+unique column lists the number of instances that only that tool could solve in
+the competition (excluding our solver). In brackets is the number of instances
+that only that tool could solve, including our solver.
 
@@
     \setlength{\tabcolsep}{8pt}
-    \begin{tabular}{l | r | r }
-        \textbf{Solver} & \textbf{Solved} & \textbf{Unique} \\
+    \begin{tabular}{l | r | r | r }
+        \textbf{Solver} & \textbf{Solved} & \textbf{Solved} & \textbf{Unique} \\
+                        & \textbf{(Competition)} & \textbf{(Benchmarks)} & \\
         \hline
-        Simple BDD Solver (2) & 195 & 10 \textit{(6)} \\
-        AbsSynthe (seq2) & 187 & 2 \\
-        Simple BDD Solver (1) & 185 & \\
-        AbsSynthe (seq3) & 179 & \\
-        Realizer (sequential) & 179 & \\
-        AbsSynthe (seq1) & 173 & 1 \\
-        Demiurge (D1real) & 139 & 5 \textit{(2)} \\
+        Simple BDD Solver (2) & 195 & 189 & 10 \textit{(6)} \\
+        AbsSynthe (seq2) & 187 & 139 & 2 \\
+        Simple BDD Solver (1) & 185 & 175 & \\
+        AbsSynthe (seq3) & 179 & 134 & \\
+        Realizer (sequential) & 179 & & \\
+        AbsSynthe (seq1) & 173 & 139 & 1 \\
+        Demiurge (D1real) & 139 & 136 & 5 \textit{(2)} \\
         Aisy & 98 & \\
-        \textit{Unbounded Synthesis} & \textit{103} & \textit{12} \\
+        \textit{Unbounded Synthesis} & & \textit{103} & \textit{12} \\
     \end{tabular}
@@
 
-We also present a cactus plot of the number of instances solved over time (Figure~\ref{fig:cactus}). For
-reference we plotted our results along with \textsc{Demiurge}~\cite{bloem2014},
-the only SAT-based tool in the competition, the winner of the sequential
-realisability track \textsc{Simple BDD Solver 2}~\cite{walker2014}, and
-AbsSynthe (seq3) \cite{brenguier2014}. Our results differ from the competition,
-we suspect due to the weaker hardware.
+We also present a cactus plot of the number of instances solved over time
+(Figure~\ref{fig:cactus}). We have plotted the best configuration of each
+solver we benchmarked. The solvers shown are
+\textsc{Demiurge}~\cite{bloem2014}, the only SAT-based tool in the competition,
+the winner of the sequential realisability track \textsc{Simple BDD Solver
+2}~\cite{walker2014}, and AbsSynthe (seq3) \cite{brenguier2014}.
 
@@
     \end{tikzpicture}
-    \caption{Number of instances solver over time by the unbounded solver, Demiurge,
-    and Simple BDD Solver.}\label{fig:cactus}
+    \caption{Number of instances solved over time.}
+    \label{fig:cactus}
 \end{figure}
